---
title: "1819_SEM"
author: "Shawna Sisler"
date: "6/23/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# LOAD PACKAGES
```{r Call packages}
library(assertive)
library(broom)
library(bmem)
library(car)
library(caret)
library(caTools)
library(clipr)
library(corrplot)
library(effectsize)
library(emmeans)
library(emulator)
library(expss)
library(FactoMineR)
library(factoextra)
library(fastDummies)
library(graphics)
library(gvlma) #
library(haven)
library(Hmisc)
library(janitor)
library(lavaan)
library(lattice)
library(leaps)
library(magrittr)
library(MBESS)
library(mediation)
library(missMDA)
library(mvord)
library(ordinal)
library(ordinalgmifs)
library(ordinalNet)
library(plspm)
library(pscl)
library(psych)
library(rcompanion)
library(readxl)
library(reshape2)
library(rstatix)
library(semPlot)
library(semTools)
library(stats)
library(stringr)
library(sure)
library(tidyverse)
library(tseries)
library(validate)
library(writexl)
```

# PREPARING FILES FOR SUBANALYSIS OF 18-19 YEARS OLD

```{r Filter sample, cache = TRUE}

#IMPORT file
# data.raw18 <- read_excel("~/Desktop/Sisler_dissertation_HMS1819_audit.xlsx",
#                        range=cell_cols(1:1820), col_types = "text")
#  
# save("data.raw18", file="~/Desktop/Sisler_dissertation_HMS1819_audit.RData")
load("~/Desktop/Sisler_dissertation_HMS1819_audit.RData")
 # source("dissertation_fx.R")
   
data.v1 <- data.raw18 %>%
  mutate(age = as.numeric(age)) %>%
  filter(age < 20 & age >= 18) %>%
  select(-c(

    # demographic info
    children_dep, educ_par1_rel, educ_par2_rel, hours_work_paid, timeclass, timestud, educ_par1, educ_par2, age_US, langpast, 
    langcur, age18:age41pl, educ_par, eductemp, educ_1:educ_4, age18_22, age31pl,
    # relationship info
    relship, relship_text, rel_sing:rel_divw,
    #  HMS-based
    SweepsIncentiveContact, endmessageSuicide, UniversityofSouthernCalifornia1a, consentFollowup1, excl_pers_oth_TEXT, 
    exp_rude,exp_accuse, exp_othersafraid, exp_ideaignore, exp_offendjoke, exp_stupid, exp_notserious, exp_superficial,
    exp_re_prof, exp_re_fear, exp_re_speakall,
    #  religious affiliations
    relig_aff_ag:relig_aff_oth, relig_aff_other_text, relig_1:relig_5,
    #  residence
    res_dorm:res_oth, residenc_text, deport_worry,
    #  degree categories
    degree_ass:degree_other_text, enroll_text, deg_ass:deg_nd,
    #  major/field choices
    field_hum:field_other, field_other_text,
    #  unused scales
    phq2_1, phq2_2,
    #  sex abuse/assault:
    ##  type
    assault_sex_stranger:assault_sex_dontknow, abuse_recent_perpet, sa_exp_touch:sa_exp_pen, 
    sa_suc_phyforc_1, sa_att_phyforc, sa_drunk_1, sa_suc_phyforc_2_1:sa_suc_phyforc_2_6, 
    sa_suc_phyforc_2_6_text, aca_phys_assault_1:aca_phys_assault_7, aca_sex_assault_1:aca_sex_assault_7,
    ##  reporting
    sa_tell_1:sa_tell_12, sa_tell_12_text, sa_nodiscl_1:sa_nodiscl_24, sa_nodiscl_24_text,
    ##  person
    sa_who_1:sa_who_12, sa_who_12_text, sa_who_stud, sa_who_employ, sa_who_gender, sa_who_gender_text,
    ##  location/context
    sa_loc_on, sa_op_substance_1:sa_op_substance_5, sa_frighten, sa_loc_off, sa_perc_1:sa_perc_12, 
    sa_train_pol, sa_train_prev,
    #  sleep variables
    sleep_wk1, sleep_wd1, sleep_wk2, sleep_wd2, sleep_np1, sleep_np2,
    #  stress variables
    stress1:compet_field,
    #  D/O SUBTYPES:
    ##  mood
    dx_dk, dx_dep_1:dx_bip_4, dep_maj, dep_oth, dx_bip_bi1:dx_bip_oth, dep_imp1:dep_imp4,
    ##  anxiety
    dx_ax_1:dx_ax_7_new, dx_ax_6_text_new, aca_anx_1:aca_add_7, dx_ax_ga:dx_ax_so,
    ##  OCD
    dx_ocd_1:dx_ocd_6, dx_ocd_6_text, dx_ocd_o:dx_ocd_oth,
    ##  trauma
    dx_trauma_1:dx_trauma_4, dx_trauma_4_text, dx_trauma_ptsd:dx_trauma_oth,
    ##  eating disorder
    dx_ea_1:dx_ea_4_text, ed_thin, weight_ideal, diet, ea_fast, purge_1:purge_4, ed_peer, 
    body_con_change, fresh15, weigh_freq, weight_ideal_text, dx_ea_an:dx_ea_av,
    ##  psychosis
    dx_psy_1:dx_psy_8, dx_psy_7_text, dx_psy_sa:dx_psy_sp,
    ##  substance-related
    dx_sa_1, dx_sa_2, dx_sa_3, dx_sa_2_text, aca_substance_1:aca_substance_7, sa_drunk_2_1:sa_drunk_2_7, 
    sa_drunk_2_6_text, sa_drunk_3:sa_drunk_6, alc_prob, sub_cig1:sub_cig5, sa_train_use:sa_sch_help,
    ##  personality d/o
    dx_perso_1:dx_perso_12, dx_perso_11_text,
    ##  chronic conditions
    aca_phys_health_1:aca_phys_health_7, dx_chronic_11_text,
    ##  SIBs
    sib_fre1:sub_cig,
    #  provider-related questions:  
    ther_helped_me, prov_1:prov_3, prov_7:prov_10, prov_9_text, anyprovi,
    #  insurance specifics
    ins_2:ins_10, ins_pare:ins_unce, ins_yesb, ins_mh1:ins_mh5, ins_ina_1:ins_ina_8_text,
    #  diet-related
    eatprac_1:eatprac_3, eatprac_frveg, eat_change, eat_change_how_1:eat_change_how_15, eat_change_how_15_text,
    #  SES
    fin_comp2, pay_worry, pay_fam, pay_self, pay_other, pay_grant, pay_loan, fincur_1:finpas_5,
    #  CCAPS
    CCAPS_1:CCAPS_34,
    #  CCMH:  
    sib_ccmh_1:suic_ccmh_6, cons_viol_ccmh_number, caus_viol_ccmh_number, exp_har_ccmh_number, 
    exp_trauma_ccmh_number, cons_viol_ccmh_recent, caus_viol_ccmh_recent, exp_har_ccmh_recent, 
    exp_trauma_ccmh_recent, trauma_ccmh_text, traumevent_ccmh_1, traumevent_ccmh_2, traumevent_ccmh_3, 
    traumevent_ccmh_4, traumevent_ccmh_5, traumevent_ccmh_6, traumevent_ccmh_7, traumevent_ccmh_8, 
    traumevent_ccmh_9, traumevent_ccmh_10, traumevent_ccmh_11, traumevent_ccmh_12, traumevent_ccmh_13, traumevent_ccmh_14, 
    traumevent_ccmh_15, traumevent_ccmh_16, traumevent_ccmh_15_text, binge_fr_ccmh,
    #  HMS-CCMH
    dep_CCMH_score_temp, dep_CCMH_score, dep_CCMH_elevated, dep_CCMH_mild, dep_CCMH_none, genanx_CCMH_score_temp,
    genanx_CCMH_score, genanx_CCMH_elevated, genanx_CCMH_mild, genanx_CCMH_none, socanx_CCMH_score_temp, 
    socanx_CCMH_score, socanx_CCMH_elevated, socanx_CCMH_mild, socanx_CCMH_none, eatconcerns_CCMH_score_temp,
    eatconcerns_CCMH_score, eatconcerns_CCMH_elevated, eatconcerns_CCMH_mild, eatconcerns_CCMH_none, dep_or_anx_CCMH,
    sib_ccmh_11:sib_ccmh_15, sib_ccmh_lifetime_any, sib_ccmh_pastyear,
    #  Qnum:  
    Q2_11, Q2_12, Q3_8, Q3_9, Q3_10_2, Q3_10_1, Q3_11_1, Q3_12_1:Q3_12_5, Q2_10_152:Q2_10_155, Q4_21_11, 
    Q9_19, Q9_20,
    #  health knowledge: 
    know_sp, dep_tx_know_1:dep_tx_know_4, dep_sx_know_1:dep_sx_know_4, anx_help_know_1:anx_help_know_4,
    ea_sx_know_1:ea_sx_know_6,
    #  therapy-related
    why_tx_1:why_tx_9_text, ther_lifetime, ther_lifetime2, ther_any, talk1_any, talk1_8_text, ther_any2, 
    knowwher1:knowwher6, bar_any:bar_plan, ther_help_me1:ther_help_me4, meds_help_me1:meds_help_me4, 
    talksup1:talksup4, talk2pro:talk2noo,
    #  school climate
    cli_disab:cli_sh_oth_TEXT, excl_env:exp_community, val_faculty:excl_pers_oth, group_TEXT:group_active,
    friend_re:friend_religid, social_re:social_religid, re_find:time_activ,
    #  SAT-related
    sat_hours_1, sat_loc_1, sat_qual_1, sat_priv_1, sat_sched_1, sat_hours_2, sat_loc_2, sat_qual_2, sat_priv_2, sat_sched_2, 
    sat_hours_3, sat_loc_3, sat_qual_3, sat_priv_3, sat_sched_3, sat_hours_4, sat_loc_4, sat_qual_4, sat_priv_4, sat_sched_4, 
    sat_hours_5, sat_loc_5, sat_qual_5, sat_priv_5, sat_sched_5, sat_hours_6, sat_loc_6, sat_qual_6, sat_priv_6, sat_sched_6, 
    sat_hours_7, sat_loc_7, sat_qual_7, sat_priv_7, sat_sched_7, sat_hours_8, sat_loc_8, sat_qual_8, sat_priv_8, sat_sched_8, 
    sat_hours_9, sat_loc_9, sat_qual_9, sat_priv_9, sat_sched_9, sat_hours_1_1:sat_hours_1_6, sat_loc_1_1:sat_loc_1_6, 
    sat_qual_1_1:sat_qual_1_6, sat_priv_1_1:sat_priv_1_6, sat_sched_1_1:sat_sched_1_6, sat_hours_2_1:sat_hours_2_6, 
    sat_loc_2_1:sat_loc_2_6, sat_qual_2_1:sat_qual_2_6, sat_priv_2_1:sat_priv_2_6, sat_sched_2_1:sat_sched_2_6, 
    sat_hours_3_1:sat_hours_3_6, sat_loc_3_1:sat_loc_3_6, sat_qual_3_1:sat_qual_3_6, sat_priv_3_1:sat_priv_3_6, 
    sat_sched_3_1:sat_sched_3_6, sat_hours_7_1:sat_hours_7_6, sat_loc_7_1:sat_loc_7_6, sat_qual_7_1:sat_qual_7_6, 
    sat_priv_7_1:sat_priv_7_6, sat_sched_7_1:sat_sched_7_6, sat_hours_8_1:sat_hours_8_6, sat_loc_8_1:sat_loc_8_6, 
    sat_qual_8_1:sat_qual_8_6, sat_priv_8_1:sat_priv_8_6, sat_sched_8_1:sat_sched_8_6,
    #  peer-related
    peer_alc_1, peer_alc_3, peer_alc_4, peer_mar_est, peer_alc_est, peer_cig_est, inf_7_text, risk_cig:risk_presc,
    #  pregnancy-related
    sex_partner, sex_partner_f, sex_partner_m, sex_partner_t, sex_30, sex_30_oral, sex_30_vag, sex_30_anal, 
    birthcontrol_1:birthcontrol_12, birthcontrol_10_text, birthcontrol_always, preg_no, preg_yes_u,preg_yes_in, 
    preg_dk, preg_cur,
    #  medication details
    meds_reason_1:meds_reason_5, meds_reason_5_text, meds_dis, meds_w_1:meds_w_5, meds_w_3_text, meds_cur_1:meds_cur_9, 
    meds_cur_7_text, meds_time_1:meds_time_7, med_help, meds_helped_me, meds_w_g:meds_di5,
    #  campus or community
    heard, camp_supp, outreach_aware, gk_1, gk_2, gkt_any, txfrf, cli_look, cli_resp_fr, timestud_1:timestud_7,
    cli_resp_cl, cli_free, st_promote, admin_listen, env_body, safe_on_day, safe_on_night, safe_off_day, 
    safe_off_night, choose_sch, persist_challenge_1:persist_challenge_11, persist_challenge_11_text, highestdeg,
    highestdeg_text, doubt_school_1, doubt_school_2, prof_support_aca, different_maj, late_assign, sleep_class,
    aca_support_serv, office_hrs, faculty_out, advisor_interact, gradstud_interact, time_manage, grade_curv,
    #  up/bystanding:
    inter_hy_dr, inter_hy_sa, inter_hy_hl, inter_y_1:inter_y_7, inter_y_6_text, inter_help, inter_n_1:inter_n_7, 
    inter_n_6_text, inter_n_why_1:inter_n_why_9, inter_n_why_9_text, witness_1:witness_7, witness_6_text, 
    inf_any:inf_oth,
    #  hypothetical:
    ther_help, talk2_1:talk2_7, talk2_7_text, talksup, percneed1:percneed_cur6, stig_t_I:stig_t_I6,
    #  misc:  
    bar_hs_1:bar_ns_8_text, smok_freq, sib_other_text, activ_other_text, knowwher_temp, exerc_changed, 
    exerc_changed_how, exerc_changed_how_text, height_ft, height_in, wgt_lbs, prov_A, prov_B, prov_C,  
    prov_loc, prov_olo, prov_oth, prov_dk, adjust_aca_1, adjust_aca_2, aca_imp1:aca_imp4, s_c_1_1:s_c_5_6, 
    s_n_1_1:s_n_5_6))
```

# Call functions
```{r Functions}
# delogodd
delogodd <- function(x) {
  oddsratio <- exp(x)
  p <- oddsratio / (1 + oddsratio)
  return(p)
}

# debug_contr_error
debug_contr_error <- function (dat, subset_vec = NULL) {
  if (!is.null(subset_vec)) {
    ## step 0
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
      }
      subset_log_vec <- subset_vec
    } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
      } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
      } 
    } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
    }
    dat <- base::subset(dat, subset = subset_log_vec)
  } else {
    ## step 1
    dat <- stats::na.omit(dat)
  }
  if (nrow(dat) == 0L) warning("no complete cases")
  ## step 2
  var_mode <- sapply(dat, mode)
  if (any(var_mode %in% c("complex", "raw"))) stop("complex or raw not allowed!")
  var_class <- sapply(dat, class)
  if (any(var_mode[var_class == "AsIs"] %in% c("logical", "character"))) {
    stop("matrix variables with 'AsIs' class must be 'numeric'")
  }
  ind1 <- which(var_mode %in% c("logical", "character"))
  dat[ind1] <- lapply(dat[ind1], as.factor)
  ## step 3
  fctr <- which(sapply(dat, is.factor))
  if (length(fctr) == 0L) warning("no factor variables to summary")
  ind2 <- if (length(ind1) > 0L) fctr[-ind1] else fctr
  dat[ind2] <- lapply(dat[ind2], base::droplevels.factor)
  ## step 4
  lev <- lapply(dat[fctr], base::levels.default)
  nl <- lengths(lev)
  ## return
  list(nlevels = nl, levels = lev)
}

# debug_contr_error2
debug_contr_error2 <- function (form, dat, subset_vec = NULL) {
  ## step 0
  if (!is.null(subset_vec)) {
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
      }
      subset_log_vec <- subset_vec
    } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
      } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
      } 
    } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
    }
    dat <- base::subset(dat, subset = subset_log_vec)
  }
  ## step 0 and 1
  dat_internal <- stats::lm(form, data = dat, method = "model.frame")
  attr(dat_internal, "terms") <- NULL
  ## rely on `debug_contr_error` for steps 2 to 4
  c(list(mf = dat_internal), debug_contr_error(dat_internal, NULL))
}

# NA_preproc
NA_preproc <- function (dat) {
  for (j in 1:ncol(dat)) {
    x <- dat[[j]]
    if (is.factor(x) && anyNA(x)) dat[[j]] <- base::addNA(x)
    if (is.character(x)) dat[[j]] <- factor(x, exclude = NULL)
  }
  dat
}

# fill.na.with.mode
## Model fitting per group and NA as factor levels
fill.na.with.mode = function(df){
  cols = colnames(df)
  for (col in cols){
    if(class(df[[col]])=='factor'){
      x = summary(df[[col]])
      mode = names(x[which.max(x)])
      df[[col]][is.na(df[[col]])]=mode
    }
    else{
      df[[col]][is.na(df[[col]])]=0
    }
  }
  return (df)
}

  # SEM visualization
plot_matrix <- function(matrix_toplot){
corrplot::corrplot(matrix_toplot, is.corr = FALSE,
               type = 'lower',
               order = "original",
               tl.col='black', tl.cex=.75)
  }

```


# OUTCOME VARIABLE

SUICIDAL IDEATION (si_type)
Creates indicator variable 'si_type' using the subtypes of suicidal ideation using sui_idea, sui_plan, sui_attempt, where: 
  0/0/0 = no SI
  1/0/0 = passive SI
  1/1/0 = active SI
  1/1/1 = volitional
+/- 'phq9_9' and 'sui_idea' into one variable -- suic_idea. Will run the end model twice.


```{r si_type}
data.v2 <- data.v1 %>%

# create si_type var
  mutate(si_type = case_when((sui_idea == 1 & suic_plan == 1 & suic_att == 1) ~ 4, # suicide attempt
                           (sui_idea == 1 & suic_plan == 1) ~ 3,                   # active SI
                           (sui_idea == 1) ~ 2,                                    # passive SI
                           sui_idea == 0 ~ 1)) %>%                                 # no SI reported
  filter(!is.na(si_type))   # filter for SI variable -- then rowwise deletion
data.v2$si_type <- as.factor(data.v2$si_type)  
  # tabulate and check output
table(data.v2$si_type)
table(data.v2$sui_idea)

#data.v2 = subset(data.v2, select= -c(suic_plan, sui_plan, suic_att, sui_att)) # clean up vars/columns

table(data.v2$si_type) # ensuring that all NAs are now clean
data.v2$si_type <- as.factor(data.v2$si_type)
plot(data.v2$si_type)
```


# DIATHESIS VARIABLES (.v3 - .v11)

Candidate variables: age, sexual, gender, finpast, sib_freq, sib, bmi, race, ed_scoff, body_sr, audit_tot, dx_dep, dx_any, dx_bi, dx_psy, dx_ea, dx_tr, dx_sa, dx_adhd, dx_pers, dx_none, med_any, med_sti, med_dep, med_psy, med_anx, med_mood, med_sle, meds_other, meds_none, alc_tx, wcs_tot, med_count, dx_chronic, disab_1, disab_2
```{r Diathesis frame}
data.v2 <- data.v2
```

## Age
```{r age}
data.v2$age <- as.numeric(data.v2$age)
hist(data.v2$age)
table(data.v2$age)
```

## Gender identity
```{r gender}
data.v3 <- data.v2 %>% 
  # Create 'gender' subtypes into one category
  mutate(gender = case_when((gender == 3 | gender == 4) ~ 3,     # transgender
                            (gender == 5 | gender == 6) ~ 2,     # gender diverse
                            (gender == 1 | gender == 2) ~ 1,     # cis-gender
                            FALSE ~ 4))                          # sets up non-answers
table(data.v3$gender)
  
  # transform into cisgender and non-cisgender
data.v3 <- data.v3 %>%  
  mutate(gender_tgd = case_when(gender == 2 ~ 1,
                                gender == 3 ~ 1,
                                gender == 1 ~ 0)) %>%
  filter(!is.na(gender))
data.v3$gender_tgd <- as.logical(data.v3$gender_tgd)


table(data.v3$gender_tgd, useNA = "always")
```

### Reconcile 'gender_text' 
```{r gender_text}
  # gender-diverse
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]emi")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]inary")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[N|n]eutral")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[Q|q]ueer")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[F|f]luid")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]gender")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]ndrogynous")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[P|p]angender")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]igender")] <- 1
data.v3$gender[data.v3$gender_text %in% c("she/they", "Closeted genderfluid (female presenting)", "Pan-gender",
                                          "female / infinite star being / gender non conforming","He or they",
                                          "Not sure, nonbinary I guess", "Non-binary trans maculine", "No preference", 
                                          "Somewhere between GNC and transfemme", "She/they",
                                          "Non-binary leaning towards trans man")] <- 1

 #transgender
data.v3$gender[data.v3$gender_text %in% c("Transmasculine", "Trans-masculine person", "Trans Masculine", 
                                          "gender non-conforming trans man")] <- 1

  # male
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]ude")] <- 0
data.v3$gender[str_detect(data.v3$gender_text, "[M|m]ale")] <- 0
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]oy")] <- 0
data.v3$gender[data.v3$gender_text %in% c("Male: there are only 2 genders", "if you have a penis you're a male", 
                                          "Cisgender man", "Male motherfucker. stop making up gender shit.", "He-male",
                                          "I have male parts so therefore I am a MALE and there are only TWO genders!!!!!!!!!!!!!!!!!",
                                          "SATSYG (Savage about to steal yo girl", "My sex is male", 
                                          "I say male, but I think gender is fluid", "Theres only two genders, I'm male",
                                          "Probably heterosexual, but so commitment-phobic that I can't actually tell if I'm asexual.",
                                          "I'm a male. Born a male, always will be a male. There should only be two choices.")] <- 0

  # female
data.v3$gender[str_detect(data.v3$gender_text, "[F|f]emale")] <- 0
data.v3$gender[str_detect(data.v3$gender_text, "[W|w]oman")]  <- 0
data.v3$gender[str_detect(data.v3$gender_text, "[G|g]irl")]  <- 0
data.v3$gender[data.v3$gender_text %in% c("Male and Female are not genders, they are sexes. I identify as a woman.", 
                                          "Female, but currently going through a gender identity crisis!")] <- 0

  # questioning/other
data.v3$gender[str_detect(data.v3$gender_text, "[Q|q]uestioning")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[U|u]nsure")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "[I|i]dk")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "Hermaphrodite")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "Questioning")] <- 1
data.v3$gender[str_detect(data.v3$gender_text, "Undecided")] <- 1
data.v3$gender[data.v3$gender_text %in% c("Literally no clue as of late", "Kynigender", "Adrogynous", "Hermaphrodite", 
                                          "idk yet", "Questioning", "Questioning; female", "Still questioning", "Not sure",
                                          "not female/questioning", "Questioning (unsure)", "Girl ? Idk", "not sure", 
                                          "Not Sure???", "I don't know, still trying to purse it out", "Undecided")] <- 1

  # non-answers jokesters
data.v3$gender[str_detect(data.v3$gender_text, "[H|h]elicopter")] <- 4
data.v3$gender[str_detect(data.v3$gender_text, "[C|c]hinook")] <- 4
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]pache")] <- 4
data.v3$gender[str_detect(data.v3$gender_text, "[U|u]nicorn")] <- 4
data.v3$gender[data.v3$gender_text %in% c("Albino stingray", "Karate Chop Action", "Flying Spaghetti Monster", "Pio Nono",
                                          "God", "None", "Baseball bat", "Mountain lion", "Oil Platform", "Chair", "Pizza", 
                                          "M1A1 Main Battle Tank", "Deppressed", "Fire hydrant/hockey puck", "Microwave",
                                          "This doesn't matter and you're an idiot for thinking it does.","REZZBIAN",
                                          "Baguette", "fairy", "a 5 year old girl", "tree", "unicorn", "His Majesty", 
                                          "Call of Duty: Modern Warfare 2", "Attack hellcopter", "Meme God", "AH-64D", 
                                          "Stainless Steel Toaster", "God himself", "baseball bat", "Mountain","Trans Former", 
                                          "Potato", "I will literally give you $100 to h*ck off", "Indian tape worm", 
                                          "Metallic Hyper-Entity", "2X4 treated lumber", "pizza", "sPecIaL SNowFlAkE", 
                                          "eh", "George Michael", "Metaknight Main", "Sea turtle", "I identify as a fruit cup",
                                          "The Senate", "1941 Nihon Giagun Yamato-Class Battleship", "Attack Hellicopter", 
                                          "penguin wearing a sombrero and sunglasses")] <- 4

  # non-answers, decline to answer
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]ecline")] <- "non-answers"
data.v3$gender[data.v3$gender_text %in% c("I am me. A human.", "I'm not yet sure where I identify", "Human-being", "genders", 
                                          "okay", "guess", "none", "Just my name.", "the above options are sex, not genders", 
                                          "the above options are sex, not genders", "its difficult", "any", "?", 
                                          "Biological male who identifies as a lesbian female.", "Prefer not to answer",
                                          "The universe conscious of itself",  "no",  "I only try to use gender when it's useful", 
                                          "fuck gender", "Gender and sex are the same", "gender is a lie", "#love", "No",
                                          "it's weird to separate out trans male and trans female if you're also asking for birth assignment.", 
                                          "2 genders", "I don't understand the question.", "Prefer not to specifiy",
                                          "'Male' and 'female' are not gender identities.",
                                          "I don't have one because it's not a legitimate concept.", 
                                          "Gender is a construct which serves no purpose.")] <- 4

data.v3 <- data.v3 %>%
  filter(!is.na(gender_tgd))
table(data.v3$gender_tgd, useNA = "always")   # check output
```

## Sexual Identity
```{r sexual}
data.v4 <- data.v3 %>%
  # aggregates all 'sexual' answers into one var
  mutate(
    sexual = case_when(sexual_queer == 1 ~ 5,                        # queer
                       sexual_bi == 1 ~ 4,                           # bisexual
                       sexual_l == 1 ~ 2,                            # lesbian 
                       sexual_g == 1 ~ 3,                            # gay
                       sexual_quest == 1 ~ 6,                        # questioning
                       sexual_h == 1 ~ 1,                            # heterosexual
                       sexual_other == 1 ~ 7,                        # other
                       FALSE ~ 8,                                    # non-answers
                       FALSE ~ 9,                                    # creating 'pansexual' var
                       FALSE ~ 10))                                  # creating 'asexual/demi' var

table(data.v4$sexual) # tabulate and check outputs
```

### Reconcile sexual_text
```{r sexual_text}
# identifying non-answers
  ## jokesters
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]elicopter")] <- 8
data.v4$sexual[data.v4$sexual_text %in% c("Yes", "i identify as a lamp", "I only like baseballs", "helisexual", "Autosexual",
                                          "now", "Miguel Garcia", "toasteroven", "must have gills", "Potsexual", "Bobcats",
                                          "Tailpipes turn me on", "Tyrannosaurus rex", "Traps are not gay", "Toaster sexual",
                                          "Crooked", "Pantsexual", "FREE", "Frozen Foods", "Giraffe", 
                                          "I'm sexually attracted to fire",  "Attracted to Vehicles", "sPecIaL SNowFlAkE",
                                          "Attracted to other self-identifying flying spaghetti monsters",  
                                          "Demiqueer quasiflux asexual foxkin", "I sleep around. Also into bestiality",  
                                          "Sexually attracted to kitchen appliances", "Elitistual. I'm into elitists",
                                          "sexually attracted to fictional characters and men twice my age")] <- 8
  ## declined to answer
  data.v4$sexual[str_detect(data.v4$sexual_text, "identi[t|f]y")] <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "label")] <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "matter")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[W|w]hatever")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "box myself")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]ecline")] <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]ormal")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]isclose")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]one")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[U|u]nknown")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]/[A|a]")]  <- 8
  data.v4$sexual[str_detect(data.v4$sexual_text, "[I|i] like")]  <- 8
  data.v4$sexual[data.v4$sexual_text %in% c("dating a trans male", "Attracted to personalities, not genders", "non",
                                            "irrelevant", "i like people", "Fuck sexuality", "personality based", 
                                            "Don't think about it too much", "Anyone that I like", "leave me alone",
                                            "define", "This is more bullshit", "No", "Me", "I am who I am", "Just there",
                                            "Don't care", "no label", "No Label", "No labels", "No", "Me",
                                            "Questions like this are the reason we are even having this debate.", 
                                            "i like who i like", "prefer not", "I do not understand why this is necessary",
                                            "Just there", "non", "Normal", "Human Being", "human being", "Closeted at home", 
                                            "It doesn't matter my question is why is this an actual question", 
                                            "I hate this PC bs", "FUCK YOU BIG TIME", "Polyamorous", "gay leaning bi",
                                            "Sexuality shouldn't matter", "No preference", "i like boys.", 
                                            "other", "Prefer not to specify", "i fall in love with a human being",
                                            "everyone's a little gay", "I prefer men, but have had women before", 
                                            "Traps are not gay", "Unknown", "#love", "Whatever I feel like at the time",
                                            "Situational",  "No one right now", "sometimes i dabble", "People", "Prefer not",
                                            "I am", "Whatever Happens", "Crooked", "love is love", "Open", "Autistic",
                                            "Human", "No one right now", "Situational", "I don't", "?", "Undefined", 
                                            "The universe conscious of itself", "Aromantic", "Repressed", "Never had sex",
                                            "Not dealing with men or women at the moment", "not sure", "Not sure",
                                            "None?", "neutral",  "no sexual orientation", "No idea, but I'm not straight",
                                            "Unlabeled", "Curious", "Anyone that I like", "Disinterested", "Do not define",
                                            "Fake Gay", "irrelevant", "People", "no particular sexual identity", 
                                            "No preference", "love is love", "there is no such thing", 
                                            "Sometimes gender doesn't matter and I don't have time to learn about these things life gets too busy",
                                            "Prefer not to answer", "just sexual (need no prefix!)", "It's complicated", 
                                            "Sexuality shouldn't matter", 
                                            "I'm just me and I love who I love. I don't like these labels.",
                                            "I'm not a fucking part of lgbtqrstuxyz bullshit association")]  <- 8
  
    # identifying queer  individuals:
data.v4$sexual[str_detect(data.v4$sexual_text, "[Q|q]ueer")] <- 5
data.v4$sexual[str_detect(data.v4$sexual_text, "[F|f]luid")] <- 5
data.v4$sexual[str_detect(data.v4$sexual_text, "[O|o]pen")] <- 5
data.v4$sexual[data.v4$sexual_text %in% c("potentially interested", "somewhere between", "homoflexible", "heteroflexible",
                                          "fluid", "Only other trans people", "non-binary", "so I guess I could check queer",
                                          "Androsexual", "Not straight", "Polysexual", "Joto", "Non-binary", "nonbinary",
                                          "Attracted to masculinity and androgony", "Androsexual")] <- 5

  # identifying questioning
data.v4$sexual[str_detect(data.v4$sexual_text, "[Q|q]uestion")] <- 6
data.v4$sexual[str_detect(data.v4$sexual_text, "[P|p]robably")] <- 6
data.v4$sexual[data.v4$sexual_text %in% c("possibly", "curious", "Unknown", "potentially interested", "don't know", "Unknown", 
                                          "don't know", "not sure", "still figuring it out", "idk", "Idk",  "don't know", 
                                          "I have a girlfriend but i am not attracted to any other females. I am attracted to males", 
                                          "Sometimes gender doesn't matter and I don't have time to learn about these things life gets too busy",
                                          "Not sure. Working on it.")] <- 6

  # identifying  pansexual write-ins
data.v4$sexual[str_detect(data.v4$sexual_text, "[P|p]an")] <- 9
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]apiosexual")] <- 9
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]apiosexual")] <- 9
data.v4$sexual[data.v4$sexual_text %in% c("all", "Polyamorous Pansexual", "Pan Sexual", "Pan-sexual", 
                                          "any but ive only had relationships with men")] <- 9

  # identifying asexual / demisexual write-ins:  
data.v4$sexual[str_detect(data.v4$sexual_text, "[G|g]r[a|e]y")] <- 10
data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]onsexual")] <-10
data.v4$sexual[str_detect(data.v4$sexual_text, "[C|c]elibate")] <- 10
data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]emi")] <- 10
data.v4$sexual[str_detect(data.v4$sexual_text, "[A|a]sex")] <- 10
data.v4$sexual[str_detect(data.v4$sexual_text, "[C|c]elibate")] <- 10
data.v4$sexual[str_detect(data.v4$sexual_text, "[A|a]ce")] <- 10
data.v4$sexual[data.v4$sexual_text %in% c("Mostly asexual", "Most likely asexual", "Masexual", "Tridium Vykon Jace 8000", 
                                          "Demi-romantic", "asexual lesbian", "Ase", "Bi-demisexual", "A sexual",
                                          "Hetroromantic Asexual", "Demiqueer quasiflux asexual foxkin", "DemiAsexual",
                                          "a-sexual", "no sexual attraction", "a sexual", "Homoromantic Demisexual",
                                          "Gay-ish, I am Demi-sexual which means I don't pay attention to sex/gender", 
                                          "experience no sexual attraction but do experience a small amount of romantic attraction", 
                                          "lithromantic", "Pan-Quoiromantic Asexual", "demiromantic asexual")] <- 10

  # identifying heterosexual individuals
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]etero")] <- 1
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]traight")] <- 1
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]etrosex")] <- 1
data.v4$sexual[data.v4$sexual_text %in% c("Kinsey scale 1", "I would put myself as a 1-1.5 on the kinsey scale", "Staight", 
                                          "Stragiht", "strait", "Str8!!!!", "Changing from gay to heterosexual",
                                          "Gynophilic", "Into females", "I fuck bitches.", "1 on the Kinsey Scale",
                                          "Gynophilic", "gynosexual", "Changing from gay to heterosexual", "100% Hetro")] <- 1

  # identified as gay or lesbian
data.v4$sexual[str_detect(data.v4$sexual_text, "[G|g]ay")] <- 3
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]omoflexible")] <- 3
data.v4$sexual[data.v4$sexual_text %in% c("Mostly gay", "soooooo gay", "Romantically straight and sexually gay")] <- 3
data.v4$sexual[str_detect(data.v4$sexual_text, "[L|l]esbian")] <- 2
data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]yke")] <- 2

  # identifying bisexual individuals
data.v4$sexual[str_detect(data.v4$sexual_text, "[B|b]isex")] <- 4
data.v4$sexual[str_detect(data.v4$sexual_text, "[B|b]icurious")] <- 4
data.v4$sexual[data.v4$sexual_text %in% c("I'm straight but I like girls too sexually, not emotionally.", "Half bisexual", 
                                          "straight but DTF sexy females", "bicurious but i like girls more", 
                                          "hetero with occasional bi tendencies", "gay leaning bi", "Bi-curious", 
                                          "Herero-romantic bisexual", "Bi with a stronger attraction to same sex",
                                          "I'm like halfway to bisexual. Never had a crush on another girl but I'm still interested. Idk.",
                                          "No boundaries in place - though probably bisexual")] <- 4

table(data.v4$sexual)   # tabulate and check outputs
```

### 'Sexual' transformed 
sexual identity: (1) heterosexual; (2) lesbian/gay/bisexual; (3) queer/pan/asex/other, and (4) non-answers
```{r sexual regroup}
data.v5 <- data.v4 %>%
    filter(!is.na(sexual)) %>% 
  # aggregates all 'sexual' answers into one var
  mutate(sexual = case_when((sexual == 9 | sexual == 10 | sexual == 5 | sexual == 6 | sexual == 4) ~ 3,   # queer, nonbinary
                             sexual == 6 ~ 3,  # queer/quest
                             sexual == 2 ~ 2,  # lesbian, gay
                             sexual == 3 ~ 2,  # bisexual
                             sexual == 1 ~ 1,  # heterosexual
                             sexual == 6 ~ 4)) %>%
  filter(!is.na(sexual))
data.v4$sexual <- as.factor(data.v4$sexual) # convert into factors

table(data.v5$sexual, useNA = "always") # tabulate and check outputs
```


## Race
```{r race}  
data.v6 <- data.v5 %>% 
# 'race' subtypes combined into one variable
  mutate(race = case_when(race_bla == 1 ~ 1,
                          race_ame == 1 ~ 5,
                          race_asi == 1 ~ 2,
                          race_his == 1 ~ 3,
                          race_pac == 1 ~ 5,
                          race_ara == 1 ~ 5,
                          race_whi == 1 ~ 4,
                          race_oth == 1 ~ 5,
                          FALSE ~ 5))
table(data.v6$race) # check output
data.v6 <- data.v6 %>%
  filter(!is.na(race))
table(data.v6$race, useNA = "always")
hist(data.v6$race)
```


### Race write-in reconcilation
Reconciles all 'other' answers vi race_other_text
```{r race reconciliation}
  #decline to answer
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]merica")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]uman")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]one")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]erson")] <- 5
data.v6$race[data.v6$race_other_text %in% c("prefer not to say", "Prefer not to answer.", "Prefer not to specify", 
                                            "Prefer not to answer :)", "FUCK YOU", "Caramel American", "Why", "irrelevant",
                                            "Prefer not to state", "Prefer no to say", "Brown", "No thx", "Texan", 
                                            "Race is a problematic social construct", "Race is a socual construct", 
                                            "Transcendent", "Race/ethnicity should not matter", "Prefer not to answer", 
                                            "why is this a thing", "White passing, ethnically other", "unfortunately", 
                                            "While I understand the socio-economic conditions that surround race, race also is a construct which serves no purpose.",
                                            "Wakanda Forever", "whatever I can use to be a victim in the given situation", 
                                            "We need to get rid of these color things",  "Me", "I do Not identify racially", 
                                            "Why is there always a race question? Can't we all just be one race? The Human Race.",  
                                            "Will not disclose",  "Race/ethnicity should not matter", "N/a", "N/A", 
                                            "There is only one human race: Homo sapien", "Citizen of the World",
                                            "don't know due to uncertainity of lineage", "I choose not to answer",
                                            "This doesn't matter and you're an idiot for thinking it does.",  
                                            "POC because now my opinion will matter in the US", 
                                            "Color doesn't mean anything right? Why do so many surveys ask it",
                                             "My skin is lighter than others, and darker than some", 
                                            "I am not a racial identifying person. I am American. My DNA says something about Korean heritage.",
                                             "I do not identify as a race in order to promote social equality for all races",
                                            "I don't think race is all that important.", 
                                            "I do not think that this information should or does make a difference ,so, I chose to not state", 
                                            "A White, Cherokee, European, Irish man who would be preferred to be called else but white",
                                            "I won't allow my race to be used in this study for potential insidious usage")]  <- 5

  # Jokesters
data.v6$race[data.v6$race_other_text %in% c("Volvo", "Vulcan", "Snowman", "tree", "General Electric", "Cat", "Earthling", 
                                            "I have rainbow skin.", "jedi", "Light skinned potato", "I'm stainless steel.", 
                                            "7-foot tall Chinese woman", "A little tan.", "Apache helicopter", "Aptenodytes",
                                            "Wood", "moon man", "Black hawk", "Aviation Heritage", "Ewok", "Aptenodytes",
                                            "Desert Camoflauge",  "Attack helicopter", "Ethnic Kekistani")] <- 5

  # white
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]ew")] <- 4  # if Israeli, could be White, Asian, or ME (e.g. Palestine)
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]shkenazi")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]aucasian")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]uropean")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]reek")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]rish")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]talian")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]lbania")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]em[e|i]tic")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[R|r]ussian")] <- 4 # could be Asian depending on region
data.v6$race[str_detect(data.v6$race_other_text, "[R|r]oma")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[T|t]urkish")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]lavic")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]etis")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]asque")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]icilian")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]olish")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]candinavian")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[F|f]innish")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]eltic")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]cottish")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]anad")] <- 4
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]editerranean")] <- 4
data.v6$race[data.v6$race_other_text %in% c("please refer to me as Caucasian; it's extremely offensive to refer to every race-- excluding Caucasians-- in their politically correct form",
                                            "Not sure if I'm Hispanic, as I am Portuguese and Spanish (from my mother's side) . I identify as white as I am of European descent.",
                                            "Portugese", "Cape Verdean", "Capeverdean", "Bulgarian-American", "Ginger", 
                                            "Anglo Saxon", "anglo/white", "Balkan for ethnicity", "Ulster-scots", 
                                            "I'm white, but I do have some Native American in me but it's not the dominant visible ethnicity.", 
                                            "White (Jewish/Jamaican-Chinese/Irish)", "ARMENIAN")] <- 4
  # Identifies as American Indian or other indigenous descent
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]ative [A|a]merican")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndigenous")] <- 5
data.v6$race[data.v6$race_other_text %in% c("I look white but im also status native", "Indigenous/Mestizo", 
                                            "american indian that isn'y legally claimed")] <- 5
  # Identifies as Asian
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]sian")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hinese")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]apanese")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndian")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]anglades")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[F|f]ilipino")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]uyanese")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ongolia")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hina")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]outheast")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]outh [A|a]sian")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]unjabi")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]iberia")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndonesia")] <- 2
data.v6$race[str_detect(data.v6$race_other_text, "[W|w]est [I|i]ndian")] <- 2
data.v6$race[data.v6$race_other_text %in% c("Sri Lankan", "Sri-Lankan", "Malays", "Chinese Ethinicy-Nuosu", "Austronesian",
                                            "third culture kid, born and raised in asia until 18yo", "Central Asian",  
                                            "Indian, UAE native", "Indian/Subcontinental", "Indians arent on the list.....", 
                                            "American Texan of Indian (South asia) Descent", "Okinawan", "Asian/Chinese",
                                            "Japanese American and Native Hawaiian Culture", "Asian. International.", 
                                            "Australian/Indian","Indo-Caribbean",  "Atlantic Islander", 
                                            "Telugu-speaking Tamil-ethnicity Indian")] <- 2
  # Identifies as Black
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]frican")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]fro")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[K|k]enya")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]lack")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]arribean")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]ahamian")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]aitian")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]aitian")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]amaica")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[Z|z]imbabwea")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]hana")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]thiopia")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[T|t]rinidad")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]ritrean")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[W|w]akanda")] <- 1
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]igerian")] <- 1
data.v6$race[data.v6$race_other_text %in% c("Cameroonian American", "Creole", "Jamacian American", "North Africa")] <- 1

  # Identifies as Hispanic
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]exican")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[X|x]ican[o|a]")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[L|l]atin[o|a]")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]uerto [R|r]ican")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]panish")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hican[o|a]")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]berian")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]estiz[o|a]")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]ortuguese")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]razil")] <- 3
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]uba")] <- 3
data.v6$race[data.v6$race_other_text %in% c("Costa Rican", "Central American Native", "Andean")] <- 3

  # Identifies as Middle Eastern or of Arab descent
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]rmenia")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]akistani")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]rab")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]ersian")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]alestin")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]srael")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]fghan")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]orth [A|a]frica")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]azara")] <- 5
data.v6$race[data.v6$race_other_text %in% c("Northern African; Moroccan; Israeli", "Arab Mix", "Berber/North African", 
                                            "Coptic Egyptian (non-Arab)", "Kekistani")] <- 5

  # Identifies all other data or mixed ethicity
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]iracial")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]i-racial")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ix")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]alf")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndo-[C|c]arribean")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ulti")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ulatto")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]anda")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[O|o]ther")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[U|u]nkown")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]mbiguous")] <- 5
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]mbiguous")] <- 5
data.v6$race[data.v6$race_other_text %in% c("Sub Asian/East African", "two or more races", "Moroccan Mexican",
                                            "Transracial", "Adopted", "Part Thetan", "Taiwanese-Greek-Cypriot American",
                                            "East Indian, Caribbean", "50/50", "Korean Ecuadorian", "Brazilian, Luso,
                                            Native American", "Central American/ Caribbean", "Asian/Indian",
                                            "Mut (more than 3)",  "mexican,filipino, italian", "Black Hispanic",
                                            "Middle Eastern, Filipino and Spanish",  "East Indian, Caribbean",
                                            "Indo-Portugesse (Fijian/Portugesse) Canadian Born", "Multiple",
                                            "White and American Indian (Choctaw)", "White, Black, and Hispanic",
                                            "White, black, Native American",  "French Lebanese", "Mixed Race", 
                                            "Lebanese, German, French", "Black/Asian", "Mixed", "Portuguese/Thai", 
                                            "An American Mutt- a bit of everything somewhere in there", 
                                            "Black and native", "Black/cuban",  "Filipino and Portuguese", "mixed race",
                                            "Eastern Indian/white", "white middle eastern", "Transracial",
                                            "German, Chinese, Arabic, Melanesian", "Hispanic and Middle Eastern",
                                            "mulatto", "mixed & south american",  "international", "Biracial",
                                            "Multiracial",  "Korean Ecuadorian", "Central American/ Caribbean",
                                            "An American Mutt- a bit of everything somewhere in there")] <- 5
  # Identifies as black and white as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("Black/White","White/African American", "White/African American")] <- 5
  # Identifies as Asian and White as 'Other/Mixed'
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]urasian")] <- 5
data.v6$race[data.v6$race_other_text %in% c("Japanese, Chinese, French, British, Greek", "Asian/White",
                                            "Asian/European", "Asian and White", "asian and white",  
                                            "White (Caucasian) and Indian (S. Asia)", "Chinese/White",
                                            "American/Asian African/American White", "3/4 White, 1/4 Japanese", 
                                            "White & Asian", "Chinese and White")] <- 5
  # Identifies as Hispanic and White as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("I am 1/4th Hispanic 3/4 German",  "Honduran, Irish", "White hispanic",
                                            "White, and Hispanic", "Hispanic/Jewish", "White and Hispanic", "White Hispanic",
                                            "French and spanish", "LATINA/WHITE", "white/hispanic",
                                            "If I identify as puerto rican does that mean i get more scholarship money?")] <- 5
	 # Identifies as MR and White as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("Half Middle Eastern half white", 
                                            "Half my family is white and the other middle eastern")] <- 5

# tabulate and check outputs
table(data.v6$race, useNA = "always")
data.v6 = subset(data.v6, select= -c(race_bla, race_asi, race_his, race_whi, race_other))   # clean up vars
hist(data.v6$race)

# consolidate race_ara, race_pi, race_ame --> other
```

## SIBs and frequency
``` {r sib_any}
data.v7 <- data.v6 %>%
  mutate(sib_freq = case_when(sib_none == 1 ~ 0,
                              sib_freq == 1 ~ 1,
                              sib_freq == 2 ~ 1,
                              sib_freq == 3 ~ 1,
                              sib_freq == 4 ~ 2,
                              sib_freq == 5 ~ 2,
                              sib_freq == 6 ~ 2)) %>%
  filter(!is.na(sib_freq))
data.v7$sib_freq <- as.numeric(data.v7$sib_freq)

table(data.v7$sib_freq, useNA = "always")
data.v7 = subset(data.v7, select= -c(sib_none:sib_oth)) # clean up vars/columns
hist(data.v7$sib_freq)
```

## Co-occuring MH diagnoses
```{r dx_mh}
# To ensure those who skipped the question are not accidentally added to No MH dx category
data.v8 <- data.v7

# set NAs to zero
data.v8$dx_dep[is.na(data.v8$dx_dep)] <- 0
data.v8$dx_anx[is.na(data.v8$dx_anx)] <- 0
data.v8$dx_bi[is.na(data.v8$dx_bi)] <- 0
data.v8$dx_psy[is.na(data.v8$dx_psy)] <- 0
data.v8$dx_ea[is.na(data.v8$dx_ea)] <- 0
data.v8$dx_tr[is.na(data.v8$dx_tr)] <- 0
data.v8$dx_sa[is.na(data.v8$dx_sa)] <- 0
data.v8$dx_neurodev_1[is.na(data.v8$dx_neurodev_1)] <- 0
data.v8$dx_pers[is.na(data.v8$dx_pers)] <- 0

data.v8$dx_adhd <- data.v8$dx_neurodev_1
data.v8$dx_adhd[is.na(data.v8$dx_adhd)] <- 0

data.v9 <- data.v8 %>%   # filters out any respondents who skipped the MH section 
  mutate(dx_mh = case_when(dx_dep == 1 ~ 1,
                           dx_anx == 1 ~ 1,
                           dx_bi == 1 ~ 1,
                           dx_psy == 1 ~ 1,
                           dx_ea == 1 ~ 1,
                           dx_tr == 1 ~ 1,
                           dx_sa == 1 ~ 1,
                           dx_adhd == 1 ~ 1,
                           dx_pers == 1 ~ 1,
                           dx_any == 1 ~ 1,
                           dx_none == 0 ~ 0)) %>%
  mutate(dx_mh = case_when(dx_mh == 1 ~ 1,
                           dx_mh == 0 ~ 0,
                           is.na(dx_mh) ~ 0))
data.v9$dx_mh <- as.logical(data.v9$dx_mh)

# Check output
table(data.v9$dx_mh, useNA = "always")
```

## Eating disorder, any type
```{r ed_any}
data.v10 <- data.v9 %>%
  filter(!is.na(ed_any)) %>%
  mutate(ed_any = case_when(ed_any == 1 ~ 1,
                            ed_any == 0 ~ 0))
data.v10$ed_any <- as.logical(data.v10$ed_any)

table(data.v10$ed_any, useNA = "always")
```

## Medication use
```{r meds}
# Create stimulant
data.v11 <- data.v10 %>%
  mutate(meds_sti = case_when(meds_1 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_1 == 0 ~ 0))
      # Reconcile stimulant write-ins from meds_7_text
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[V|v]yvan[s|c]e")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[T|t]enex")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[C|c]oncerta")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[A|a]dderall")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[R|r]italin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[P|p]rovigil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[A|a]rmodafinil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[F|f]ocalin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[M|m]etadat")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[F|f]ocalin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[N|n]uvigil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "ADHD")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[M|m]odafinil")] <- 1
      data.v11$meds_sti[data.v11$meds_7_text %in% c("Vyance", "Vivance adhd", "Stratera", "Aderol", "Aderal", "Adderal",
                                                    "Adderrall (for ADHD)", "Adderal, vyvance",  "Celltech")] <- 1
      
# Create antidepressant
data.v11 <- data.v11 %>%
  mutate(meds_dep = case_when(meds_2 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_2 == 0 ~ 0))
      # reconcile write-ins for depression from med_7_text
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]elexa")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]italopram")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]ymbalta")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[E|e]scitalopram")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[L|l]exapro")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[S|s]ertraline")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[Z|z]oloft")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "SSRI")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[E|e]lavil")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]aroxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]rozac")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]ntidepressant")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]mitriptyline")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]uloxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c][e|i]pralex")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[W|w]ellbutrin")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[S|s]trattera")] <- 1 
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]tomoxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[F|f]luvoxamine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]enlafaxine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]ristiq")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]esipramine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]iibryd")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]enlafaxin")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]lomipramine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]esvenflaxine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[M|m]irtazapine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[R|r][e|i]meron")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[T|t]raz[a|o|i]done")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[N|n]ortriptyline")] <- 1
      data.v11$meds_dep[data.v11$meds_7_text %in% c("CYMBALTA", "Cymbolta", "Lexipro", "Lexxapro", "Sertaline",
                                                    "remiron", "Mirtazepene", "bupropin", "Lexapro, Wellbutrin",
                                                    "Lexapro and Paxil", "Fuoxetine", "Duluxotine", 
                                                    "Forgot the name, anti depressant for anxiety, daily med", 
                                                    "Atomoxitine (nonstimulant ADHD)", "Fluvoximine (luvox)",
                                                    "Anafranil", "amatryptaline for migraines", "amnitryptaline", 
                                                    "Buproprion (may or may not count as a psychostim?)", 
                                                    "amatryptaline for migraines", "Minipress")] <- 1
      
# Create antipsychotic
data.v11 <- data.v11 %>%
  mutate(meds_psy = case_when(meds_3 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_3 == 0 ~ 0))
      # Reconcile anti-psychotic entries from med_7_text
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]bilify")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]bility")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[V|v]raylar")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[O|o]lanzapine")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]ripiprazole")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[L|l]atuda")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[S|s]eroqu[e|o]l")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[G|g]eodon")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[Q|q]uetiapine")] <- 1
      data.v11$meds_psy[data.v11$meds_7_text %in% c("Ziprazidone", "Quitiapin", "Latuda, Giadon")] <- 1 
      
# Create anti-anxiety  
data.v11 <- data.v11 %>%            
  mutate(meds_anx = case_when(meds_4 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_4 == 0 ~ 0))
      # Reconcile all anti-anxiety drugs from the write-ins from meds_7_text
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[X|x]anax")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[V|v]alium")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]ropranolol")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]ropanol")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[L|l]orazepam")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[N|n]eurontin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[A|a]tarax")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[L|l]yrica")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[A|a]tivan")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[C|c]lonidine")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]razosin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[B|b]uspar")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[V|v]istaril")] <- 1  # antihistamine
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[I|i]nderal")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[H|h]ydroxyzine")] <- 1 # antihistamine
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[K|k]lonopin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[D|d]iazepam")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[B|b]uspirone")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[G|g]abapentin")] <- 1
      data.v11$meds_anx[data.v11$meds_7_text %in% c("Valuim", "you didnt list vallium under anti-anxiety", "Alphrazolam", 
                                                   "Hydroxizine--prescribed as needed for anxiety, only took one because of adverse reaction", 
                                                   "Hydrodoxine, a blood pressure medicine for anxiety that I don't remember the name of")] <- 1

# Create mood stabilizer
data.v11 <- data.v11 %>%
  mutate(meds_mood = case_when(meds_5 == 1 ~ 1,
                               meds_8 == 1 ~ 0,
                               meds_5 == 0 ~ 0))
      # Reconcile mood stabilizers from write-in in med_7_text
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]ithium")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]amictal")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]amotrigine")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[D|d]epakote")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[T|t]rileptal")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[O|o]xcarbarzepine")] <- 1
      data.v11$meds_mood[data.v11$meds_7_text %in% c("Mood Stabilizer/Birth Control")] <- 1
      
# Create sleep med
data.v11 <- data.v11 %>%
  mutate(meds_sle = case_when(meds_6 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_6 == 0 ~ 0))
      # Reconcile sleep medications from write-ins in meds_7_text
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[X|x]yrem")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[S|s]leep")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[M|m]elatonin")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[Z|z]quil")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[N|n]yquil")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[D|d]iphenhydramine")] <- 1
      data.v11$meds_sle[data.v11$meds_7_text %in% c("gabapentin, trazodone (both for sleep on separate occasions)", 
                                                    "generic sleep medication: melatonin supplements", 
                                                    "cold medicine to induce sleep")] <- 1
      
# Create "Other" meds
data.v11 <- data.v11 %>%
  mutate(meds_oth = case_when(meds_7 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_7 == 0 ~ 0)) 
      # Reconcile "other" entries from med_7_text
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]annabis")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[M|m]arijuana")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[W|w]eed")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[N|n]altrexone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[S|s]ynthroid")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]opomax")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]estosterone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]hyroid")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[H|h]ydrodocone")] <- 1 
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]lonidine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[G|g]uanfacine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[K|k]etamine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[F|f]lexeril")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[P|p]ant[o|a]loc")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[S|s]pironolactone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[H|h]ormone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[O|o]xycodone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]annabic")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[B|b]eta")] <- 1
      data.v11$meds_oth[data.v11$meds_7_text %in% c("Canabis", "Vistiril", "Vistral", "Thyroxine", "The Reefer",
                                                    "Topimitrate", "topiramate", "Estradiol, Spironactone", 
                                                    "muscle relaxants to help me sleep", "Ondansetron" ,
                                                    "Sumatriptan for migraines", "Psilocybin Mushrooms", "clonodine",
                                                    "Muscle Relaxers", "anti-siezure medication", "Codeine", 
                                                    "Androgen Blockers", "Anti-seizure to control racing thoughts",
                                                    "Anti seizure", "Buprenorphine (Suboxone) for opiate dependence",
                                                    "Hydroxine", "Hydroxcyzine", "Alyacen", "Mirapex", 
                                                    "hypothyroid medication", "Epilepsy drugs", "Birth control",
                                                    "Methadone", "Vimpat", "Contrave", "Benztropine", "Intuniv",
                                                    "Medicinal marijuana", "Pantaloc, for psychogenic vomiting",
                                                    "medicinal Cannabis", "Testosterone Enanthate", "testosterone",
                                                    "Cannabis", "Medical marijuana")] <- 1
      
# Create a group for students who do not take any psychotropic meds
data.v11 <- data.v11 %>%
  mutate(meds_none = case_when(meds_none == 1 ~ 1,
                               meds_none == 0 ~ 0,
                               meds_sti == 1 ~ 0,
                               meds_dep == 1 ~ 0,
                               meds_psy == 1 ~ 0,
                               meds_anx == 1 ~ 0,
                               meds_mood == 1 ~ 0,
                               meds_sle == 1 ~ 0,
                               meds_oth == 1 ~ 0))

# tabulate and check outputs
table(data.v11$meds_sti)
table(data.v11$meds_dep)
table(data.v11$meds_psy)
table(data.v11$meds_anx)
table(data.v11$meds_mood)
table(data.v11$meds_sle)
table(data.v11$meds_oth)
table(data.v11$meds_none)

```

## Psychopharm medication count
Create 'meds_count' to be used in future regression to capture quantity and diversity of medications 
to see what is associated (i.e. the count is meant to find the common meds in suicidality)
```{r med_count}
data.v11$meds_1 <- as.numeric(data.v11$meds_1)
data.v11$meds_2 <- as.numeric(data.v11$meds_2)
data.v11$meds_3 <- as.numeric(data.v11$meds_3)
data.v11$meds_4 <- as.numeric(data.v11$meds_4)
data.v11$meds_5 <- as.numeric(data.v11$meds_5)
data.v11$meds_6 <- as.numeric(data.v11$meds_6)
data.v11$meds_7 <- as.numeric(data.v11$meds_7)
data.v11$meds_8 <- as.numeric(data.v11$meds_8)

data.v11$meds_1[is.na(data.v11$meds_1)] <- 0
data.v11$meds_2[is.na(data.v11$meds_2)] <- 0
data.v11$meds_3[is.na(data.v11$meds_3)] <- 0
data.v11$meds_4[is.na(data.v11$meds_4)] <- 0
data.v11$meds_5[is.na(data.v11$meds_5)] <- 0
data.v11$meds_6[is.na(data.v11$meds_6)] <- 0
data.v11$meds_7[is.na(data.v11$meds_7)] <- 0

# create meds_count var
data.v11 <- data.v11 %>%
  mutate(meds_count = meds_1 + meds_2 + meds_3 + meds_4 + meds_5 + meds_6 + meds_7) %>%
  mutate(meds_count = case_when(meds_count == 0 ~ 0,
                                meds_count == 1 ~ 1,
                                meds_count == 2 ~ 2, 
                                meds_count == 3 ~ 3,
                                meds_count == 4 ~ 3,
                                meds_count == 5 ~ 3,
                                meds_count == 6 ~ 3))
data.v11$meds_count <- as.numeric(data.v11$meds_count)

table(data.v11$meds_count, useNA = "always")  # check output
data.v11 = subset(data.v11, select= -c(meds_1:meds_7)) # clean up vars/columns
hist(data.v11$meds_count)
```

## DIATHESIS ANALYSES
```{r Diathesis analyses}
data.v12 <- data.v11 # create new dataframe for summary stats. Will jump to v13 for environmental
```

### Diathesis correlation matrix
```{r Diathesis correlation}

# Define diathesis vars for correlation matrix
data_dia <- data.v12 %>% select(age, sexual, gender_tgd, finpast, sib_freq, race, ed_scoff, body_sr,
                                dx_mh, meds_count, ed_any, meds_sti, meds_dep, meds_psy, meds_anx, meds_mood,
                                meds_sle, meds_oth)

# Corr_simple function
corr_simple <- function(data=data_dia, sig=0.6){
  # convert data to numeric in order to run correlations 
  data_dia <- data_dia %>% mutate_if(is.character, as.factor)
  data_dia <-  data_dia %>% mutate_if(is.factor, as.numeric)
  # run a correlation and drop the insignificant ones
  corr <- cor(data_dia, use = "pairwise.complete.obs")                               
  corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
  corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
  corr <- as.data.frame(as.table(corr))                   # turn into a 3-column table
  corr <- na.omit(corr)                                   # remove the NA values from above 
  corr <- corr[order(-abs(corr$Freq)),]                   # sort by highest correlation
  print(corr)
  
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}
corr_simple()

# Extracting overlapping vars to visualize correlation for regression
data_dia <- data.v12 %>% select(age, sexual, gender_tgd, finpast, sib_freq, race, 
                                      body_sr, ed_any, dx_mh, meds_count, meds_sti, meds_dep, 
                                      meds_psy, meds_anx, meds_mood, meds_sle, meds_oth)

# Corr_simple function
corr_simple <- function(data=data_dia, sig=0.6){
  # convert data to numeric in order to run correlations 
  data_dia <- data_dia %>% mutate_if(is.character, as.factor)
  data_dia <-  data_dia %>% mutate_if(is.factor, as.numeric)
  # run a correlation and drop the insignificant ones
  corr <- cor(data_dia, use = "pairwise.complete.obs")                               
  corr[lower.tri(corr,diag=TRUE)] <- NA                   # prep to drop duplicates and correlations of 1    
  corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
  corr <- as.data.frame(as.table(corr))                   # turn into a 3-column table
  corr <- na.omit(corr)                                   # remove the NA values from above 
  corr <- corr[order(-abs(corr$Freq)),]                   # sort by highest correlation
  print(corr)
  
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}
corr_simple()

```

### EFA diathesis
```{r EFA diathesis}
data_dia <- data.v12 %>% select(age, sexual, gender_tgd, finpast, sib_freq, race,
                                dx_mh, ed_any, meds_count)

data_dia <- na.omit(data_dia)

# create corr matrix, with dummy vars
data_dia <- data_dia %>% 
  mutate_if(is.numeric, as.factor) %>% 
  mutate(
    age = as.numeric(age),
    meds_count = as.numeric(meds_count),
  ) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_dia   # check output

results_dia <- dummy_cols(data_dia, select_columns = c("age", "sexual", "gender_tgd", "finpast", 
                                                       "sib_freq", "race", "dx_mh", "ed_any",
                                                       "meds_count"), 
                          remove_first_dummy = TRUE)

# factoextra -- https://rpkgs.datanovia.com/factoextra/
dia_pca <- PCA(results_dia, graph=FALSE)  # apply PCA
eigval_dia <- get_eigenvalue(dia_pca)
eigval_dia

fviz_screeplot(dia_pca, addlabels = TRUE, ylim = c(0, 50))

fviz_pca_var(dia_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(dia_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(dia_pca, choice = "var", axes = 1, 
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

### CLM diathesis regression
```{r clm_dia}
clm_dia <- data.v12 %>% select(si_type, age, sexual, gender_tgd, sib_freq, race,
                                  ed_any, dx_mh, meds_count)
clm_dia <- na.omit(clm_dia)

view(clm_dia)
dia_mod <- as.data.frame(clm_dia)

dia_mod <- clm(si_type ~ ., data = dia_mod)

summary(dia_mod)
coef(dia_mod)

ci_dia <- confint(dia_mod)
ci_dia

model.dia <- tidy(dia_mod)  # Convert model to dataframe for easy manipulation
model.dia
model.dia  %>%
   mutate(or = exp(estimate),  # Odds ratio/gradient
          var.diag = diag(vcov(dia_mod)),  # Variance of each coefficient
          or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# ENVIRONMENT VARIABLES
Candidate variables: res_status, military, activ_yn, residenc, discrim, env_mh, 
ther_vis, school2, medtherc, medther, talk, religios, residenc, satisfied_overall, firstgen


```{r Env frame}
data.v13 <- data.v11
```

## Religion
```{r religios}
# Label levels of religious importance
data.v13 <- data.v13 %>%
  filter(!is.na(religios))
data.v13$religios <- as.numeric(data.v13$religios)

table(data.v13$religios, useNA = "always")       # Check output
```

## Distress & social support
Group talk var into subcategories of (1) none,  (2) formal, (3) informal, or (4) combo of support
```{r talk}
data.v13$talk1_1[is.na(data.v13$talk1_1)] <- 0
data.v13$talk1_2[is.na(data.v13$talk1_2)] <- 0
data.v13$talk1_3[is.na(data.v13$talk1_3)] <- 0
data.v13$talk1_4[is.na(data.v13$talk1_4)] <- 0
data.v13$talk1_5[is.na(data.v13$talk1_5)] <- 0
data.v13$talk1_6[is.na(data.v13$talk1_6)] <- 0
data.v13$talk1_7[is.na(data.v13$talk1_7)] <- 0
data.v13$talk1_8[is.na(data.v13$talk1_8)] <- 0

# change talk_ var to numeric
data.v14 <- data.v13 %>% 
  # aggregate all talk vars under 3 'talk' categories -- none, informal help, formal help, formal+informal help
  mutate(talk = case_when(((talk1_1 == 1 | talk1_6 == 1 | talk1_7 == 1 | talk1_8 == 1) &        # combined
                             (talk1_2 == 1 | talk1_3 == 1 | talk1_4 == 1 | talk1_5 == 1)) ~ 4,
                          (talk1_9 == 1 | talk1no == 1) ~ 1,                                    # no support
                          (talk1_1 == 1 | talk1_6 == 1 | talk1_7 == 1 | talk1_8 == 1) ~ 2,      # formal support
                          (talk1_2 == 1 | talk1_3 == 1 | talk1_4 == 1 | talk1_5 == 1) ~ 3)) %>% # informal support
  mutate(talk = case_when(talk == 1 ~ 1,
                          talk == 2 ~ 2,
                          talk == 3 ~ 3,
                          talk == 4 ~ 4)) %>%
  filter(!is.na(talk))
data.v14$talk <- as.factor(data.v14$talk)

table(data.v14$talk, useNA = "always") # check output

data.v14 = subset(data.v14, select= -c(talk1_1:talk1_9, talk1sig, talk1fr, 
talk1rel, talk1no, talk1sup, talk1rm, talk1pro, talk1fam)) # clean up
```

## Overall school satisfaction
```{r satisfied_overall}
data.v14 <- data.v14 %>%
  mutate(satisfied_overall = case_when(satisfied_overall == 1 ~ 1,
                                       satisfied_overall == 2 ~ 2,
                                       satisfied_overall == 3 ~ 3,
                                       satisfied_overall == 4 ~ 4,
                                       satisfied_overall == 5 ~ 5,
                                       satisfied_overall == 6 ~ 6)) %>% 
  filter(!is.na(satisfied_overall))
data.v14$satisfied_overall <- as.numeric(data.v14$satisfied_overall)

table(data.v14$satisfied_overall, useNA = "always")
```

## Activity Involvement
```{r Activities}
# set all NAs in actvities equal to 0
data.v14$activ_ac[is.na(data.v14$activ_ac)] <- 0
data.v14$activ_athc[is.na(data.v14$activ_athc)] <- 0
data.v14$activ_athv[is.na(data.v14$activ_athv)] <- 0
data.v14$activ_athi[is.na(data.v14$activ_athi)] <- 0
data.v14$activ_cs[is.na(data.v14$activ_cs)] <- 0
data.v14$activ_cu[is.na(data.v14$activ_cu)] <- 0
data.v14$activ_da[is.na(data.v14$activ_da)] <- 0
data.v14$activ_fs[is.na(data.v14$activ_fs)] <- 0
data.v14$activ_gs[is.na(data.v14$activ_gs)] <- 0
data.v14$activ_gov[is.na(data.v14$activ_gov)] <- 0
data.v14$activ_mp[is.na(data.v14$activ_md)] <- 0
data.v14$activ_md[is.na(data.v14$activ_none)] <- 0
data.v14$activ_soc[is.na(data.v14$activ_soc)] <- 0
data.v14$activ_rel[is.na(data.v14$activ_rel)] <- 0
data.v14$activ_art[is.na(data.v14$activ_art)] <- 0
data.v14$activ_other[is.na(data.v14$activ_other)] <- 0

# Create one activity variable (dichotomous) based on involment in campus activities
data.v14 <- data.v14 %>% 
  mutate(activ_yn = case_when(activ_none == 1 ~ 0,
                              activ_ac == 1 ~ 1,
                              activ_athc == 1 ~ 1,
                              activ_athv == 1 ~ 1,
                              activ_athi == 1 ~ 1,
                              activ_cs == 1 ~ 1,
                              activ_cu == 1 ~ 1,
                              activ_da == 1 ~ 1,
                              activ_fs == 1 ~ 1,
                              activ_gs == 1 ~ 1,
                              activ_gov == 1 ~ 1,
                              activ_mp == 1 ~ 1,
                              activ_md == 1 ~ 1,
                              activ_rel == 1 ~ 1,
                              activ_soc == 1 ~ 1,
                              activ_art == 1 ~ 1,
                              activ_other == 1 ~ 1)) %>%
  filter(!is.na(activ_yn))
data.v14$activ_yn <- as.logical(data.v14$activ_yn)

table(data.v14$activ_yn, useNA = "always")     # tabulate and check output

#data.v14 = subset(data.v14, select= -c(activ_none, activ_ac:activ_other))    # clean up vars
```

## School variable
```{r school2}
# insert each school as a dichotomous var
data.v14 <- data.v14 %>%
  mutate(var_AlbertaCAD = ifelse((school2 == "AlbertaCAD"), 1, 0)) %>%
  mutate(var_Boston = ifelse((school2 == "Boston"), 1, 0)) %>%
  mutate(var_CIA = ifelse((school2 == "CIA"), 1, 0)) %>%
  mutate(var_CalPoly = ifelse((school2 == "CalPoly"), 1, 0)) %>%
  mutate(var_ColoradoCollege = ifelse((school2 == "ColoradoCollege"), 1, 0)) %>%
  mutate(var_ColoradoMountain = ifelse((school2 == "ColoradoMountain"), 1, 0)) %>%
  mutate(var_ColumbusCAD = ifelse((school2 == "ColumbusCAD"), 1, 0)) %>%
  mutate(var_Delaware = ifelse((school2 == "Delaware"), 1, 0)) %>%
  mutate(var_DetroitMercy = ifelse((school2 == "DetroitMercy"), 1, 0)) %>%
  mutate(var_Emerson = ifelse((school2 == "Emerson"), 1, 0)) %>%
  mutate(var_GeorgeMason = ifelse((school2 == "GeorgeMason"), 1, 0)) %>%
  mutate(var_GeorgiaTech = ifelse((school2 == "GeorgiaTech"), 1, 0)) %>%
  mutate(var_Kalamazoo = ifelse((school2 == "Kalamazoo"), 1, 0)) %>%
  mutate(var_Kansas = ifelse((school2 == "Kansas"), 1, 0)) %>%
  mutate(var_MDInstituteCollegeArt = ifelse((school2 == "MDInstituteCollegeArt"), 1, 0)) %>%
  mutate(var_MassArt = ifelse((school2 == "MassArt"), 1, 0)) %>%
  mutate(var_Merrimack = ifelse((school2 == "Merrimack"), 1, 0)) %>%
  mutate(var_MichiganState = ifelse((school2 == "MichiganState"), 1, 0)) %>%
  mutate(var_MinnesotaCAD = ifelse((school2 == "MinnesotaCAD"), 1, 0)) %>%
  mutate(var_MontclairState = ifelse((school2 == "MontclairState"), 1, 0)) %>%
  mutate(var_NHInstituteArt = ifelse((school2 == "NHInstituteArt"), 1, 0)) %>%
  mutate(var_NevadaReno = ifelse((school2 == "NevadaReno"), 1, 0)) %>%
  mutate(var_OaklandCC = ifelse((school2 == "OaklandCC"), 1, 0)) %>%
  mutate(var_OklahomaCityCC = ifelse((school2 == "OklahomaCityCC"), 1, 0)) %>%
  mutate(var_PaloAltoCollege = ifelse((school2 == "PaloAltoCollege"), 1, 0)) %>%
  mutate(var_PennCollegeTech = ifelse((school2 == "PennCollegeTech"), 1, 0)) %>%
  mutate(var_Pratt = ifelse((school2 == "RedlandsCC"), 1, 0)) %>%
  mutate(var_RhodeIsland = ifelse((school2 == "RhodeIsland"), 1, 0)) %>%
  mutate(var_RedlandsCC = ifelse((school2 == "RedlandsCC"), 1, 0)) %>%
  mutate(var_Ringling = ifelse((school2 == "Ringling"), 1, 0)) %>%
  mutate(var_Rollins = ifelse((school2 == "Rollins"), 1, 0)) %>%
  mutate(var_SAIC = ifelse((school2 == "SAIC"), 1, 0)) %>%
  mutate(var_Sewanee = ifelse((school2 == "Sewanee"), 1, 0)) %>%
  mutate(var_Smith = ifelse((school2 == "Smith"), 1, 0)) %>%
  mutate(var_SouthernNazarene = ifelse((school2 == "SouthernNazarene"), 1, 0)) %>%
  mutate(var_SouthwesternOklahomaState = ifelse((school2 == "SouthwesternOklahomaState"), 1, 0)) %>%
  mutate(var_StJohns = ifelse((school2 == "StJohns"), 1, 0)) %>%
  mutate(var_StMarys = ifelse((school2 == "StMarys"), 1, 0)) %>%
  mutate(var_StRose = ifelse((school2 == "StRose"), 1, 0)) %>%
  mutate(var_TrumanState = ifelse((school2 == "TrumanState"), 1, 0)) %>%
  mutate(var_TuftsHealthSciences = ifelse((school2 == "TuftsHealthSciences"), 1, 0)) %>%
  mutate(var_TuftsMainCampus = ifelse((school2 == "TuftsMainCampus"), 1, 0)) %>%
  mutate(var_UChicago = ifelse((school2 == "UChicago"), 1, 0)) %>%
  mutate(var_UFlorida = ifelse((school2 == "UFlorida"), 1, 0)) %>%
  mutate(var_UMDearborn = ifelse((school2 == "UMDearborn"), 1, 0)) %>%
  mutate(var_UMich = ifelse((school2 == "UMich"), 1, 0)) %>%
  mutate(var_UNCSchoolArt = ifelse((school2 == "UNCSchoolArt"), 1, 0)) %>%
  mutate(var_UTKnoxville = ifelse((school2 == "UTKnoxville"), 1, 0)) %>%
  mutate(var_UniversityofSouthernCalifornia = ifelse((school2 == "UniversityofSouthernCalifornia"), 1, 0)) %>%  
  mutate(var_Utah = ifelse((school2 == "Utah"), 1, 0)) %>%
  mutate(var_VATech = ifelse((school2 == "VATech"), 1, 0)) %>%
  mutate(var_VirginiaCommonwealth = ifelse((school2 == "VirginiaCommonwealth"), 1, 0)) %>%
  mutate(var_WakeForest = ifelse((school2 == "WakeForest"), 1, 0)) %>%
  mutate(var_Watkins = ifelse((school2 == "Watkins"), 1, 0)) %>%
  mutate(var_WestVirginia = ifelse((school2 == "WestVirginia"), 1, 0)) %>%
  mutate(var_WesternCarolina = ifelse((school2 == "WesternCarolina"), 1, 0)) %>%
  mutate(var_WesternMichigan = ifelse((school2 == "WesternMichigan"), 1, 0)) %>%
  mutate(var_WesternWA = ifelse((school2 == "WesternWA"), 1, 0)) %>%
  mutate(var_Xavier = ifelse((school2 == "Xavier"), 1, 0))
```

### School type levels

(1) 4-year: var_Boston, var_CalPoly, var_ColoradoCollege, var_ColoradoMountain, var_Delaware, var_DetroitMercy, var_Emerson, var_GeorgeMason, var_GeorgiaTech, var_Kalamazoo, var_Kansas, var_Merrimack, var_MichiganState, var_MontclairState, var_NevadaReno, var_PennCollegeTech, var_RhodeIsland, var_Rollins, var_Sewanee, var_Smith, var_SouthernNazarene, var_SouthwesternOklahomaState, var_StJohns, var_StMarys, var_StRose, var_TrumanState, var_TuftsHealthSciences, var_TuftsMainCampus, var_UChicago, var_UFlorida, var_UMDearborn, var_UMich, var_UTKnoxville,  var_UniversityofSouthernCalifornia, var_Utah, var_VATech, var_VirginiaCommonwealth, var_WakeForest,var_WestVirginia, var_WesternCarolina, var_WesternMichigan, var_WesternWA, var_Xavier

(2) Community College: var_OaklandCC, var_OklahomaCityCC, var_PaloAltoCollege, var_RedlandsCC 

(3) Art & Design, Culinary: var_AlbertaCAD, var_CIA, var_ColumbusCAD, var_MDInstituteCollegeArt, var_MassArt, var_MinnesotaCAD, var_NHInstituteArt, var_Pratt, , var_Ringling, var_SAIC, var_UNCSchoolArt, var_Watkins, 

```{r school2_type}
data.v14 <- data.v14 %>%
  # 4 year college
  mutate(school2_4yr = ifelse(var_Boston + var_CalPoly + var_ColoradoCollege + var_ColoradoMountain + var_Delaware +
                                var_DetroitMercy + var_Emerson + var_GeorgeMason + var_GeorgiaTech + var_Kalamazoo + 
                                var_Kansas + var_Merrimack + var_MichiganState + var_MontclairState + var_NevadaReno +
                                var_PennCollegeTech + var_RhodeIsland + var_Rollins + var_Sewanee + var_Smith +
                                var_SouthernNazarene + var_SouthwesternOklahomaState + var_StJohns + var_StMarys + var_StRose +
                                var_TrumanState + var_TuftsHealthSciences + var_TuftsMainCampus + var_UChicago + var_UFlorida +
                                var_UMDearborn + var_UMich + var_UTKnoxville + var_UniversityofSouthernCalifornia + var_Utah +
                                var_VATech + var_VirginiaCommonwealth + var_WakeForest + var_WestVirginia + 
                                var_WesternCarolina + var_WesternMichigan + var_WesternWA + var_Xavier == 1, TRUE,
                              ifelse(
                                var_Boston + var_CalPoly + var_ColoradoCollege + var_ColoradoMountain + var_Delaware +
                                  var_DetroitMercy + var_Emerson + var_GeorgeMason + var_GeorgiaTech + var_Kalamazoo +
                                  var_Kansas + var_Merrimack + var_MichiganState + var_MontclairState + var_NevadaReno +
                                  var_PennCollegeTech + var_RhodeIsland + var_Rollins + var_Sewanee + var_Smith +
                                  var_SouthernNazarene + var_SouthwesternOklahomaState + var_StJohns + var_StMarys + 
                                  var_StRose + var_TrumanState + var_TuftsHealthSciences + var_TuftsMainCampus + var_UChicago +
                                  var_UFlorida + var_UMDearborn + var_UMich + var_UTKnoxville +
                                  var_UniversityofSouthernCalifornia + var_Utah + var_VATech + var_VirginiaCommonwealth +
                                  var_WakeForest + var_WestVirginia + var_WesternCarolina + var_WesternMichigan +
                                  var_WesternWA + var_Xavier == 0, FALSE, NA))) %>%
    
  # Community colleges
    mutate(school2_cc = ifelse(var_OaklandCC + var_OklahomaCityCC + var_PaloAltoCollege + var_RedlandsCC == 1, TRUE, 
                               ifelse(var_OaklandCC + var_OklahomaCityCC + var_PaloAltoCollege + var_RedlandsCC == 0, FALSE, NA))) %>%
    
  # Art, design, culinary
    mutate(school2_ad = ifelse(var_AlbertaCAD + var_CIA + var_ColumbusCAD + var_MDInstituteCollegeArt + 
                                 var_MassArt + var_MinnesotaCAD + var_NHInstituteArt + var_Pratt + var_Ringling +
                                 var_SAIC + var_UNCSchoolArt + var_Watkins == 1, TRUE, 
                               ifelse(
                                 var_AlbertaCAD + var_CIA + var_ColumbusCAD + var_MDInstituteCollegeArt +
                                   var_MassArt + var_MinnesotaCAD + var_NHInstituteArt + var_Pratt + 
                                   var_Ringling + var_SAIC + var_UNCSchoolArt + var_Watkins == 0, FALSE, NA)))

# Binary variable for school type
table(data.v14$school2_4yr)
table(data.v14$school2_cc)
table(data.v14$school2_ad)

#  create one school type variables
data.v14 <- data.v14 %>%
  mutate(school2_type = case_when(school2_4yr == TRUE ~ 1,
                                  school2_cc == TRUE ~ 2,
                                  school2_ad == TRUE ~ 3)) %>%
  filter(!is.na(school2_type))
data.v14$school2_type <- as.factor(data.v14$school2_type)

table(data.v14$school2_type, useNA = "always") # check output

# Clean-up school2 var
data.v14 = subset(data.v14, select= -c(var_Boston, var_CalPoly, var_ColoradoCollege, var_ColoradoMountain,
                                       var_Delaware, var_DetroitMercy, var_Emerson, var_GeorgeMason,
                                       var_GeorgiaTech, var_Kalamazoo, var_Kansas, var_Merrimack, 
                                       var_MichiganState, var_MontclairState, var_NevadaReno,
                                       var_PennCollegeTech, var_RhodeIsland, var_Rollins, var_Sewanee, var_Smith,
                                       var_SouthernNazarene, var_SouthwesternOklahomaState, var_StJohns,
                                       var_StMarys, var_StRose, var_TrumanState, var_TuftsHealthSciences,
                                       var_TuftsMainCampus, var_UChicago, var_UFlorida, var_UMDearborn, var_UMich,
                                       var_UTKnoxville, var_UniversityofSouthernCalifornia, var_Utah, var_VATech,
                                       var_VirginiaCommonwealth, var_WakeForest, var_WestVirginia,
                                       var_WesternCarolina, var_WesternMichigan, var_WesternWA, var_Xavier,
                                       var_OaklandCC, var_OklahomaCityCC, var_PaloAltoCollege,
                                       var_RedlandsCC, var_AlbertaCAD, var_CIA, var_ColumbusCAD, 
                                       var_MDInstituteCollegeArt, var_MassArt, var_MinnesotaCAD, 
                                       var_NHInstituteArt, var_Pratt, var_Ringling, var_SAIC,
                                       var_UNCSchoolArt, var_Watkins))
```

## ENVIRONMENT ANALYSES
```{r}
data.v15 <- data.v14  # new dataframe for analysis
```

### Environmental correlation
```{r}
# Define environment vars for correlation matrix
data_env <- data.v15 %>% select(discrim, religios, residenc, activ_yn, divers, 
                                school2_type, satisfied_overall, talk, military, international)

 # Convert data to numeric in order to run correlations
data_env <- data_env %>% mutate_if(is.character, as.factor)
data_env <-  data_env %>% mutate_if(is.factor, as.numeric)
data_env <-  data_env %>% mutate_if(is.logical, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_env, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)), ]                   #sort by highest correlation
print(corr)
  
# turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
```

##EFA Environment
```{r EFA Environment}
data_env <- data.v15 %>% select(religios, residenc, activ_yn, school2_type, 
                                      satisfied_overall, talk)

data_env <- na.omit(data_env)
# create corr matrix, with dummy vars
data_env <- data_env %>% 
  mutate_if(is.numeric, as.factor) %>% 
  mutate(
    age = as.numeric(satisfied_overall)) %>%
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_env   # check output

# PCA map with eigenvalues
results_env <- dummy_cols(data_env, select_columns = c("religios", "residenc", "activ_yn",
                                                         "school2_type", "satisfied_overall", "talk"),
                          remove_first_dummy = TRUE)

env_pca <- PCA(results_env, graph=FALSE)  # apply PCA
names(env_pca)
summary(env_pca)

eigval_env <- get_eigenvalue(env_pca)
eigval_env

# Visualization
fviz_screeplot(env_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(env_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(env_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(env_pca, choice = "var", axes = 1,
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

### CLM environmental regression
```{r CLM environment}
clm_env <- data.v15 %>% select(si_type, religios, residenc, activ_yn, 
                                  school2_type, satisfied_overall, talk)

clm_env <- clm_env %>% 
  NA_preproc() %>%
  as.data.frame()
view(clm_env)

clm_env2 <- clm(si_type ~ religios + residenc + activ_yn +
                  school2_type + satisfied_overall + talk,
                data = clm_env)
summary(clm_env2)
coef(clm_env2)
```


# LIFE EVENTS
Candidate variables: ins_cover, gpa_sr, abuse_year, psyhx, isi_risk, binge_fr, fincur, drug_mar, drug_coc, drug_other, drug_met, drug_none, drug_opi, drug_her, drug_stim, drug_ect, assault_sex, assault_phys, assault_emo, isi_tot


```{r LE frame}
data.v16 <- data.v14
```

## Insurance coverage
```{r ins_cover}
# create one insurance var from related questions in 'ins_' module
data.v16 <- data.v16 %>% 
  mutate(ins_cover = case_when((ins_1 == 1 | ins_mh == 4 | ins_mh == 5 | ins_ade == 3) ~ 0,   # inadequate coverage
                               (ins_1 == 0 | ins_mh == 1 | ins_mh == 2 | ins_ade == 2) ~ 1)) %>% # adequate MH coverage
  filter(!is.na(ins_cover))   
data.v16$ins_cover <- as.logical(data.v16$ins_cover)
table(data.v16$ins_cover, useNA = "always")   # check output

#data.v16 = subset(data.v16, select= -c(ins_1, ins_mh, ins_ade, ins_inade)) # clean up var
```

## Illicit drug use
Creates illicit drug use var (limitation -- 30 days reflection)
```{r drugs_yn}
## add summary of each
table(data.v16$drug_none)
table(data.v16$drug_mar)  
table(data.v16$drug_coc)
table(data.v16$drug_her)  # too small
table(data.v16$drug_met)  # too small
table(data.v16$drug_stim)
table(data.v16$drug_ect)  # too small
table(data.v16$drug_opi)   # too small
table(data.v16$drug_other)  # too vague, small

# Sample is too small to analyze individual drugs, except for marijuana. Transform to illicit drug y/n
data.v16 <- data.v16 %>%
  mutate(drugs_yn = case_when(drug_none == 1 ~ 0,
                              drug_coc == 1 ~ 1,
                              drug_her == 1 ~ 1,
                              drug_met == 1 ~ 1,
                              drug_stim == 1 ~ 1,
                              drug_ect == 1 ~ 1,
                              drug_opi == 1 ~ 1,
                              drug_other == 1 ~ 1)) %>%
  mutate(drugs_yn = case_when(drugs_yn == 1 ~ 1,
                              drugs_yn == 0 ~ 0,
                              is.na(drugs_yn) ~ 0)) %>%
  mutate(drug_mar = case_when(drug_mar == 1 ~ 1, 
                              drug_mar == 0 ~ 0,
                              is.na(drug_mar) ~ 0))
data.v16$drug_mar <- as.logical(data.v16$drug_mar)
data.v16$drugs_yn <- as.logical(data.v16$drugs_yn)
  
table(data.v16$drugs_yn, useNA = "always")
table(data.v16$drug_mar, useNA = "always")
```

## Current financial stressors
```{r fincur}
data.v16 <- data.v16 %>%
  filter(!is.na(fincur)) %>%
  mutate(fincur = case_when(fincur == 3 ~ 0,  # sometimes
                            fincur == 4 ~ 0,  # rarely or never
                            fincur == 5 ~ 0,
                            fincur == 1 ~ 1,  # always or often stressful
                            fincur == 2 ~ 1))
data.v16$fincur <- as.logical(data.v16$fincur)
table(data.v16$fincur, useNA = "always")
```

## Abuse or assault, in the last year only
```{r assault}
table(data.v16$abuse_life)
table(data.v16$abuse_recent)

# create dichotomous variable called 'abuse_year' to reflect abuse that occurred in last 12 months
data.v17 <- data.v16 %>%
  mutate(abuse_year = case_when(abuse_recent == 1 ~ 1,
                                abuse_recent == 2 ~ 1,
                                abuse_recent == 3 ~ 1,
                                abuse_recent == 4 ~ 0,
                                abuse_recent == 5 ~ 0,
                                abuse_life == 0 ~ 0,
                                is.na(abuse_life) ~ 0))
table(data.v17$abuse_year)   # check output

# restructure sexual assault var
data.v17 <- data.v17 %>%
  mutate(assault = case_when(
                             assault_sex == 1 ~ 1,
                             assault_emo == 1 ~ 1,
                             assault_phys == 1 ~ 1, 
                             abuse_year == 1 ~ 1,
                             abuse_year == 0 ~ 0,
                             assault_sex == 2 ~ 0,
                             assault_emo == 0 ~ 0,
                             assault_phys == 1 ~ 0)) %>%
  mutate(assault = case_when(assault == 1 ~ 1,
                             assault == 0 ~ 0,
                             is.na(assault) ~ 0))
data.v17$assault <- as.logical(data.v17$assault)

# tabulate and check output
table(data.v17$assault, useNA = "always") 
                                         
data.v17 = subset(data.v17, select= -c(abuse_life, abuse_recent)) # clean up var
```

## Binging frequency
```{r binge_fr}
# set all binge_fr values equal to 0
data.v17$binge_fr_f[is.na(data.v17$binge_fr_f)] <- 0
data.v17$binge_fr_m[is.na(data.v17$binge_fr_m)] <- 0
data.v17$binge_fr_o[is.na(data.v17$binge_fr_o)] <- 0

# create one binge var
data.v18 <- data.v17 %>% 
  mutate(binge_fr = as.numeric(binge_fr_f) + as.numeric(binge_fr_m) + as.numeric(binge_fr_o))
table(data.v18$binge_fr)

data.v18 <- data.v18 %>%
  mutate(binge = case_when(binge_fr >= 4 ~ 1,
                           binge_fr < 4 ~ 0)) %>%
  filter(!is.na(binge))
data.v18$binge <-as.logical(data.v18$binge)
table(data.v18$binge, useNA = "always") # check output

data.v18 = subset(data.v18, select= -c(binge_fr_f, binge_fr_m, binge_fr_o)) # clean up var
```

## Recent Psychiatric Hospitalization, last 12 months
```{r psyhx}
# Create variable for any psychiatric hospitalization in the last 12 months
data.v18 <- data.v18 %>% 
  mutate(psyhx = case_when((prov_pes == 1 | prov_inp == 1 | prov_par == 1) ~ 1, 
                           (prov_pes == 0 & prov_inp == 0 & prov_par == 0) ~ 0,
                           ther_ever == 0 ~ 0)) %>%
  filter(!is.na(psyhx))
data.v18$psyhx <- as.logical(data.v18$psyhx)

table(data.v18$psyhx, useNA = "always")   # check output
data.v18 = subset(data.v18, select= -c(prov_pes, prov_inp, prov_par)) # clean up var
```

## Current GPA
```{r gpa_sr}
data.v18 <- data.v18 %>%
  mutate(gpa_sr = case_when(gpa_sr == 0 ~ 1,   # A+
                            gpa_sr == 1 ~ 1,   # A
                            gpa_sr == 2 ~ 1,   # A-
                            gpa_sr == 3 ~ 2,   # B+
                            gpa_sr == 4 ~ 2,   # B
                            gpa_sr == 5 ~ 2,   # B-
                            gpa_sr == 6 ~ 3,   # C+
                            gpa_sr == 7 ~ 3,   # C
                            gpa_sr == 8 ~ 3,   # C-
                            gpa_sr == 9 ~ 4,   # D+ and below
                            gpa_sr == 10 ~ 5,  # Don't know
                            is.na(gpa_sr) ~ 5))
data.v18$gpa_sr <- as.numeric(data.v18$gpa_sr)
table(data.v18$gpa_sr, useNA = "always")
```
 
```{r LE data analysis}
data.v19 <- data.v18
```
 
## LE correlation
```{r LE correlation}
# Define life event var candidates for correlation matrix
data_le <- data.v19 %>% select(ins_cover, gpa_sr, aca_impa, fincur, binge,
                                     assault, drugs_yn, drug_mar, psyhx)

# convert data to numeric in order to run correlations
data_le <- data_le %>% mutate_if(is.character, as.factor)
data_le <-  data_le %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_le, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop dups and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
```

## EFA LE
```{r le_pca}
data_le <- data.v19 %>% select(ins_cover, gpa_sr, aca_impa, fincur, 
                                     binge, drugs_yn, drug_mar, assault, psyhx)
data_le <- na.omit(data_le)



# create corr matrix, with dummy vars
data_le <- data_le %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_le   # check output

results_le <- dummy_cols(data_le, select_columns = c("ins_cover", "gpa_sr", "aca_impa", "fincur",
                                                        "binge","drugs_yn", "drug_mar", "assault", "psyhx"),
                         remove_first_dummy = TRUE)

le_pca <- PCA(results_le, graph=FALSE)  # apply PCA
le_pca$eig   # output matrix with eigenvalues

eigval_le <- get_eigenvalue(le_pca)
eigval_le

# visualization for PCA, FAMD
fviz_screeplot(le_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(le_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(le_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(le_pca, choice = "var", axes = 1,
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

## CLM Life event regression
```{r clm_LE}
clm_le <- data.v19 %>% select(si_type, ins_cover, gpa_sr, aca_impa, fincur, 
                                 assault, binge, drugs_yn, drug_mar, psyhx)
clm_le <- na.omit(clm_le)

# run regression
clm_le <- clm(si_type ~ ins_cover + gpa_sr + aca_impa + fincur + assault + binge + 
                drugs_yn + drug_mar + psyhx,
              data = clm_le)
summary(clm_le)
coef(clm_le)

ci_le <- confint(clm_le)
ci_le

model.le <- tidy(clm_le)  # Convert model to dataframe for easy manipulation
model.le
model.le  %>%
   mutate(or = exp(estimate),  # Odds ratio/gradient
          var.diag = diag(vcov(clm_le)),  # Variance of each coefficient
          or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# DEFEAT
Candidate variables: aca_impa, failed, men_ill_ash, deprawsc


```{r Defeat frame}
data.v20 <- data.v18
```

## PHQ risk/severity
```{r deprawsc}
table(data.v20$deprawsc)
data.v20$deprawsc <- as.numeric(data.v20$deprawsc)

# Categorize by clinical translation for later --
data.v20 <- data.v20 %>%
  mutate(dep_none = case_when(deprawsc < 5 ~ 1,
                              deprawsc > 4 ~ 0)) %>%
  mutate(dep_mild = case_when((deprawsc > 4 & deprawsc < 10) ~ 1,
                              (deprawsc < 5 | deprawsc > 9) ~ 0)) %>%
  mutate(dep_mod = case_when((deprawsc > 9 & deprawsc < 15) ~ 1,
                             (deprawsc > 14 | deprawsc < 10) ~ 0)) %>%
  mutate(dep_modsev = case_when((deprawsc > 14 & deprawsc < 20) ~ 1,
                               (deprawsc > 19 | deprawsc < 15) ~ 0)) %>%
  mutate(dep_sev = case_when(deprawsc > 19 ~ 1,
                             deprawsc < 20 ~ 0)) %>%
  mutate(dep_na = case_when(is.na(deprawsc) ~ 1,
                            !is.na(deprawsc) ~ 0))

# Check outputs
table(data.v20$dep_none)
table(data.v20$dep_mild)
table(data.v20$dep_mod)
table(data.v20$dep_modsev)
table(data.v20$dep_sev)
table(data.v20$dep_na)

data.v20 <- data.v20 %>%
  mutate(dep_type = case_when(dep_none == 1 ~ 1,
                              dep_mild == 1 ~ 2,
                              dep_mod == 1 ~ 3,
                              dep_modsev == 1 ~ 4,
                              dep_sev == 1 ~ 5)) %>%
  filter(!is.na(dep_type)) # filters out all NAs
data.v20$deprawsc <- as.numeric(data.v20$deprawsc)

table(data.v20$dep_type, useNA = "always")
```

## GAD7 total
```{r anx_score}
data.v20$anx_score <- as.numeric(data.v20$anx_score)

# Convert to severity
data.v20 <- data.v20 %>%
  filter(!is.na(anx_score)) %>% # filters out all NAs
  mutate(anx_type = case_when(anx_score > 14 ~ 4,
                              anx_score < 5 ~ 1,                       # anx none
                              anx_score < 10 ~ 2,                      # anx mild
                              anx_score < 15 ~ 3))                    # anx severe
data.v20$anx_score <- as.numeric(data.v20$anx_score)

table(data.v20$anx_type, useNA = "always") # check output
```

## Academic impact
```{r aca_impa}
data.v20 <- data.v20 %>%
  filter(!is.na(aca_impa)) %>%
  mutate(aca_impa = case_when(aca_impa == 1 ~ 1,
                              aca_impa == 2 ~ 2,
                              aca_impa == 3 ~ 3,
                              aca_impa == 4 ~ 4)) 
data.v20$aca_impa <- as.numeric(data.v20$aca_impa)

table(data.v20$aca_impa, useNA = "always")
```

## Defeat analysis
```{r Defeat analyses}
data.v21 <- data.v20
```

### Defeat correlation
```{r Defeat cor}
# Define defeat variable candidates for correlation matrix
data_def <- data.v21 %>% select(aca_impa, dep_type, anx_type)

 # convert data to numeric in order to run correlations
data_def <- data_def %>% mutate_if(is.character, as.factor)
data_def <-  data_def %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_def, use = "pairwise.complete.obs")                               
corr[lower.tri(corr, diag=TRUE)] <- NA        # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                # drop perfect correlations
corr <- as.data.frame(as.table(corr))         # turn into a 3-column table
corr <- na.omit(corr)                         # remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]         # sort by highest correlation
print(corr)
```

### EFA defeat
```{r EFA_def}
data_def <- data.v21 %>% select(aca_impa, dep_type, anx_type)
data_def <- na.omit(data_def)

# create corr matrix, with dummy vars 
data_def <- data_def %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_def   # check output

results_def <- dummy_cols(data_def, select_columns = c("aca_impa", "dep_type", "anx_type"),
                         remove_first_dummy = TRUE)

def_pca <- PCA(results_def, graph=FALSE)  # apply PCA
def_pca$eig   # output matrix with eigenvalues

eigval_def <- get_eigenvalue(def_pca)
eigval_def

# visualization for PCA, FAMD
fviz_screeplot(def_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(def_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(def_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(def_pca, choice = "var", axes = 1, 
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

### CLM defeat
```{r CLM defeat}
results2_def <- data.v21 %>% select(si_type, aca_impa, dep_type, anx_type)
clm_def <- na.omit(results2_def)

# run defeat regression
clm_def <- clm(si_type ~ aca_impa + dep_type + anx_type, data = clm_def)

summary(clm_def)
coef(clm_def)

ci_def <- confint(clm_def)
ci_def

model.def <- tidy(clm_def)  # Convert model to dataframe for easy manipulation
model.def
model.def  %>%
   mutate(or = exp(estimate),  # Odds ratio/gradient
          var.diag = diag(vcov(clm_def)),  # Variance of each coefficient
          or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# THREATS TO SELF-MODERATION
Candidate variables: persist, anx_score, dep_secret, aaq_yn, achieve4


```{r TSM frame}
data.v22 <- data.v20
```

## Depression secret
```{r dep_secret}
## Secretive MH
data.v22 <- data.v22 %>%
  mutate(dep_secret = case_when(dep_secret == 1 ~ 1,
                                dep_secret == 2 ~ 2,
                                dep_secret == 3 ~ 3,
                                dep_secret == 4 ~ 4,
                                dep_secret == 5 ~ 5,
                                dep_secret == 6 ~ 6))
  # filter(!is.na(dep_secret))
data.v22$dep_secret <- as.numeric(data.v22$dep_secret)
table(data.v22$dep_secret, useNA = "always")  # not sure what to do about this one.... NA = 7000+
```

## AAQ total
```{r aaq_dum}
# create AAQ_tot variable to sum up imputed AAQ questions
data.v22 <- data.v22 %>%
  mutate(aaq_tot = as.numeric(AAQ_1) + as.numeric(AAQ_2) + as.numeric(AAQ_3) + as.numeric(AAQ_4) + 
           as.numeric(AAQ_5) + as.numeric(AAQ_6) + as.numeric(AAQ_7))

table(data.v22$aaq_tot)

#  for loop for predictive cutoff point needed
 rs_df <- data.frame("cutoff" = c(), "rs" = c())
 for (i in 8:48) {
   aaq_bin <- ifelse(data.v22$aaq_tot >= i, TRUE, FALSE)
   model = lm(data.v22$sui_idea ~ aaq_bin)
   rs_df <- rbind(rs_df, c(i, summary(model)$r.squared))

 }
 names(rs_df) <- c("cutoff", "rs")
 rs_df_arranged <- rs_df %>%
   arrange(desc(rs))
 rs_df_arranged[1,1]

# deem 'rs_df_arranged[1,1]' as SI cutoff
 data.v22 <- data.v22 %>%
   mutate(aaq_dum = ifelse(aaq_tot > rs_df_arranged[1,1], 1, 0)) %>%
   mutate(aaq_dum = case_when(aaq_dum == 1 ~ 1,
                              aaq_dum == 0 ~ 0)) %>%
   filter(!is.na(aaq_dum))
data.v22$aaq_dum <- as.logical(data.v22$aaq_dum)

table(data.v22$aaq_dum, useNA = "always")
```

## Persist
```{r persist}
data.v22 <- data.v22 %>%
  mutate(persist = case_when(persist == 1 ~ 1,
                             persist == 2 ~ 2, 
                             persist == 3 ~ 3, 
                             persist == 4 ~ 4, 
                             persist == 5 ~ 5, 
                             persist == 6 ~ 6))
data.v22$persist <- as.numeric(data.v22$persist)

table(data.v22$persist, useNA = "always") # check output
```

## TSM Analysis
```{r TSM var analyses}
data.v23 <- data.v22
```

### Threats to self-moderation correlation
```{r CORR TSM}
# Define TSM variable candidates for correlation matrix
data_tsm <- data.v23 %>% select(persist, aaq_dum)

# convert data to numeric in order to run correlations 
data_tsm <- data_tsm %>% mutate_if(is.character, as.factor)
data_tsm <-  data_tsm %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_tsm, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
# Re-run corr with non-overlapping vars
data_tsm <- data.v23 %>% select(persist, aaq_dum)

# convert data to numeric in order to run correlations 
data_tsm <- data_tsm %>% mutate_if(is.character, as.factor)
data_tsm <-  data_tsm %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_tsm, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
```

### EFA TSM
```{r EFA TSM}
data_tsm <- data.v23 %>% select(persist, aaq_dum)
data_tsm <- na.omit(data_tsm)

# create corr matrix, with dummy vars 
data_tsm <- data_tsm %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_tsm   # check output

results_tsm <- dummy_cols(data_tsm, select_columns = c("persist", "aaq_dum"),
                         remove_first_dummy = TRUE)

tsm_pca <- PCA(results_tsm, graph=FALSE)  # apply PCA
eigval_tsm <- get_eigenvalue(tsm_pca)
eigval_tsm

# Visualization of PCA, FAMD
fviz_screeplot(tsm_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(tsm_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(tsm_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(tsm_pca, choice = "var", axes = 1,
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

### CLM TSM
```{r CLM TSM}
clm_tsm <- data.v23 %>% select(si_type, persist, aaq_dum)
clm_tsm <- na.omit(clm_tsm)

clm_tsm <- clm(si_type ~ persist + aaq_dum, data = clm_tsm) 

summary(clm_tsm)
coef(clm_tsm)

ci_tsm <- confint(clm_tsm)
ci_tsm

model.tsm <- tidy(clm_tsm)  # Convert model to dataframe for easy manipulation
model.tsm

model.tsm  %>%
    mutate(or = exp(estimate),  # Odds ratio/gradient
           var.diag = diag(vcov(clm_tsm)),  # Variance of each coefficient
           or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# ENTRAPMENT
Candidate variables: dep_impa, gad7_impa, perneed_cur


```{r Entrap frame}
data.v24 <- data.v22
```

## Depression impact
```{r dep_impa}
data.v24 <- data.v24 %>%
  mutate(dep_impa = case_when(dep_impa == 1 ~ 1,
                              dep_impa == 2 ~ 2,
                              dep_impa == 3 ~ 3,
                              dep_impa == 4 ~ 4)) %>%
  filter(!is.na(dep_impa))
data.v24$dep_impa <- as.numeric(data.v24$dep_impa)

table(data.v24$dep_impa, useNA = "always")

```

## Anxiety impact
```{r gad7_impa}
data.v24 <- data.v24 %>%
  mutate(gad7_impa = case_when(gad7_impa == 1 ~ 1,
                               gad7_impa == 2 ~ 2,
                               gad7_impa == 3 ~ 3,
                               gad7_impa == 4 ~ 4)) %>%
  filter(!is.na(gad7_impa))
data.v24$gad7_impa <- as.numeric(data.v24$gad7_impa)

table(data.v24$gad7_impa, useNA = "always")
```

## Current needs
```{r percneed_cur}
data.v24 <- data.v24 %>%
  mutate(percneed_cur = case_when(percneed_cur == 1 ~ 1,
                                 percneed_cur == 2 ~ 2,
                                 percneed_cur == 3 ~ 3,
                                 percneed_cur == 4 ~ 4,
                                 percneed_cur == 5 ~ 5,
                                 percneed_cur == 6 ~ 6))
data.v24$percneed_cur <- as.numeric(data.v24$percneed_cur)

table(data.v24$percneed_cur, useNA = "always")
```

## ENTRAPMENT ANALYSES
```{r Entrapment analyses}
data.v25 <- data.v24
```

### Entrapment correlation
```{r Entrapment corr}
# Define entrapment variable candidates for correlation matrix
data_entrap <- data.v25 %>% select(dep_impa, gad7_impa)

 # convert data to numeric in order to run correlations
data_entrap <- data_entrap %>% mutate_if(is.character, as.factor)
data_entrap <-  data_entrap %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_entrap, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
```

### EFA Entrapment
```{r CFA entrap}
data_entrap <- data.v25 %>% select(dep_impa, gad7_impa)
data_entrap <- na.omit(data_entrap)

# create corr matrix, with dummy vars 
results_entrap <- data_entrap %>% 
  mutate_if(is.numeric, as.factor) %>% 
  mutate_if(is.character, as.factor) %>% 
  data.matrix()
results_entrap   # check output

entrap_pca <- PCA(results_entrap, graph=FALSE)  # apply PCA
entrap_pca$eig   # output matrix with eigenvalues

eigval_entrap <- get_eigenvalue(entrap_pca)
eigval_entrap

# Visualization of PCA, FAMD
fviz_screeplot(entrap_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(entrap_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(entrap_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(entrap_pca, choice = "var", axes = 1, 
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

###  CLM Entrapment regression
```{r clm_entrap}
clm_entrap <- data.v25 %>% select(si_type, dep_impa, gad7_impa)
clm_entrap <- na.omit(clm_entrap)

# run entrapment regression
clm_entrap <- clm(si_type ~ dep_impa + gad7_impa, data = clm_entrap)

summary(clm_entrap)
coef(clm_entrap)

ci_entrap <- confint(clm_entrap)
ci_entrap

model.entrap <- tidy(clm_entrap)  # Convert model to dataframe for easy manipulation
model.entrap

model.entrap  %>%
   mutate(or = exp(estimate),  # Odds ratio/gradient
         var.diag = diag(vcov(clm_entrap)),  # Variance of each coefficient
         or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# MOTIVATIONAL MODERATORS
Candidate variables: flourish, belong2, belong8, belong9, achieve1, achieve14, 
achieve18, stig_pcv_2, stig_pcv_3, friend_devel, inf, brs_tot


```{r MM frame}
data.v26 <- data.v24
```

## Belonging
```{r belong}
data.v26 <- data.v26 %>%
  mutate(belong1 = case_when(belong1 == 1 ~ 1,
                             belong1 == 2 ~ 2,
                             belong1 == 3 ~ 3,
                             belong1 == 4 ~ 4,
                             belong1 == 5 ~ 5,
                             belong1 == 6 ~ 6)) %>%
  filter(!is.na(belong1))
table(data.v26$belong1, useNA = "always")

data.v26$belong1 <- as.numeric(data.v26$belong1)
summary(data.v26$belong1)

table(data.v26$belong2)  # elective question, but common
table(data.v26$belong8)  # nested x 2
table(data.v26$belong9)  # not as good of a proxy
```

## Flourishing Scale
```{r flourish}
data.v26$flourish <- as.numeric(data.v26$flourish)
table(data.v26$flourish)

#  for loop for predictive cutoff point needed
rs_df <- data.frame("cutoff" = c(), "rs" = c())
for (i in 9:55) {
  flourish_bin <- ifelse(data.v26$flourish >= i, TRUE, FALSE)
  model = lm(data.v26$sui_idea ~ flourish_bin)
  rs_df <- rbind(rs_df, c(i, summary(model)$r.squared))

}
names(rs_df) <- c("cutoff", "rs")
rs_df_arranged <- rs_df %>% 
  arrange(desc(rs)) 
rs_df_arranged[1,1] # chi-squared high

# create levels for scoring
data.v26 <- data.v26 %>%
  mutate(flourish_dum = case_when(flourish > 50 ~ 5,
                                  flourish > 40 ~ 4,
                                  flourish > 30 ~ 3,
                                  flourish > 20 ~ 2,
                                  flourish < 14 ~ 1)) %>%
  filter(!is.na(flourish_dum))
data.v26$flourish_dum <- as.numeric(data.v26$flourish_dum)

table(data.v26$flourish_dum, useNA = "always")
```

## Informal Mental Health Support
```{r inf}
data.v26$inf_1[is.na(data.v26$inf_1)] <- 0
data.v26$inf_2[is.na(data.v26$inf_2)] <- 0
data.v26$inf_3[is.na(data.v26$inf_3)] <- 0
data.v26$inf_4[is.na(data.v26$inf_4)] <- 0
data.v26$inf_5[is.na(data.v26$inf_5)] <- 0
data.v26$inf_6[is.na(data.v26$inf_6)] <- 0
data.v26$inf_7[is.na(data.v26$inf_7)] <- 0

# change inf_ var to numeric
data.v26 <- data.v26 %>% 
  # merge all var into inf_type
  mutate(inf = case_when(inf_8 == 1 ~ 1,
                         inf_1 == 1 ~ 0,
                         inf_2 == 1 ~ 0,
                         inf_3 == 1 ~ 0,
                         inf_4 == 1 ~ 0,
                         inf_5 == 1 ~ 0,
                         inf_6 == 1 ~ 0,
                         inf_7 == 1 ~ 0)) %>%
  filter(!is.na(inf))
data.v26$inf <- as.logical(data.v26$inf)
table(data.v26$inf, useNA ="always")

# clean up vars
#data.v26 = subset(data.v26, select= -c(inf_1:inf_help))
```

## Perceived MH as failure, stigma
```{r stig_pcv_2}
# original likert
table(data.v26$stig_pcv_2)
table(data.v26$stig_pcv_3)
table(data.v26$stig_per_2)
table(data.v26$stig_per_3)

# transform into y/n 
data.v26 <- data.v26 %>%
  mutate(stig_pcv2 = case_when((stig_pcv_2 == 1 | stig_pcv_2 == 2 | stig_pcv_2 == 3) ~ 1,
                                (stig_pcv_2 == 4 | stig_pcv_2 == 5 | stig_pcv_2 == 6) ~ 0)) %>%
  mutate(stig_pcv3 = case_when((stig_pcv_3 == 1 | stig_pcv_3 == 2 | stig_pcv_3 == 3) ~ 1,
                                (stig_pcv_3 == 4 | stig_pcv_3 == 5 | stig_pcv_3 == 6) ~ 0)) %>%  
  mutate(stig_per2 = case_when((stig_per_2 == 1 | stig_per_2 == 2 | stig_per_2 == 3) ~ 1,
                                (stig_per_2 == 4 | stig_per_2 == 5 | stig_per_2 == 6) ~ 0)) %>%
  mutate(stig_per3 = case_when((stig_per_3 == 1 | stig_per_3 == 2 | stig_per_3 == 3) ~ 1,
                                (stig_per_3 == 4 | stig_per_3 == 5 | stig_per_3 == 6) ~ 0)) 

data.v26$stig_pcv2 <- as.logical(data.v26$stig_pcv2)
data.v26$stig_pcv3 <- as.logical(data.v26$stig_pcv3)
data.v26$stig_per2 <- as.logical(data.v26$stig_per2)
data.v26$stig_per3 <- as.logical(data.v26$stig_per3)

table(data.v26$stig_pcv2)  # Most people feel that MH tx is a sign of personal failure.
table(data.v26$stig_pcv3)  # Most people think less of a person who has received MH tx.
table(data.v26$stig_per2)  # I feel that receiving MH tx is a sign of personal failure.
table(data.v26$stig_per3)  # I would think less of a person who has received MH tx.

# combine scales
data.v26 <- data.v26 %>%
  mutate(mh_stigma = case_when(stig_pcv2 == 1 ~ 1,
                               stig_pcv3 == 1 ~ 1,
                               stig_per2 == 1 ~ 1,
                               stig_per3 == 1 ~ 1,
                               stig_pcv2 == 0 ~ 0,
                               stig_pcv3 == 0 ~ 0,
                               stig_per2 == 0 ~ 0,
                               stig_per3 == 0 ~ 0))
data.v26$mh_stigma <- as.logical(data.v26$mh_stigma)

table(data.v26$mh_stigma, useNA = "always")
```

## Motivational moderator analyses
```{r MM var analyses}
data.v27 <- data.v26
```

### Motivational Moderator correlation
```{r MM_corr}
# Define MM variable candidates for correlation matrix
data_mm <- data.v27 %>% select(inf, flourish_dum, belong1, mh_stigma)

 # convert data to numeric in order to run correlations
data_mm <- data_mm %>% mutate_if(is.character, as.factor)
data_mm <-  data_mm %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_mm, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   # turn into a 3-column table
corr <- na.omit(corr)                                   # remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   # sort by highest correlation
print(corr)

```

### EFA MM
```{r EFA MM}
data_mm <- data.v27 %>% select(inf, flourish_dum, belong1, mh_stigma)
data_mm <- na.omit(data_mm)

# create corr matrix, with dummy vars 
data_mm <- data_mm %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_mm   # check output
results_mm <- dummy_cols(data_mm, select_columns = c("inf", "flourish_dum", "belong1"),
                         remove_first_dummy = TRUE)
mm_pca <- PCA(results_mm, graph=FALSE)  # apply PCA
eigval_mm <- get_eigenvalue(mm_pca)
eigval_mm

# Visualization of PCA, FAMD
fviz_screeplot(mm_pca, addlabels = TRUE, ylim = c(0, 50))
fviz_pca_var(mm_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_famd_var(mm_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fviz_cos2(mm_pca, choice = "var", axes = 1, 
          fill = "steelblue", color = "steelblue",
          sort.val = c("desc", "asc", "none"),
          top = 40,
          xtickslab.rt = 45,
          ggtheme = theme_minimal())
```

### CLM Motivational moderators regression
```{r CLM mm}
clm_mm <- data.v27 %>% select(si_type, inf, flourish_dum, belong1)
clm_mm <- na.omit(clm_mm)

# run MM regression
clm_mm <- clm(si_type ~ inf + flourish_dum + belong1, data = clm_mm)

summary(clm_mm)
coef_mm <- coef(summary(clm_mm))

ci_mm <- confint(clm_mm)
ci_mm

model.mm <- tidy(clm_mm)  # Convert model to dataframe for easy manipulation
model.mm

model.mm  %>%
  mutate(or = exp(estimate),  # Odds ratio/gradient
        var.diag = diag(vcov(clm_mm)),  # Variance of each coefficient
        or.se = sqrt(or^2 * var.diag))  # Odds-ratio adjusted
```


# MISSINGNESS


## Absolute missingness, by school
```{r Structural missingness, by school}
data.v28 <- data.v26
# school structural analysis
var_names <- c(
  # diathesis
  "age", "sexual", "gender_tgd", "sib_freq", "race", "ed_any", "dx_mh", "meds_count",
  # environment
  "satisfied_overall", "talk", "religios", "school2_type", "activ_yn",
  # life events
  "ins_cover", "gpa_sr", "binge", "psyhx", "fincur", "drugs_yn", "assault", "drug_mar",
  # defeat
  "aca_impa", "dep_type", "anx_type",
  # threats to self-moderation
  "persist", "aaq_dum",
  # entrapment
  "dep_impa", "gad7_impa",
  # motivational moderators
  "inf", "flourish_dum", "belong1")  
var_names <- sort(var_names)

# build table for absolute missingness based on school2
school2 <- count(data.v28, school2)
school2_ncol <- ncol(school2)

for (j in 1:length(var_names)) {
  school2[, school2_ncol + j] <- 999
}

# create column headers and vectors for school, total students, then each variable count
names <- c("school2_values", "total", var_names)
names(school2) <- names

# input values for each school missingness across all variables
for (j in 1:(length(var_names))) {
for (i in 1:nrow(school2)) {   # i corresponds to rows, j to columns
  school2[i, 2 + j] <- sum(!is.na(data.v28[data.v28$school2 == school2[[i,1]], var_names[j]]))
}
}
# add percentage in each column
school2_m <- as.matrix(school2[-1]) 
props <- school2_m[,2:ncol(school2_m)] / school2_m[,1]
school2 <- data.frame(school2[, 1:2], props)

# view missingness matrix
view(school2)

# save matrix
write_xlsx(school2, path = "sch2_missing.xlsx", col_names = TRUE, format_headers = TRUE)
```

## Missingness of each var w/in each var
```{r Missingness between vars}
missing_dfs <- list()
for (k in 1:length(var_names)) {

#create skeleton of missingness matrix
var <- count(data.v28, data.v28[var_names[k]])
var_ncol <- ncol(var)

for (j in 1:length(var_names)) {
  var[, var_ncol + j] <- 999
}
# create column headers and vectors for school, total students, then each variable count
names <- c(paste0(var_names[k], "_values"), "total", var_names)
names(var) <- names

# input values for each school missingness across all variables
for (j in 1:(length(var_names))) {
for (i in 1:nrow(var)) {   # i corresponds to rows, j to columns
  var[i, 2 + j] <- sum(!is.na(data.v28[data.v28[[var_names[k]]] == var[[i,1]], var_names[j]]))
}
}
# add percentage in each column
var_m <- as.matrix(var[-1]) 
props <- var_m[,2:ncol(var_m)] / var_m[,1]
var <- data.frame(var[, 1:2], props)
replace_na("NA")
missing_dfs[[k]] <- var
}

names(missing_dfs) <- var_names
save(missing_dfs, file = "missing_dfs.RData")
load("missing_dfs.RData")
write_xlsx(missing_dfs, path = "missing.xlsx", col_names = TRUE, format_headers = TRUE)
```


# DATA PREP


## Select all vars for SEM
```{r train_df_sem2}
# Prep SEM data vars
data_sem <- data.v28 %>% select(
  # outcome var
  si_type,
  # diathesis
  age, sexual, gender_tgd, sib_freq, race, ed_any, dx_mh, meds_count,
  # environment
  satisfied_overall, talk, religios, school2_type, activ_yn,
  # life events
  ins_cover, gpa_sr, binge, psyhx, fincur, drugs_yn, assault, drug_mar,
  # defeat
  aca_impa, dep_type, anx_type,
  # threats to self-moderation
  persist, aaq_dum,
  # entrapment
  dep_impa, gad7_impa,
  # motivational moderators
  inf, flourish_dum, belong1)  #mh_stigma, 

# create sem matrix 
data_sem2 <- data_sem %>% 
  # mutate_if(is.numeric, as.factor) %>% 
  data.matrix()
data_sem2 <- na.omit(data_sem2)
head(data_sem2)   # check output

# # xlsx output
data_sem_out <- as.data.frame(data_sem2)
save(data_sem_out , file = "data_nondummy.RData")
load("data_nondummy.RData")
write_xlsx(data_sem_out, path = "data_nondummy.xls", col_names = TRUE, format_headers = TRUE)
```

## Create dummy vars
```{r dummy_vars}
results_tr_sem <- dummy_cols(data_sem2, select_columns = c(
    # outcome var
    "si_type",
    # diathesis
    "age", "sexual", "gender_tgd", "sib_freq", "race", "ed_any", "dx_mh", "meds_count",
    # environment
    "satisfied_overall", "talk", "religios", "school2_type", "activ_yn",
    # life events
    "ins_cover", "gpa_sr", "binge", "psyhx", "fincur", "drugs_yn", "assault", "drug_mar",
    # defeat
    "aca_impa", "dep_type", "anx_type",
    # threats to self-moderation
    "persist", "aaq_dum",
    # entrapment
    "dep_impa", "gad7_impa",
    # motivational moderators
    "inf", "flourish_dum", "belong1"), 
                         remove_first_dummy = TRUE, remove_selected_columns = TRUE)

results_tr_sem <- na.omit(results_tr_sem)
results_tr_df <- as.data.frame(results_tr_sem)
head(results_tr_df)

#save full var set for SEM
save(results_tr_df, file = "data_sem.RData")
load("data_sem.RData")
write_xlsx(results_tr_df, path = "data_sem.xls", col_names = TRUE, format_headers = TRUE)
```


# SUMMARY STATISTICS


## Descriptive stats
```{r Var_sum}
# training data
var_desc18 <- describe(data_sem2)
var_desc18
```

## Model w/ levels correlations
```{r model_lev_corr}
results_corr <- NA_preproc(data_sem2)
sig <- 0.4

# run a correlation and drop the insignificant ones 
corr <- cor(results_corr, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   # turn into a 3-column table
corr <- na.omit(corr)                                   # remove the NA values from above  
corr <- subset(corr, abs(Freq) > sig)                   # select significant values  
corr <- corr[order(-abs(corr$Freq)),]                   # sort by highest correlation
print(corr)
  
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
```


# SEM ANALYSIS


## Prep for SEM dummy cols
```{r results_tr_com2}
# Prep SEM data vars
data_sem5 <- data.v28 %>% select(
  # outcome var
  si_type,
  # diathesis
  age, sexual, gender_tgd, sib_freq, race, ed_any, dx_mh, meds_count,
  # environment
  satisfied_overall, talk, religios, school2_type, activ_yn,
  # life events
  ins_cover, gpa_sr, binge, psyhx, fincur, drugs_yn, assault, drug_mar,
  # defeat
  aca_impa, anx_type, dep_type,
  # threats to self-moderation
  persist, aaq_dum,
  # entrapment
  dep_impa, gad7_impa,
  # motivational moderators
  inf, flourish_dum, belong1)  #mh_stigma, 

# create sem matrix 
data_sem5 <- data_sem5 %>% 
  # mutate_if(is.numeric, as.factor) %>% 
  data.matrix()
data_sem5 <- na.omit(data_sem5)
head(data_sem5)   # check output

results_tr_com <- dummy_cols(data_sem5, select_columns = c("si_type", "sexual", "inf", "anx_type", "dep_type",
                                                              "race", "talk", "school2_type"),
                         remove_first_dummy = TRUE, remove_selected_columns = TRUE)
results_tr_com2 <- cbind(data_sem5, results_tr_com)
```

# SEM model
```{r si_model_tr2}
## Dummy col SEM w/ continuous
SI_model_tr2 <- '
                # measurement model
                  si =~  si_type_2 + si_type_3 + si_type_4
    
                  prem  =~ age + sexual_2 + sexual_3 + gender_tgd + sib_freq + 
                           race_2 + race_3 + race_4 + race_5 + ed_any + dx_mh +
                           meds_count + satisfied_overall + talk_2 + talk_3 + talk_4 + religios + 
                           school2_type_2 + school2_type_3 + activ_yn +
                           ins_cover + gpa_sr + binge + psyhx + fincur + drugs_yn + assault + drug_mar         

                   def =~ aca_impa + dep_type + anx_type
                              
                   tsm =~ persist + aaq_dum

                   entrap =~ dep_impa + gad7_impa

                   mm =~  inf + flourish_dum + belong1

                # structural model - regressions
                   si ~ prem + def + tsm + def*tsm + entrap + mm + entrap*mm 

                # residual covars
                  dx_mh ~~ ed_any
                  dx_mh ~~ meds_count
                  assault ~~ dx_mh
                  drugs_yn ~~ drug_mar '

fit_train3 <- sem(SI_model_tr2, data=results_tr_com2)
summary_train <- summary(fit_train3, fit.measures=TRUE, standardized = TRUE)
```


## Model prediction
```{r SEM test model}
# model diagnostics
fitmeasures(fit_train3)
fitted(fit_train3)
inspect(fit_train3, what="cor.all")

# compare to observed
lavCor(fit_train3)

# residuals in coef unit
resid(fit_train3, "cor")

# modification indices
modificationindices(fit_train3, minimum.value = 20)

fit_train4 <- update(fit_train3)
summary(fit_train4, fit.measures=TRUE)
```

## Parameter estimates
```{r parameter_est}
#train
# semPaths(fit_train3, "model", "est", layout = "circle", curvePivot = FALSE, edge.label.cex = 0.5)
semPaths(fit_train3, what = "col", whatLabels = "std", style = "mx", rotation = 1, layout = "spring", 
    nCharNodes = 8, shapeMan = "rectangle", sizeMan = 8, sizeMan2 = 8)

parameterEstimates(fit_train3)
fitted(fit_train3)
resid(fit_train3)
fitMeasures(fit_train3, c("cfi","chisq", "df", "pvalue","rmsea","srmr"), output = "matrix")

print(fitMeasures(fit_train3, c("chisq", "df", "pvalue", "cfi", "rmsea"),
                  output = "text"), add.h0 = TRUE)

lavInspect(fit_train3, what = "start")
lavInspect(fit_train3, what = "list")
```


```{r mediation}

# TSM as a mediator?
fit.totaleffect=lm(entrap~def, results_tr_com2)
summary(fit.totaleffect)
fit.mediator <- lm(tsm~def, results_tr_com2)
summary(fit.mediator)
fit.dv <- lm(entrap~def+tsm, results_tr_com2)
summary(fit.dv)

results = mediate(fit.mediator, fit.dv, sims = 1000, boot=F, treat = "def", mediator = "tsm")
summary(results)
sens.bout <- medsens(results, rho.by = 0.05, sims = 1000)
plot(sens.bout, sens.par = "rho")
plot(sens.bout, sens.par = "R2", r.type = 2, sign.prod = 1)

# TSM as a moderator
xz<-def*tsm
summary(lm(entrap~def+tsm+xz))


# # ENTRAPMENT AS A MEDIATOR
# fit.totaleffect2 <- lm(si~def, results_tr_com2)
# summary(fit.totaleffect2)
# 
# fit.mediator2 <- lm(entrap~def, results_tr_com2)
# summary(fit.mediator2)
# 
# fit.dv2 <- lm(si~def+entrap, results_tr_com2)
# summary(fit.dv2)
# 
# results2 <- mediate(fit.mediator2, fit.dv2, sims = 1000, boot= FALSE, treat = "def", mediator = "entrap")
# summary(results2)
# sens.bout <- medsens(results2, rho.by = 0.05, sims = 1000)
# plot(sens.bout, sens.par = "rho")
# plot(sens.bout, sens.par = "R2", r.type = 2, sign.prod = 1)


# # MM as a mediator or moderator?
# fit.totaleffect3 <- lm(si~entrap, results_tr_com2)
# summary(fit.totaleffect3)
# 
# fit.mediator3 <- lm(mm~entrap, results_tr_com2)
# summary(fit.mediator3)
# 
# fit.dv3 <- lm(si~entrap+mm, results_tr_com2)
# summary(fit.dv3)
# 
# results3 = mediate(fit.mediator3, fit.dv2, sims = 1000, boot=F, treat = "entrap", mediator = "mm", robustSE=T)
# summary(results3)
# sens.bout <- medsens(results2, rho.by = 0.05, sims = 1000)
# plot(sens.bout, sens.par = "rho")
# plot(sens.bout, sens.par = "R2", r.type = 2, sign.prod = 1)

```

## PLS-PM
```{r pls path}

data.pls <- read_excel("~/Desktop/sem_combo.xlsx", range=cell_cols(1:46), col_types = "numeric")
save("data.pls", file="~/Desktop/sem_combo.RData")
load("~/Desktop/sem_combo.RData")

# INNER MODEL
# rows of path matrix
prem =   c(0, 1, 0, 0, 0, 0)
def =    c(0, 0, 1, 1, 0, 0)
tsm =    c(0, 0, 0, 1, 0, 0)
entrap = c(0, 0, 0, 0, 1, 1)
mm =     c(0, 0, 0, 0, 0, 1)
si =     c(0, 0, 0, 0, 0, 0)

# matrix (by row binding)
imv_path = rbind (prem, def, tsm, entrap, mm, si)

imv_path = t(imv_path)
View(imv_path)
view(data.pls)

# plot the inner matrix
innerplot(imv_path, box.size = 0.1)

# outer model
imv_blocks = list(4:31, 32:39, 40:41, 42:43, 44:46, 1:3)
 
# modes
imv_modes = c("B", "A", "B", "A", "B", "B")

# apply plspm
imv_pls1 = plspm(data.pls, imv_path, imv_blocks, modes = imv_modes)

# print diss_pls1
imv_pls1

## Partial Least Squares Path Modeling (PLS-PM)
## ---------------------------------------------
##    NAME                   DESCRIPTION
## 1  $outer_model           outer model
## 2  $inner_model           inner model
## 3  $path_coefs            path coefficients matrix
## 4  $scores                latent variable scores
## 5  $crossloadings         cross-loadings
## 6  $inner_summary         summary inner model
## 7  $effects               total effects
## 8  $unidim                uni-dimensionality
## 9  $gof                   goodness-of-fit
## 10 $boot                  bootstrap results
## 11 $data                  data matrix
## ---------------------------------------------

# check unidimensionality
imv_pls1$unidim

# check coefs
imv_pls1$path_coefs
imv_pls1$inner_model

# plotting the loadings
plot(imv_pls1, what = "loadings")

# check outer model
imv_pls1$outer_model

# summarized results
summary(imv_pls1)
plot(imv_pls1)
plot(imv_pls1, what = "loadings", arr.width = 0.1)
head(imv_pls1$scores, n = 5)
tail(imv_pls1$scores, n = 5)
```


