---
title: "Sisler Initial Analysis"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# LOAD PACKAGES
```{r Call packages}
library(assertive)
library(bestglm)
library(broom)
library(car)
library(carat)
library(caret)
library(caTools)
library(clipr)
library(corrplot)
library(cowplot)
library(fastDummies)
library(effsize)
library(effectsize)
library(emulator)
library(esc)
library(expss)
library(FactoMineR)
library(finalfit)
library(FactoMineR)
library(glmnet)
library(glmnetcr)
library(haven)
library(Hmisc)
library(ISLR)
library(janitor)
library(lavaan)
library(lattice)
library(leaps)
library(lsr)
library(magrittr)
library(MASS)
library(missForest)
library(missMDA)
library(MissMech)
library(mvord)
library(ordinal)
library(ordinalForest)
library(ordinalgmifs)
library(ordinalNet)
library(PerformanceAnalytics)
library(psych)
library(purrr)
library(randomForest)
library(readr)
library(repr)
library(rcompanion)
library(readxl)
library(reshape2)
library(rococo)
library(rpart)
library(rpart.plot)
library(rstatix)
library(stats)
library(stringr)
library(sure)
library(tidyverse)
library(validate)
library(VIM)
library(writexl)

```

# PREPARING FILES

Notes: LEADING VARIABLE FOR SEM SELECTION

Premotivational variables
  - Diathesis: age, sexual, gender, finpast, sib_freq, race, body_sr, ed_scoff, audit_tot, dx_dep, dx_bi, dx_psy, dx_ea, dx_tr,  wcs_tot
    dx_sa, dx_adhd, dx_pers, dx_none, meds_sti, meds_dep, meds_psy, meds_anx, meds_mood, meds_sle, meds_oth, meds_none, meds_count, disab_1, alc_tx
  - Environment: env_mh, fam_support_aca, ther_vis, divers, satisfied_overall, residenc, talk, activ_yn, religios, discrim
  - Life Events: ins_cover, gpa_sr, abuse_year, assault_phys, binge_fr, assault_sex, assault_emo, psyhx, isi_tot, fincur, drug_poly,
    drug_mar, drug_coc, drug_her, drug_met, drug_stim, drug_ect, drug_opi, drug_other, drug_none

Motivational variables
  - Defeat: aca_impac, men_ill_ash, deprawsc **
  - Threats to Self-Moderation: persist, anx_score, dep_secret, AAQ_tot, achieve4 **
  - Entrapment: dep_impa, gad7_impa, percneed_cur **
  - Motivational Moderators: inf, flourish, belong1, belong2, belong8, belong9, stig_pcv_2, stig_pcv_3, BRS_tot 
  
* Will show demographics from original datasets with descriptive stats.
demographic tables... show up on the survey - acknowledge diversity
  didnt want to erase or collapse, but for the purpose of the analysis
  look at college literature -- categories higher risk
  
run correlation matrix for each  
  

```{r Filter sample, cache = TRUE}

#IMPORT file
 # data.raw <- read_excel("~/Desktop/Desktop_Home/UTAH/Stats_I/R_Studio/Dissertation_files/Sisler_dissertation_HMS1819_audit.xlsx",
 #                        range=cell_cols(1:1820), col_types = "text")
 # 
 # save("data.raw", file="~/Desktop/Desktop_Home/UTAH/Stats_I/R_Studio/Dissertation_files/Sisler_dissertation_HMS1819_audit.RData")
 load("~/Desktop/Desktop_Home/UTAH/Stats_I/R_Studio/Dissertation_files/Sisler_dissertation_HMS1819_audit.RData")
 # source("dissertation_fx.R")
   
data.v1 <- data.raw %>%
  mutate(age = as.numeric(age)) %>%
  filter(age < 26 & age >= 18) %>%
  select(-c(

    # demographic info
    children_dep, educ_par1_rel, educ_par2_rel, hours_work_paid, timeclass, timestud, educ_par1, educ_par2, age_US, langpast, langcur, 
    age18:age41pl, educ_par, eductemp, educ_1:educ_4, age18_22, age31pl,
    # relationship info
    relship, relship_text, rel_sing:rel_divw,
    #  HMS-based
    SweepsIncentiveContact, endmessageSuicide, UniversityofSouthernCalifornia1a, consentFollowup1, excl_pers_oth_TEXT, exp_rude, 
    exp_accuse, exp_othersafraid, exp_ideaignore, exp_offendjoke, exp_stupid, exp_notserious, exp_superficial, exp_re_prof, exp_re_fear, 
    exp_re_speakall,
    #  religious affiliations
    relig_aff_ag:relig_aff_oth, relig_aff_other_text, relig_1:relig_5,
    #  residence
    res_dorm:res_oth, residenc_text, deport_worry,
    #  degree categories
    degree_ass:degree_other_text, enroll_text, deg_ass:deg_nd,
    #  major/field choices
    field_hum:field_other, field_other_text,
    #  unused scales
    phq2_1, phq2_2,
    #  sex abuse/assault:
    ##  type
    assault_sex_stranger:assault_sex_dontknow, abuse_recent_perpet, sa_exp_touch:sa_exp_pen, sa_suc_phyforc_1, sa_att_phyforc, 
    sa_drunk_1, sa_suc_phyforc_2_1:sa_suc_phyforc_2_6, sa_suc_phyforc_2_6_text, aca_phys_assault_1:aca_phys_assault_7, 
    aca_sex_assault_1:aca_sex_assault_7,
    ##  reporting
    sa_tell_1:sa_tell_12, sa_tell_12_text, sa_nodiscl_1:sa_nodiscl_24, sa_nodiscl_24_text,
    ##  person
    sa_who_1:sa_who_12, sa_who_12_text, sa_who_stud, sa_who_employ, sa_who_gender, sa_who_gender_text,
    ##  location/context
    sa_loc_on, sa_op_substance_1:sa_op_substance_5, sa_frighten, sa_loc_off, sa_perc_1:sa_perc_12, sa_train_pol, sa_train_prev,
    #  sleep variables
    sleep_wk1, sleep_wd1, sleep_wk2, sleep_wd2, sleep_np1, sleep_np2,
    #  stress variables
    stress1:compet_field,
    #  D/O SUBTYPES:
    ##  mood
    dx_dk, dx_dep_1:dx_bip_4, dep_maj, dep_oth, dx_bip_bi1:dx_bip_oth, dep_imp1:dep_imp4,
    ##  anxiety
    dx_ax_1:dx_ax_7_new, dx_ax_6_text_new, aca_anx_1:aca_add_7, dx_ax_ga:dx_ax_so,
    ##  OCD
    dx_ocd_1:dx_ocd_6, dx_ocd_6_text, dx_ocd_o:dx_ocd_oth,
    ##  trauma
    dx_trauma_1:dx_trauma_4, dx_trauma_4_text, dx_trauma_ptsd:dx_trauma_oth,
    ##  eating disorder
    dx_ea_1:dx_ea_4_text, ed_thin, weight_ideal, diet, ea_fast, purge_1:purge_4, ed_peer, body_con_change, fresh15, weigh_freq, 
    weight_ideal_text, dx_ea_an:dx_ea_av,
    ##  psychosis
    dx_psy_1:dx_psy_8, dx_psy_7_text, dx_psy_sa:dx_psy_sp,
    ##  substance-related
    dx_sa_1, dx_sa_2, dx_sa_3, dx_sa_2_text, aca_substance_1:aca_substance_7, sa_drunk_2_1:sa_drunk_2_7, 
    sa_drunk_2_6_text, sa_drunk_3:sa_drunk_6, alc_prob, sub_cig1:sub_cig5, sa_train_use:sa_sch_help,
    ##  personality d/o
    dx_perso_1:dx_perso_12, dx_perso_11_text,
    ##  chronic conditions
    aca_phys_health_1:aca_phys_health_7, dx_chronic_11_text,
    ##  SIBs
    sib_fre1:sub_cig,
    #  provider-related questions:  
    ther_helped_me, prov_1:prov_3, prov_7:prov_10, prov_9_text, anyprovi,
    #  insurance specifics
    ins_2:ins_10, ins_pare:ins_unce, ins_yesb, ins_mh1:ins_mh5, ins_ina_1:ins_ina_8_text,
    #  diet-related
    eatprac_1:eatprac_3, eatprac_frveg, eat_change, eat_change_how_1:eat_change_how_15, eat_change_how_15_text,
    #  SES
    fin_comp2, pay_worry, pay_fam, pay_self, pay_other, pay_grant, pay_loan, fincur_1:finpas_5,
    #  CCAPS
    CCAPS_1:CCAPS_34,
    #  CCMH:  
    sib_ccmh_1:suic_ccmh_6, cons_viol_ccmh_number, caus_viol_ccmh_number, exp_har_ccmh_number, exp_trauma_ccmh_number, 
    cons_viol_ccmh_recent, caus_viol_ccmh_recent, exp_har_ccmh_recent, exp_trauma_ccmh_recent, trauma_ccmh_text, traumevent_ccmh_1, 
    traumevent_ccmh_2, traumevent_ccmh_3, traumevent_ccmh_4, traumevent_ccmh_5, traumevent_ccmh_6, traumevent_ccmh_7, traumevent_ccmh_8, 
    traumevent_ccmh_9, traumevent_ccmh_10, traumevent_ccmh_11, traumevent_ccmh_12, traumevent_ccmh_13, traumevent_ccmh_14, 
    traumevent_ccmh_15, traumevent_ccmh_16, traumevent_ccmh_15_text, binge_fr_ccmh,
    #  HMS-CCMH
    dep_CCMH_score_temp, dep_CCMH_score, dep_CCMH_elevated, dep_CCMH_mild, dep_CCMH_none, genanx_CCMH_score_temp, genanx_CCMH_score, 
    genanx_CCMH_elevated, genanx_CCMH_mild, genanx_CCMH_none, socanx_CCMH_score_temp, socanx_CCMH_score, socanx_CCMH_elevated, 
    socanx_CCMH_mild, socanx_CCMH_none, eatconcerns_CCMH_score_temp, eatconcerns_CCMH_score, eatconcerns_CCMH_elevated, 
    eatconcerns_CCMH_mild, eatconcerns_CCMH_none, dep_or_anx_CCMH, sib_ccmh_11:sib_ccmh_15, sib_ccmh_lifetime_any, sib_ccmh_pastyear,
    #  Qnum:  
    Q2_11, Q2_12, Q3_8, Q3_9, Q3_10_2, Q3_10_1, Q3_11_1, Q3_12_1:Q3_12_5, Q2_10_152:Q2_10_155, Q4_21_11, Q9_19, Q9_20,
    #  health knowledge: 
    know_sp, dep_tx_know_1:dep_tx_know_4, dep_sx_know_1:dep_sx_know_4, anx_help_know_1:anx_help_know_4, ea_sx_know_1:ea_sx_know_6,
    #  therapy-related
    why_tx_1:why_tx_9_text, ther_lifetime, ther_lifetime2, ther_any, talk1_any, talk1_8_text, ther_any2, knowwher1:knowwher6, 
    bar_any:bar_plan, ther_help_me1:ther_help_me4, meds_help_me1:meds_help_me4, talksup1:talksup4, talk2pro:talk2noo,
    #  school climate
    cli_disab:cli_sh_oth_TEXT, excl_env:exp_community, val_faculty:excl_pers_oth, group_TEXT:group_active, friend_re:friend_religid, 
    social_re:social_religid, re_find:time_activ,
    #  SAT-related
    sat_hours_1, sat_loc_1, sat_qual_1, sat_priv_1, sat_sched_1, sat_hours_2, sat_loc_2, sat_qual_2, sat_priv_2, sat_sched_2, 
    sat_hours_3, sat_loc_3, sat_qual_3, sat_priv_3, sat_sched_3, sat_hours_4, sat_loc_4, sat_qual_4, sat_priv_4, sat_sched_4, 
    sat_hours_5, sat_loc_5, sat_qual_5, sat_priv_5, sat_sched_5, sat_hours_6, sat_loc_6, sat_qual_6, sat_priv_6, sat_sched_6, 
    sat_hours_7, sat_loc_7, sat_qual_7, sat_priv_7, sat_sched_7, sat_hours_8, sat_loc_8, sat_qual_8, sat_priv_8, sat_sched_8, 
    sat_hours_9, sat_loc_9, sat_qual_9, sat_priv_9, sat_sched_9, sat_hours_1_1:sat_hours_1_6, sat_loc_1_1:sat_loc_1_6, 
    sat_qual_1_1:sat_qual_1_6, sat_priv_1_1:sat_priv_1_6, sat_sched_1_1:sat_sched_1_6, sat_hours_2_1:sat_hours_2_6, 
    sat_loc_2_1:sat_loc_2_6, sat_qual_2_1:sat_qual_2_6, sat_priv_2_1:sat_priv_2_6, sat_sched_2_1:sat_sched_2_6, 
    sat_hours_3_1:sat_hours_3_6, sat_loc_3_1:sat_loc_3_6, sat_qual_3_1:sat_qual_3_6, sat_priv_3_1:sat_priv_3_6, 
    sat_sched_3_1:sat_sched_3_6, sat_hours_7_1:sat_hours_7_6, sat_loc_7_1:sat_loc_7_6, sat_qual_7_1:sat_qual_7_6, 
    sat_priv_7_1:sat_priv_7_6, sat_sched_7_1:sat_sched_7_6, sat_hours_8_1:sat_hours_8_6, sat_loc_8_1:sat_loc_8_6, 
    sat_qual_8_1:sat_qual_8_6, sat_priv_8_1:sat_priv_8_6, sat_sched_8_1:sat_sched_8_6,
    #  peer-related
    peer_alc_1, peer_alc_3, peer_alc_4, peer_mar_est, peer_alc_est, peer_cig_est, inf_7_text, risk_cig:risk_presc,
    #  pregnancy-related
    sex_partner, sex_partner_f, sex_partner_m, sex_partner_t, sex_30, sex_30_oral, sex_30_vag, sex_30_anal, 
    birthcontrol_1:birthcontrol_12, birthcontrol_10_text, birthcontrol_always, preg_no, preg_yes_u,preg_yes_in, preg_dk, preg_cur,
    #  medication details
    meds_reason_1:meds_reason_5, meds_reason_5_text, meds_dis, meds_w_1:meds_w_5, meds_w_3_text, meds_cur_1:meds_cur_9, 
    meds_cur_7_text, meds_time_1:meds_time_7, med_help, meds_helped_me, meds_w_g:meds_di5,
    #  campus or community
    heard, camp_supp, outreach_aware, gk_1, gk_2, gkt_any, txfrf, cli_look, cli_resp_fr, 
    cli_resp_cl, cli_free, st_promote, admin_listen, env_body, safe_on_day, safe_on_night, safe_off_day, 
    safe_off_night, choose_sch, persist_challenge_1:persist_challenge_11, persist_challenge_11_text, highestdeg, highestdeg_text,
    doubt_school_1, doubt_school_2, prof_support_aca, different_maj, late_assign, sleep_class, aca_support_serv, office_hrs, faculty_out,
    advisor_interact, gradstud_interact, time_manage, grade_curv, timestud_1:timestud_7,
    #  up/bystanding:
    inter_hy_dr, inter_hy_sa, inter_hy_hl, inter_y_1:inter_y_7, inter_y_6_text, inter_help, inter_n_1:inter_n_7, inter_n_6_text, 
    inter_n_why_1:inter_n_why_9, inter_n_why_9_text, witness_1:witness_7, witness_6_text, inf_any:inf_oth,
    #  hypothetical:
    ther_help, talk2_1:talk2_7, talk2_7_text, talksup, percneed1:percneed_cur6, stig_t_I:stig_t_I6,
    #  misc:  
    bar_hs_1:bar_ns_8_text, smok_freq, sib_other_text, activ_other_text, knowwher_temp, exerc_changed, exerc_changed_how, 
    exerc_changed_how_text, height_ft, height_in, wgt_lbs, prov_A, prov_B, prov_C,  prov_loc, prov_olo, prov_oth, prov_dk, adjust_aca_1, 
    adjust_aca_2, aca_imp1:aca_imp4, s_c_1_1:s_c_5_6, s_n_1_1:s_n_5_6))
```

# Call functions
```{r Functions}
# delogodd
delogodd <- function(x) {
  oddsratio <- exp(x)
  p <- oddsratio / (1 + oddsratio)
  return(p)
}

# debug_contr_error
debug_contr_error <- function (dat, subset_vec = NULL) {
  if (!is.null(subset_vec)) {
    ## step 0
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
      }
      subset_log_vec <- subset_vec
    } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
      } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
      } 
    } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
    }
    dat <- base::subset(dat, subset = subset_log_vec)
  } else {
    ## step 1
    dat <- stats::na.omit(dat)
  }
  if (nrow(dat) == 0L) warning("no complete cases")
  ## step 2
  var_mode <- sapply(dat, mode)
  if (any(var_mode %in% c("complex", "raw"))) stop("complex or raw not allowed!")
  var_class <- sapply(dat, class)
  if (any(var_mode[var_class == "AsIs"] %in% c("logical", "character"))) {
    stop("matrix variables with 'AsIs' class must be 'numeric'")
  }
  ind1 <- which(var_mode %in% c("logical", "character"))
  dat[ind1] <- lapply(dat[ind1], as.factor)
  ## step 3
  fctr <- which(sapply(dat, is.factor))
  if (length(fctr) == 0L) warning("no factor variables to summary")
  ind2 <- if (length(ind1) > 0L) fctr[-ind1] else fctr
  dat[ind2] <- lapply(dat[ind2], base::droplevels.factor)
  ## step 4
  lev <- lapply(dat[fctr], base::levels.default)
  nl <- lengths(lev)
  ## return
  list(nlevels = nl, levels = lev)
}

# debug_contr_error2
debug_contr_error2 <- function (form, dat, subset_vec = NULL) {
  ## step 0
  if (!is.null(subset_vec)) {
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
      }
      subset_log_vec <- subset_vec
    } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
      } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
      } 
    } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
    }
    dat <- base::subset(dat, subset = subset_log_vec)
  }
  ## step 0 and 1
  dat_internal <- stats::lm(form, data = dat, method = "model.frame")
  attr(dat_internal, "terms") <- NULL
  ## rely on `debug_contr_error` for steps 2 to 4
  c(list(mf = dat_internal), debug_contr_error(dat_internal, NULL))
}

# NA_preproc
NA_preproc <- function (dat) {
  for (j in 1:ncol(dat)) {
    x <- dat[[j]]
    if (is.factor(x) && anyNA(x)) dat[[j]] <- base::addNA(x)
    if (is.character(x)) dat[[j]] <- factor(x, exclude = NULL)
  }
  dat
}

# fill.na.with.mode
## Model fitting per group and NA as factor levels
fill.na.with.mode = function(df){
  cols = colnames(df)
  for (col in cols){
    if(class(df[[col]])=='factor'){
      x = summary(df[[col]])
      mode = names(x[which.max(x)])
      df[[col]][is.na(df[[col]])]=mode
    }
    else{
      df[[col]][is.na(df[[col]])]=0
    }
  }
  return (df)
}

# corr_simple -- by https://towardsdatascience.com/how-to-create-a-correlation-matrix-with-too-many-variables-309cc0c0a57
corr_simple <- function(data=corr_vars,sig=0.5){
  #convert data to numeric in order to run correlations
  #convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA
  df_cor <- data %>% mutate_if(is.character, as.factor)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)
  #run a correlation and drop the insignificant ones
  corr <- cor(df_cor)
  #prepare to drop duplicates and correlations of 1     
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  #drop perfect correlations
  corr[corr == 1] <- NA 
  #turn into a 3-column table
  corr <- as.data.frame(as.table(corr))
  #remove the NA values from above 
  corr <- na.omit(corr) 
  #select significant values  
  corr <- subset(corr, abs(Freq) > sig) 
  #sort by highest correlation
  corr <- corr[order(-abs(corr$Freq)),] 
  #print table
  print(corr)
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")
  
  #plot correlations visually
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}

```


# OUTCOME VARIABLE

SUICIDAL IDEATION (si_type)
Creates indicator variable 'si_type' using the subtypes of suicidal ideation using sui_idea, sui_plan, sui_attempt, where: 
  0/0/0 = no SI
  1/0/0 = passive SI
  1/1/0 = active SI
  1/1/1 = volitional
+/- 'phq9_9' and 'sui_idea' into one variable -- suic_idea. Will run the end model twice.


```{r si_type}

data.v2 <- data.v1 %>%

# create si_type var
  mutate(si_type = case_when((sui_idea == 1 & suic_plan == 1 & suic_att == 1) ~ 4,   # volitional, suicide attempt or related bx
                           (sui_idea == 1 & suic_plan == 1) ~ 3,                   # active SI
                           (sui_idea == 1) ~ 2,                                    # passive SI
                           sui_idea == 0 ~ 1)) %>%                                 # no SI reported
  mutate(si_type = factor(si_type, levels=1:4,
                        labels=c("No SI", "passive SI", "active SI", "SA")))
  
# tabulate and check output
table(data.v2$si_type, useNA = "always")
table(data.v2$sui_idea)
data.v2 = subset(data.v2, select= -c(suic_plan, sui_plan, suic_att, sui_att)) # clean up vars/columns

# select if SI variable is missing -- then rowwise deletion
data.v2 <- data.v2 %>%
  filter(!is.na(si_type)) 
table(data.v2$si_type, useNA = "always") # ensuring that all NAs are now clean

# Define starting sample (n = 51137)
length(unique(data.v2$responseid))

```


# DIATHESIS VARIABLES

Candidate variables: age, sexual, gender, finpast, sib_freq, sib, bmi, race, ed_scoff, body_sr, audit_tot, dx_dep, dx_any, dx_bi, dx_psy, dx_ea, dx_tr, dx_sa, dx_adhd, dx_pers, dx_none, med_any, med_sti, med_dep, med_psy, med_anx, med_mood, med_sle, meds_other, meds_none, alc_tx, wcs_tot, med_count, dx_chronic, disab_1


```{r Diathesis frame}
data.v3 <- data.v2
```

## Age
```{r age}
table(data.v3$age)
```

## Gender identity

- Original (n=55138): 1=Male; 2=Female; 3=Trans male/Trans man; 4=Trans female/Trans woman; 5=Genderqueer/Gender non-conforming, 6=Other, self-identify
- Transformed (n=55138): 1 = cis-gender, 2=transgender 3=gender diverse, 4=non-answer (NA, jokesters, decline to answer)

```{r gender}
data.v3 <- data.v3 %>%
  filter(!is.na(gender)) %>% 
  # Create 'gender' subtypes into one category
  mutate(gender = case_when((gender == 3 | gender == 4) ~ 3,     # transgender
                            (gender == 5 | gender == 6) ~ 2,     # gender diverse
                            (gender == 1 | gender == 2) ~ 1,     # cis-gender
                            FALSE ~ 4)) %>%                      # sets up non-answers
  # label broad gender identities
  mutate(gender = factor(gender, levels=1:4,
                         labels=c("cis-gender", "gender diverse", "transgender", "non-answers")))
table(data.v3$gender) # Check output
```

### Reconcile 'gender_text' 
```{r gender_text}
  # gender-diverse
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]emi")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]inary")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[N|n]eutral")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[Q|q]ueer")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[F|f]luid")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]gender")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]ndrogynous")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[P|p]angender")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]igender")] <- "gender diverse"
data.v3$gender[data.v3$gender_text %in% c("she/they", "Closeted genderfluid (female presenting)", "Pan-gender",
                                          "female / infinite star being / gender non conforming","He or they",
                                          "Not sure, nonbinary I guess", "Non-binary trans maculine", "No preference", 
                                          "Somewhere between GNC and transfemme", "She/they",
                                          "Non-binary leaning towards trans man")] <- "gender diverse"

 #transgender
data.v3$gender[data.v3$gender_text %in% c("Transmasculine", "Trans-masculine person", "Trans Masculine", 
                                          "gender non-conforming trans man")] <- "transgender"

  # male
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]ude")] <- "cis-gender"
data.v3$gender[str_detect(data.v3$gender_text, "[M|m]ale")] <- "cis-gender"
data.v3$gender[str_detect(data.v3$gender_text, "[B|b]oy")] <- "cis-gender"
data.v3$gender[data.v3$gender_text %in% c("Male: there are only 2 genders", "if you have a penis you're a male", 
                                          "Cisgender man", "Male motherfucker. stop making up gender shit.", "He-male",
                                          "I have male parts so therefore I am a MALE and there are only TWO genders!!!!!!!!!!!!!!!!!",
                                          "SATSYG (Savage about to steal yo girl", "My sex is male", 
                                          "I say male, but I think gender is fluid", "Theres only two genders, I'm male",
                                          "Probably heterosexual, but so commitment-phobic that I can't actually tell if I'm asexual.",
                                          "I'm a male. Born a male, always will be a male. There should only be two choices.")] <- "cis-gender"

  # female
data.v3$gender[str_detect(data.v3$gender_text, "[F|f]emale")] <- "cis-gender"
data.v3$gender[str_detect(data.v3$gender_text, "[W|w]oman")]  <- "cis-gender" 
data.v3$gender[str_detect(data.v3$gender_text, "[G|g]irl")]  <- "cis-gender"
data.v3$gender[data.v3$gender_text %in% c("Male and Female are not genders, they are sexes. I identify as a woman.", 
                                          "Female, but currently going through a gender identity crisis!")] <- "cis-gender"

  # questioning/other
data.v3$gender[str_detect(data.v3$gender_text, "[Q|q]uestioning")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[U|u]nsure")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "[I|i]dk")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "Hermaphrodite")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "Questioning")] <- "gender diverse"
data.v3$gender[str_detect(data.v3$gender_text, "Undecided")] <- "gender diverse"
data.v3$gender[data.v3$gender_text %in% c("Literally no clue as of late", "Kynigender", "Adrogynous", "Hermaphrodite", "idk yet", 
                                          "Questioning", "Questioning; female", "Still questioning", "not female/questioning", 
                                          "Questioning (unsure)", "Girl ? Idk", "not sure", "Not Sure???", "Not sure", 
                                          "I don't know, still trying to purse it out", "Undecided")] <- "gender diverse"

  # non-answers jokesters
data.v3$gender[str_detect(data.v3$gender_text, "[H|h]elicopter")] <- "non-answers"
data.v3$gender[str_detect(data.v3$gender_text, "[C|c]hinook")] <- "non-answers" 
data.v3$gender[str_detect(data.v3$gender_text, "[A|a]pache")] <- "non-answers"
data.v3$gender[str_detect(data.v3$gender_text, "[U|u]nicorn")] <- "non-answers"
data.v3$gender[data.v3$gender_text %in% c("Albino stingray", "Karate Chop Action", "Flying Spaghetti Monster", "Pio Nono", "God", 
                                          "None", "Baseball bat", "Mountain lion", "Oil Platform", "Chair", "M1A1 Main Battle Tank",
                                          "This doesn't matter and you're an idiot for thinking it does.","REZZBIAN",
                                          "Baguette", "fairy", "a 5 year old girl", "tree", "unicorn", "His Majesty", 
                                          "Call of Duty: Modern Warfare 2", "Attack hellcopter", "Meme God", "AH-64D", 
                                          "Stainless Steel Toaster", "God himself", "baseball bat", "Mountain","Trans Former", 
                                          "Potato", "I will literally give you $100 to h*ck off", "Indian tape worm", 
                                          "Metallic Hyper-Entity", "2X4 treated lumber", "pizza", "sPecIaL SNowFlAkE", "Deppressed",
                                          "eh", "George Michael", "Metaknight Main", "Microwave", "Pizza", "Sea turtle", 
                                          "1941 Nihon Giagun Yamato-Class Battleship", "penguin wearing a sombrero and sunglasses", 
                                          "Fire hydrant/hockey puck", "Attack Hellicopter", "The Senate", 
                                          "I identify as a fruit cup")] <- "non-answers"

  # non-answers, decline to answer
data.v3$gender[str_detect(data.v3$gender_text, "[D|d]ecline")] <- "non-answers"
data.v3$gender[data.v3$gender_text %in% c("I am me. A human.", "I'm not yet sure where I identify", "Human-being", "genders", 
                                          "okay", "guess", "none", "Just my name.", "the above options are sex, not genders", 
                                          "the above options are sex, not genders", "its difficult", "any", "?", 
                                          "Biological male who identifies as a lesbian female.", "Prefer not to answer",
                                          "The universe conscious of itself",  "no",  "I only try to use gender when it's useful", 
                                          "fuck gender", "Gender and sex are the same", "gender is a lie", "#love", "No",
                                          "it's weird to separate out trans male and trans female if you're also asking for birth assignment.", 
                                          "2 genders", "I don't understand the question.", "Prefer not to specifiy",
                                          "'Male' and 'female' are not gender identities.",
                                          "I don't have one because it's not a legitimate concept.", 
                                          "Gender is a construct which serves no purpose.")] <- "non-answers"

table(data.v3$gender, useNA = "always")   # check output
```

## Sexual Identity
  - Original: 1=heterosexual, 2=lesbian, 3=gay, 4=bisexual, 5=queer, 6=questioning, 7=other
  - Transformed_1: 1=heterosexual, 2=lesbian/gay 3=bisexual, 4=queer/questioning, 5= other, 6=pansexual, 7=asexual/gray/demi, 8=non-answers
  -- can do interaction with gender and sexuality post-analysis
```{r sexual}
data.v4 <- data.v3 %>%
  # aggregates all 'sexual' answers into one var
  mutate(
    sexual = case_when(sexual_queer == 1 ~ 5,                        # queer
                       sexual_bi == 1 ~ 4,                           # bisexual
                       sexual_l == 1 ~ 2,                            # lesbian 
                       sexual_g == 1 ~ 3,                            # gay
                       sexual_quest == 1 ~ 6,                        # questioning
                       sexual_h == 1 ~ 1,                            # heterosexual
                       sexual_other == 1 ~ 7,                        # other
                       FALSE ~ 8,                                    # non-answers
                       FALSE ~ 9,                                    # creating 'pansexual' var
                       FALSE ~ 10)) %>%                              # creating 'asexual/demi' var
  mutate(sexual = factor(sexual, levels=1:10,
                         labels=c("Heterosexual", "Gay", "Lesbian", "Bisexual", "Queer", "Questioning",
                                  "Other", "Non-answers", "Pansexual", "Asexual/Demi")))

table(data.v4$sexual, useNA = "always") # tabulate and check outputs

```

### Reconcile sexual_text
```{r sexual_text}

# identifying non-answers
  ## jokesters
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]elicopter")] <- "Non-answers"
data.v4$sexual[data.v4$sexual_text %in% c("Yes", "i identify as a lamp", "I only like baseballs", "helisexual", "Autosexual",
                                          "now", "Miguel Garcia", "toasteroven", "must have gills", "Potsexual", "Bobcats",
                                          "Tailpipes turn me on", "Tyrannosaurus rex", "Traps are not gay", "Toaster sexual", 
                                          "Crooked", "Pantsexual", "FREE", "Frozen Foods", "Giraffe", "I'm sexually attracted to fire",
                                          "Attracted to other self-identifying flying spaghetti monsters", "sPecIaL SNowFlAkE", 
                                          "Demiqueer quasiflux asexual foxkin", "Attracted to Vehicles", "Elitistual. I'm into elitists",
                                          "I sleep around. Also into bestiality",  "Sexually attracted to kitchen appliances",
                                          "sexually attracted to fictional characters and men twice my age")] <- "Non-answers"
  ## declined to answer
  data.v4$sexual[str_detect(data.v4$sexual_text, "identi[t|f]y")] <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "label")] <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "matter")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[W|w]hatever")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "box myself")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]ecline")] <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]ormal")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]isclose")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]one")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[U|u]nknown")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]/[A|a]")]  <- "Non-answers"
  data.v4$sexual[str_detect(data.v4$sexual_text, "[I|i] like")]  <- "Non-answers"
  data.v4$sexual[data.v4$sexual_text %in% c("dating a trans male", "Attracted to personalities, not genders", "non", "irrelevant",
                                            "i like people", "Fuck sexuality", "personality based", "Don't think about it too much",
                                            "Anyone that I like", "leave me alone", "define", "This is more bullshit", "No", "Me",
                                            "I am who I am", "Just there", "Don't care", "no label", "No Label", "No labels",
                                            "Questions like this are the reason we are even having this debate.", "i like who i like", 
                                            "prefer not", "I do not understand why this is necessary", "Just there", "non", "Normal",
                                            "Human Being", "It doesn't matter my question is why is this an actual question", 
                                            "I hate this PC bs", "FUCK YOU BIG TIME", "No", "Me", "human being", "Closeted at home", 
                                            "Polyamorous", "Sexuality shouldn't matter", "No preference", "i like boys.", 
                                            "gay leaning bi", "other", "Prefer not to specify", "i fall in love with a human being",
                                            "everyone's a little gay", "I prefer men, but have had women before", "Traps are not gay",
                                            "Unknown", "#love", "Whatever I feel like at the time", "Situational",  "No one right now", 
                                            "sometimes i dabble", "People", "Prefer not", "I am", "Whatever Happens", "Crooked", 
                                            "love is love", "Open", "Autistic", "Human", "No one right now", "Situational", "I don't",
                                            "The universe conscious of itself", "Aromantic", "Not dealing with men or women at the moment",
                                            "Repressed", "?", "Undefined", "Never had sex", "None?", "neutral", "not sure", "Not sure",
                                            "no sexual orientation", "No idea, but I'm not straight", "Unlabeled", "Curious",  
                                            "Anyone that I like", "Disinterested", "Do not define", "Fake Gay", "irrelevant", "People", 
                                            "No preference", "love is love", "no particular sexual identity", "there is no such thing", 
                                            "Sometimes gender doesn't matter and I don't have time to learn about these things life gets too busy",
                                            "Prefer not to answer", "just sexual (need no prefix!)", "It's complicated", 
                                            "Sexuality shouldn't matter", "I'm just me and I love who I love. I don't like these labels.",
                                            "I'm not a fucking part of lgbtqrstuxyz bullshit association")]  <- "Non-answers"
  
    # identifying queer  individuals:
data.v4$sexual[str_detect(data.v4$sexual_text, "[Q|q]ueer")] <- "Queer"
data.v4$sexual[str_detect(data.v4$sexual_text, "[F|f]luid")] <- "Queer"
data.v4$sexual[str_detect(data.v4$sexual_text, "[O|o]pen")] <- "Queer"
data.v4$sexual[data.v4$sexual_text %in% c("potentially interested", "somewhere between", "homoflexible", "heteroflexible", "fluid",
                                          "Only other trans people", "non-binary", "so I guess I could check queer", 
                                          "Only other trans people", "Androsexual", "Not straight", "Polysexual", "Joto",
                                          "Attracted to masculinity and androgony", "Androsexual", "Non-binary", "nonbinary")] <- "Queer"

  # identifying questioning
data.v4$sexual[str_detect(data.v4$sexual_text, "[Q|q]uestion")] <- "Questioning"
data.v4$sexual[str_detect(data.v4$sexual_text, "[P|p]robably")] <- "Questioning"
data.v4$sexual[data.v4$sexual_text %in% c("possibly", "curious", "Unknown", "potentially interested", "don't know", "Unknown", 
                                          "don't know", "not sure", "still figuring it out", "idk", "Idk",  "don't know", 
                                          "I have a girlfriend but i am not attracted to any other females. I am attracted to males", 
                                          "Sometimes gender doesn't matter and I don't have time to learn about these things life gets too busy",
                                          "Not sure. Working on it.")] <- "Questioning"

# identifying  pansexual write-ins
data.v4$sexual[str_detect(data.v4$sexual_text, "[P|p]an")] <- "Pansexual"
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]apiosexual")] <- "Pansexual"
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]apiosexual")] <- "Pansexual"
data.v4$sexual[data.v4$sexual_text %in% c("all", "Polyamorous Pansexual", "Pan Sexual", "Pan-sexual", 
                                          "any but ive only had relationships with men")] <- "Pansexual"

# identifying asexual / demisexual write-ins:  
data.v4$sexual[str_detect(data.v4$sexual_text, "[G|g]r[a|e]y")] <- "Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[N|n]onsexual")] <-"Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[C|c]elibate")] <- "Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]emi")] <- "Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[A|a]sex")] <- "Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[C|c]elibate")] <- "Asexual/Demi"
data.v4$sexual[str_detect(data.v4$sexual_text, "[A|a]ce")] <- "Asexual/Demi"
data.v4$sexual[data.v4$sexual_text %in% c("Mostly asexual", "Most likely asexual", "Masexual", "Tridium Vykon Jace 8000", 
                                          "Demi-romantic", "asexual lesbian", "Ase", "Bi-demisexual",  "Hetroromantic Asexual",
                                          "Demiqueer quasiflux asexual foxkin", "DemiAsexual", "a-sexual", "no sexual attraction", 
                                          "Gay-ish, I am Demi-sexual which means I don't pay attention to sex/gender", "a sexual",
                                          "Pan-Quoiromantic Asexual", "demiromantic asexual", "Homoromantic Demisexual",
                                          "experience no sexual attraction but do experience a small amount of romantic attraction", 
                                          "lithromantic", "A sexual")] <- "Asexual/Demi"

  # identifying heterosexual individuals
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]etero")] <- "Heterosexual"
data.v4$sexual[str_detect(data.v4$sexual_text, "[S|s]traight")] <- "Heterosexual"
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]etrosex")] <- "Heterosexual"
data.v4$sexual[data.v4$sexual_text %in% c("Kinsey scale 1", "I would put myself as a 1-1.5 on the kinsey scale", "Staight", 
                                          "Stragiht", "strait", "Str8!!!!", "Changing from gay to heterosexual", "Gynophilic",
                                          "Into females", "I fuck bitches.", "1 on the Kinsey Scale", "Gynophilic", "gynosexual",
                                          "Changing from gay to heterosexual", "100% Hetro")] <- "Heterosexual"

  # identified as gay or lesbian
data.v4$sexual[str_detect(data.v4$sexual_text, "[G|g]ay")] <- "Gay"
data.v4$sexual[str_detect(data.v4$sexual_text, "[H|h]omoflexible")] <- "Gay"
data.v4$sexual[data.v4$sexual_text %in% c("Mostly gay", "soooooo gay", "Romantically straight and sexually gay")] <- "Gay"
data.v4$sexual[str_detect(data.v4$sexual_text, "[L|l]esbian")] <- "Lesbian"
data.v4$sexual[str_detect(data.v4$sexual_text, "[D|d]yke")] <- "Lesbian"

  # identifying bisexual individuals
data.v4$sexual[str_detect(data.v4$sexual_text, "[B|b]isex")] <- "Bisexual"
data.v4$sexual[str_detect(data.v4$sexual_text, "[B|b]icurious")] <- "Bisexual"
data.v4$sexual[data.v4$sexual_text %in% c("I'm straight but I like girls too sexually, not emotionally.", "Half bisexual", 
                                          "straight but DTF sexy females", "bicurious but i like girls more", 
                                          "hetero with occasional bi tendencies", "gay leaning bi", "Bi-curious", 
                                          "Herero-romantic bisexual", "Bi with a stronger attraction to same sex",
                                          "I'm like halfway to bisexual. Never had a crush on another girl but I'm still interested. Idk.",
                                          "No boundaries in place - though probably bisexual")] <- "Bisexual"

table(data.v4$sexual, useNA = "always")   # tabulate and check outputs
```

### 'Sexual' transformed 
sexual identity: (1) heterosexual; (2) lesbian or gay; (3) bisexual; (4) queer/pan/asex/other, and (5) non-answers
```{r sexual regroup}

data.v5 <- data.v4 %>%
    filter(!is.na(sexual)) %>% 
  # aggregates all 'sexual' answers into one var
  mutate(
    sexual = case_when(sexual %in% c("Pansexual", "Queer", "Asexual/Demi") ~ 4, # queer, nonbinary
    sexual == "Questioning" ~ 5,                                                # questioning
    sexual %in% c("Gay", "Lesbian") ~ 2,                                        # gay, lesbian
    sexual == "Bisexual" ~ 3,                                                   # bisexual
    sexual == "Heterosexual" ~ 1,                                               # heterosexual
    sexual == "Other" ~ 4,                                                      # other
    sexual == "Non-answers" ~ 6)) %>%                                           # non-answers
  mutate(sexual = factor(sexual, levels=1:6,
                       labels=c("Heterosexual", "Gay/Lesbian", "Bisexual", "Queer/non-binary", "Questioning", "Non-answers")))
table(data.v5$sexual) # tabulate and check outputs


```

## BMI
```{r bmi}
data.v5$bmi <- as.integer(data.v5$bmi)
table(data.v5$bmi)
```

## Chronic conditions
n = 622, pos  ** too small to include
```{r Chronic conditions}
table(data.v5$dx_chronic_1)
table(data.v5$dx_chronic_2)
table(data.v5$dx_chronic_3)
table(data.v5$dx_chronic_4)
table(data.v5$dx_chronic_5)
table(data.v5$dx_chronic_6)
table(data.v5$dx_chronic_7)
table(data.v5$dx_chronic_8)
table(data.v5$dx_chronic_9)
table(data.v5$dx_chronic_11)
table(data.v5$dx_chronic_12)
table(data.v5$dx_chronic_13)

data.v5 = subset(data.v5, select= -c(dx_chronic_1:dx_chronic_13)) # drop chronic as possibly flawed variable and low sample
```

## Subjective weight and body image
```{r body_sr}
table(data.v5$body_sr)
```

## Race
```{r race}  
data.v6 <- data.v5 %>% 
# 'race' subtypes combined into one variable
  mutate(race = case_when(race_bla == 1 ~ 1,
                          race_ame == 1 ~ 2,
                          race_asi == 1 ~ 3,
                          race_his == 1 ~ 4,
                          race_pac == 1 ~ 5,
                          race_ara == 1 ~ 6,
                          race_whi == 1 ~ 7,
                          race_oth == 1 ~ 8,
                          FALSE ~ 9)) %>%
    mutate(race = factor(race, levels=1:9,
                         labels=c("Black", "AI/AN", "Asian", "Hispanic", "Pac Island", "ME/Arab", 
                                  "White", "Other/Mixed", "Non-answers")))
table(data.v6$race) # check output
```


### Race write-in reconcilation
Reconciles all 'other' answers vi race_other_text
```{r race reconciliation}
  #decline to answer
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]merica")] <- "Non-answers"
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]uman")] <- "Non-answers"
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]one")] <- "Non-answers"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]erson")] <- "Non-answers"
data.v6$race[data.v6$race_other_text %in% c("prefer not to say", "Prefer not to answer.", "Prefer not to specify", 
                                            "Prefer not to answer :)", "FUCK YOU", "Caramel American", "Why", "irrelevant",
                                            "Prefer not to state", "Prefer no to say", "Race is a problematic social construct",
                                            "Transcendent", "Texan", "Race is a socual construct", "Race/ethnicity should not matter",
                                            "why is this a thing", "Brown", "No thx", "White passing, ethnically other",
                                            "While I understand the socio-economic conditions that surround race, race also is a construct which serves no purpose.",
                                            "Wakanda Forever", "unfortunately", "Prefer not to answer", 
                                            "whatever I can use to be a victim in the given situation", 
                                            "We need to get rid of these color things",  "Me", "I do Not identify racially", 
                                            "Why is there always a race question? Can't we all just be one race? The Human Race.",  
                                            "Will not disclose",  "Race/ethnicity should not matter", "N/a", "N/A", 
                                            "There is only one human race: Homo sapien", "don't know due to uncertainity of lineage",
                                            "This doesn't matter and you're an idiot for thinking it does.",  
                                            "POC because now my opinion will matter in the US", "I choose not to answer",
                                            "Color doesn't mean anything right? Why do so many surveys ask it",
                                            "Citizen of the World", "My skin is lighter than others, and darker than some", 
                                            "I am not a racial identifying person. I am American. My DNA says something about Korean heritage.",
                                             "I do not identify as a race in order to promote social equality for all races",
                                            "I don't think race is all that important.", 
                                            "I do not think that this information should or does make a difference ,so, I chose to not state", 
                                            "A White, Cherokee, European, Irish man who would be preferred to be called else but white",
                                            "I won't allow my race to be used in this study for potential insidious usage")]  <- "Non-answers"

  # Jokesters
data.v6$race[data.v6$race_other_text %in% c("Volvo", "Vulcan", "Snowman", "tree", "General Electric", "Cat", "Earthling", 
                                            "I have rainbow skin.", "jedi", "Light skinned potato", "I'm stainless steel.", 
                                            "7-foot tall Chinese woman", "A little tan.", "Apache helicopter", "Aptenodytes", 
                                            "Wood", "moon man", "Black hawk", "Aviation Heritage", "Ewok", "Desert Camoflauge",
                                            "Aptenodytes", "Attack helicopter", "Ethnic Kekistani")] <- "Non-answers"

  # white
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]ew")] <- "White"  # if Israeli, could be White, Asian, or ME (e.g. Palestine)
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]shkenazi")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]aucasian")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]uropean")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]reek")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]rish")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]talian")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]lbania")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]em[e|i]tic")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[R|r]ussian")] <- "White" # could be Asian depending on region
data.v6$race[str_detect(data.v6$race_other_text, "[R|r]oma")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[T|t]urkish")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]lavic")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]etis")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]asque")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]icilian")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]olish")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]candinavian")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[F|f]innish")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]eltic")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]cottish")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]anad")] <- "White"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]editerranean")] <- "White"
data.v6$race[data.v6$race_other_text %in% c("please refer to me as Caucasian; it's extremely offensive to refer to every race-- excluding Caucasians-- in their politically correct form",
                                            "Not sure if I'm Hispanic, as I am Portuguese and Spanish (from my mother's side) . I identify as white as I am of European descent.",
                                            "Portugese", "Cape Verdean", "Capeverdean", "Bulgarian-American", "Ginger", 
                                            "Anglo Saxon", "anglo/white", "Balkan for ethnicity", "Ulster-scots", 
                                            "I'm white, but I do have some Native American in me but it's not the dominant visible ethnicity.", 
                                            "White (Jewish/Jamaican-Chinese/Irish)", "ARMENIAN")] <- "White"
  # Identifies as American Indian or other indigenous descent
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]ative [A|a]merican")] <- "AI/AN"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndigenous")] <- "AI/AN"
data.v6$race[data.v6$race_other_text %in% c("I look white but im also status native", "Indigenous/Mestizo", 
                                            "american indian that isn'y legally claimed")] <- "AI/AN"
  # Identifies as Asian
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]sian")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hinese")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]apanese")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndian")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]anglades")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[F|f]ilipino")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]uyanese")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ongolia")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hina")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]outheast")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]outh [A|a]sian")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]unjabi")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]iberia")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndonesia")] <- "Asian"
data.v6$race[str_detect(data.v6$race_other_text, "[W|w]est [I|i]ndian")] <- "Asian"
data.v6$race[data.v6$race_other_text %in% c("Sri Lankan", "Sri-Lankan", "third culture kid, born and raised in asia until 18yo",
                                            "Central Asian", "Malays", "Chinese Ethinicy-Nuosu", "Austronesian", 
                                            "Indian, UAE native", "Indian/Subcontinental", "Indians arent on the list.....", 
                                            "American Texan of Indian (South asia) Descent", "Okinawan", "Asian/Chinese",
                                            "Japanese American and Native Hawaiian Culture", "Asian. International.", 
                                            "Australian/Indian","Indo-Caribbean",  "Atlantic Islander", 
                                            "Telugu-speaking Tamil-ethnicity Indian")] <- "Asian"
  # Identifies as Black
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]frican")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]fro")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[K|k]enya")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]lack")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]arribean")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]ahamian")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]aitian")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]aitian")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[J|j]amaica")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[Z|z]imbabwea")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[G|g]hana")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]thiopia")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[T|t]rinidad")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]ritrean")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[W|w]akanda")] <- "Black"
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]igerian")] <- "Black"
data.v6$race[data.v6$race_other_text %in% c("Cameroonian American", "Creole", "Jamacian American", "North Africa")] <- "Black"

  # Identifies as Hispanic
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]exican")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[X|x]ican[o|a]")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[L|l]atin[o|a]")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]uerto [R|r]ican")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[S|s]panish")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]hican[o|a]")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]berian")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]estiz[o|a]")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]ortuguese")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]razil")] <- "Hispanic"
data.v6$race[str_detect(data.v6$race_other_text, "[C|c]uba")] <- "Hispanic"
data.v6$race[data.v6$race_other_text %in% c("Costa Rican", "Central American Native", "Andean")] <- "Hispanic"

  # Identifies as Middle Eastern or of Arab descent
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]rmenia")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]akistani")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]rab")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]ersian")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]alestin")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]srael")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]fghan")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[N|n]orth [A|a]frica")] <- "ME/Arab"
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]azara")] <- "ME/Arab"
data.v6$race[data.v6$race_other_text %in% c("Northern African; Moroccan; Israeli", "Arab Mix", "Berber/North African", 
                                            "Coptic Egyptian (non-Arab)", "Kekistani")] <- "ME/Arab"

  # Identifies all other data or mixed ethicity
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]iracial")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[B|b]i-racial")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ix")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[H|h]alf")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[I|i]ndo-[C|c]arribean")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ulti")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[M|m]ulatto")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[P|p]anda")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[O|o]ther")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[U|u]nkown")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]mbiguous")] <- "Other/Mixed"
data.v6$race[str_detect(data.v6$race_other_text, "[A|a]mbiguous")] <- "Other/Mixed"
data.v6$race[data.v6$race_other_text %in% c("Sub Asian/East African", "two or more races", "Moroccan Mexican", "Transracial", "Adopted",
                                             "Part Thetan", "Taiwanese-Greek-Cypriot American", "East Indian, Caribbean", "50/50",
                                            "Korean Ecuadorian", "Brazilian, Luso, Native American", "Central American/ Caribbean", 
                                            "Mut (more than 3)",  "mexican,filipino, italian", "Middle Eastern, Filipino and Spanish",
                                            "Indo-Portugesse (Fijian/Portugesse) Canadian Born", "White and American Indian (Choctaw)",
                                            "White, black, Native American", "White, Black, and Hispanic", "French Lebanese", 
                                            "Lebanese, German, French", "Asian/Indian", "Black Hispanic", "Black/Asian", 
                                            "An American Mutt- a bit of everything somewhere in there", "Black and native", 
                                            "Black/cuban",  "East Indian, Caribbean", "Filipino and Portuguese", "mixed race",
                                            "Eastern Indian/white", "white middle eastern", "Transracial", "Portuguese/Thai", 
                                            "German, Chinese, Arabic, Melanesian", "Hispanic and Middle Eastern", "mulatto", 
                                            "mixed & south american", "Multiple", "international", "Mixed Race", 
                                            "Biracial", "Multiracial", "Mixed", "Korean Ecuadorian", "Central American/ Caribbean",
                                            "An American Mutt- a bit of everything somewhere in there")] <- "Other/Mixed" 
  # Identifies as black and white as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("Black/White","White/African American", "White/African American")] <- "Other/Mixed"
  # Identifies as Asian and White as 'Other/Mixed'
data.v6$race[str_detect(data.v6$race_other_text, "[E|e]urasian")] <- "Other/Mixed"
data.v6$race[data.v6$race_other_text %in% c("Japanese, Chinese, French, British, Greek", "Asian/White", "Asian/European", 
                                            "Asian and White", "asian and white",  "White (Caucasian) and Indian (S. Asia)", 
                                            "American/Asian African/American White", "3/4 White, 1/4 Japanese", 
                                            "White & Asian",  "Chinese and White", "Chinese/White")] <- "Other/Mixed"
  # Identifies as Hispanic and White as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("I am 1/4th Hispanic 3/4 German",  "Honduran, Irish", "White hispanic",
                                            "White, and Hispanic", "Hispanic/Jewish", "White and Hispanic", "White Hispanic",
                                            "French and spanish", "LATINA/WHITE", "white/hispanic",
                                            "If I identify as puerto rican does that mean i get more scholarship money?")] <- "Other/Mixed"
	 # Identifies as MR and White as 'Other/Mixed'
data.v6$race[data.v6$race_other_text %in% c("Half Middle Eastern half white", 
                                            "Half my family is white and the other middle eastern")] <- "Other/Mixed"

# tabulate and check outputs
table(data.v6$race)
data.v6 = subset(data.v6, select= -c(race_bla, race_ame, race_asi, race_his, race_pi, race_ara, race_whi, race_other))   # clean up vars
```

## SIBs and frequency
Create dichotomous var for SIBs -- sib_any and sib_none -- as well as frequency measure
``` {r sib_any}
data.v7 <- data.v6 %>%
  mutate(sib_any = case_when(sib_none == 1 ~ 0,
                             sib_cut == 1 ~ 1,
                             sib_burn == 1 ~ 1,
                             sib_punc == 1 ~ 1,
                             sib_scra == 1 ~ 1,
                             sib_pull == 1 ~ 1,
                             sib_bite == 1 ~ 1,
                             sib_woun == 1 ~ 1,
                             sib_carv == 1 ~ 1,
                             sib_rub == 1 ~ 1,
                             sib_wall == 1 ~ 1,
                             sib_oth == 1 ~ 1)) %>%
  mutate(sib_freq = case_when(sib_any == 0 ~ 0,
                              sib_freq == 1 ~ 1,
                              sib_freq == 2 ~ 2,
                              sib_freq == 3 ~ 3,
                              sib_freq == 4 ~ 4,
                              sib_freq == 5 ~ 5,
                              sib_freq == 6 ~ 6)) %>%
  mutate(sib_freq = factor(sib_freq, levels=0:6,
                    labels=c("No SIBs", "1-2 tot", "<1 mo", "2-3x mo", "1-2x wk", "3-5 d/wk", "every")))
table(data.v7$sib_freq, useNA = "always")  # tabulate and check output
data.v7 = subset(data.v7, select= -c(sib_none:sib_oth)) # clean up vars/columns
```

## Disability
```{r disab_1}
table(data.v7$disab_1)  # missing question codebook will likely drop if insignificant
```

## SCOFF scale
```{r SCOFF}
table(data.v7$ed_scoff)
data.v7 = subset(data.v7, select= -c(scoff_1:scoff_5)) # clean up vars/columns  
```

## Past finances
```{r finpast}
table(data.v7$finpast)
```

## Co-occuring MH diagnoses
```{r dx_mh}

# To ensure those who skipped the question are not accidentally added to No MH dx category
data.v8 <- data.v7
  
data.v8$dx_adhd = data.v8$dx_neurodev_1

data.v8 <- data.v8 %>%
  mutate(dx_any = case_when(dx_dep == 1 ~ 1,   
                            dx_anx  == 1 ~ 1,
                            dx_bi == 1 ~ 1,
                            dx_psy == 1 ~ 1,
                            dx_ea == 1 ~ 1,
                            dx_tr == 1 ~ 1,
                            dx_sa == 1 ~ 1,
                            dx_adhd == 1 ~ 1,
                            dx_pers == 1 ~ 1,
                            dx_none == 1 ~ 0))
                            
# MH diagnoses sub-category counts
data.v9 <- data.v8 %>%
  filter(!is.na(dx_any)) %>%   # filters out any respondents who skipped the MH section 
  mutate(dx_dep = case_when(dx_dep == 1 ~ 1,
                            dx_none == 1 ~ 0),     # depression
         dx_dep = factor(dx_dep, levels=0:1,
                         labels=c("No dep dx", "Depression"))) %>%
  mutate(dx_anx = case_when(dx_anx == 1 ~ 1,
                            dx_none == 1 ~ 0),     # anxiety
         dx_anx = factor(dx_anx, levels=0:1,
                         labels=c("No anx dx", "Anxiety"))) %>%
  mutate(dx_bi = case_when(dx_bi == 1 ~ 1,
                           dx_none == 1 ~ 0),     # bipolar
         dx_bi = factor(dx_bi, levels=0:1, 
                        labels=c("No BPD dx", "Bipolar"))) %>%
  mutate(dx_psy = case_when(dx_psy == 1 ~ 1,
                            dx_none == 1 ~ 0),    # psychotic
         dx_psy = factor(dx_psy, levels=0:1,
                         labels=c("No psychosis", "Psychosis"))) %>%
  mutate(dx_ea = case_when(dx_ea == 1 ~ 1,
                           dx_none == 1 ~ 0),     # eating d/o
         dx_ea = factor(dx_ea, levels=0:1,
                        labels=c("No EA dx", "Eating"))) %>%
  mutate(dx_tr = case_when(dx_tr == 1 ~ 1,
                           dx_none == 1 ~ 0),     # trauma
         dx_tr = factor(dx_tr, levels=0:1,
                        labels=c("No trauma dx", "Trauma"))) %>%
  mutate(dx_sa = case_when(dx_sa == 1 ~ 1,
                           dx_none == 1 ~ 0),     # substance abuse
         dx_sa = factor(dx_sa, levels=0:1,
                        labels=c("No SA dx", "Substance"))) %>%
  mutate(dx_adhd = case_when(dx_adhd == 1 ~ 1,
                             dx_none == 1 ~ 0),   # adhd
         dx_adhd = factor(dx_adhd, levels=0:1,
                          labels=c("No ADHD dx", "ADHD"))) %>%
  mutate(dx_pers = case_when(dx_pers == 1 ~ 1,
                             dx_none == 1 ~ 0),   # personality
         dx_pers = factor(dx_pers, levels=0:1,
                          labels=c("No personality d/o dx", "Personality"))) %>%
  mutate(dx_none = factor(dx_none, levels=0:1,
                          labels=c("Any dx", "No dx")))

# Check output
table(data.v9$dx_any)
table(data.v9$dx_dep)
table(data.v9$dx_anx)
table(data.v9$dx_bi)
table(data.v9$dx_psy)
table(data.v9$dx_ea)
table(data.v9$dx_tr)
table(data.v9$dx_sa)
table(data.v9$dx_adhd)
table(data.v9$dx_pers)
table(data.v9$dx_none)
```

## Religion
```{r religios}
# Label levels of religious importance
data.v9 <- data.v9 %>%
  filter(!is.na(religios)) %>% 
  mutate(religios = factor(religios, levels=1:5,
                    labels=c("Very Important", "Important", "Neutral", "Unimportant", "Very unimportant")))
table(data.v9$religios, useNA = "always")       # Check output
```

## AUDIT  
** too small for inclusion
```{r audit_tot}
data.v9 <- data.v9 %>%
  # Combine audit_3_f, audit_3_m, audit_3_o to create 'audit_3'
  mutate(audit_3 = case_when(audit_3_f == 1 ~ 1,
                             audit_3_m == 1 ~ 1,
                             audit_3_o == 1 ~ 1,
                             audit_3_f == 0 ~ 0,
                             audit_3_m == 0 ~ 0,
                             audit_3_o == 0 ~ 0)) %>%
  # All outputs needs to be numeric for future addition of AUDIT total
  mutate(audit_1 = as.numeric(audit_1),
         audit_2 = as.numeric(audit_2),
         audit_3 = as.numeric(audit_3),
         audit_4 = as.numeric(audit_4),
         audit_5 = as.numeric(audit_5),
         audit_6 = as.numeric(audit_5),
         audit_7 = as.numeric(audit_7),
         audit_8 = as.numeric(audit_8),
         audit_9 = as.numeric(audit_9),
         audit_10 = as.numeric(audit_10)) %>%
  # Find audit_tot
  mutate(audit_tot = as.numeric(audit_1) + as.numeric(audit_2) + as.numeric(audit_3) + 
           as.numeric(audit_4) + as.numeric(audit_5) + as.numeric(audit_6) + as.numeric(audit_7) + 
           as.numeric(audit_8) + as.numeric(audit_9) + as.numeric(audit_10))
table(data.v9$audit_tot)

```

## ETOH treatment
n= 3885 ** sample is too small
```{r alc_tx}
table(data.v9$alc_tx)
```

## Eating disorder, any type
```{r ed_any}
table(data.v9$ed_any)
```

## WCS
Combine wcs_1_f, wcs_1_m, wcs_1_o to create 'wcs_1'
** too small
```{r WCS data merge}

table(data.v9$wcs_1_f)
table(data.v9$wcs_1_m)
table(data.v9$wcs_1_o)

table(data.v9$wcs_2)
table(data.v9$wcs_3)
table(data.v9$wcs_4)
table(data.v9$wcs_5)

data.v9 = subset(data.v9, select= -c(wcs_1_f, wcs_1_m, wcs_1_o, wcs_2, 
                                     wcs_3, wcs_4, wcs_5)) # clean up vars/columns

```

## Medication use
```{r meds}
# Create stimulant
data.v11 <- data.v9 %>%
  mutate(meds_sti = case_when(meds_1 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_1 == 0 ~ 0))
      # Reconcile stimulant write-ins from meds_7_text
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[V|v]yvan[s|c]e")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[T|t]enex")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[C|c]oncerta")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[A|a]dderall")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[R|r]italin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[P|p]rovigil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[A|a]rmodafinil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[F|f]ocalin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[M|m]etadat")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[F|f]ocalin")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[N|n]uvigil")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "ADHD")] <- 1
      data.v11$meds_sti[str_detect(data.v11$meds_7_text, "[M|m]odafinil")] <- 1
      data.v11$meds_sti[data.v11$meds_7_text %in% c("Vyance", "Vivance adhd", "Stratera", "Aderol", "Aderal", "Adderal",
                                                    "Adderrall (for ADHD)", "Adderal, vyvance",  "Celltech")] <- 1
      
# Create antidepressant
data.v11 <- data.v11 %>%
  mutate(meds_dep = case_when(meds_2 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_2 == 0 ~ 0))
      # reconcile write-ins for depression from med_7_text
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]elexa")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]italopram")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]ymbalta")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[E|e]scitalopram")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[L|l]exapro")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[S|s]ertraline")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[Z|z]oloft")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "SSRI")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[E|e]lavil")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]aroxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]rozac")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]ntidepressant")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]mitriptyline")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]uloxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c][e|i]pralex")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[W|w]ellbutrin")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[S|s]trattera")] <- 1 
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[A|a]tomoxetine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[F|f]luvoxamine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]enlafaxine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[P|p]ristiq")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]esipramine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]iibryd")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[V|v]enlafaxin")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[C|c]lomipramine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[D|d]esvenflaxine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[M|m]irtazapine")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[R|r][e|i]meron")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[T|t]raz[a|o|i]done")] <- 1
      data.v11$meds_dep[str_detect(data.v11$meds_7_text, "[N|n]ortriptyline")] <- 1
      data.v11$meds_dep[data.v11$meds_7_text %in% c("CYMBALTA", "Cymbolta", "Lexipro", "Lexxapro", "Sertaline", "remiron",
                                                    "Mirtazepene", "bupropin", "Lexapro, Wellbutrin", "Lexapro and Paxil", "Fuoxetine", 
                                                    "Forgot the name, anti depressant for anxiety, daily med", "Duluxotine", 
                                                    "Atomoxitine (nonstimulant ADHD)", "Fluvoximine (luvox)", "Anafranil", 
                                                    "amatryptaline for migraines", "amnitryptaline", 
                                                    "Buproprion (may or may not count as a psychostim?)", "amatryptaline for migraines",
                                                    "Minipress")] <- 1
      
# Create antipsychotic
data.v11 <- data.v11 %>%
  mutate(meds_psy = case_when(meds_3 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_3 == 0 ~ 0))
      # Reconcile anti-psychotic entries from med_7_text
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]bilify")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]bility")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[V|v]raylar")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[O|o]lanzapine")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[A|a]ripiprazole")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[L|l]atuda")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[S|s]eroqu[e|o]l")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[G|g]eodon")] <- 1
      data.v11$meds_psy[str_detect(data.v11$meds_7_text, "[Q|q]uetiapine")] <- 1
      data.v11$meds_psy[data.v11$meds_7_text %in% c("Ziprazidone", "Quitiapin", "Latuda, Giadon")] <- 1 
      
# Create anti-anxiety  
data.v11 <- data.v11 %>%            
  mutate(meds_anx = case_when(meds_4 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_4 == 0 ~ 0))
      # Reconcile all anti-anxiety drugs from the write-ins from meds_7_text
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[X|x]anax")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[V|v]alium")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]ropranolol")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]ropanol")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[L|l]orazepam")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[N|n]eurontin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[A|a]tarax")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[L|l]yrica")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[A|a]tivan")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[C|c]lonidine")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[P|p]razosin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[B|b]uspar")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[V|v]istaril")] <- 1  # antihistamine
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[I|i]nderal")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[H|h]ydroxyzine")] <- 1 # antihistamine
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[K|k]lonopin")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[D|d]iazepam")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[B|b]uspirone")] <- 1
      data.v11$meds_anx[str_detect(data.v11$meds_7_text, "[G|g]abapentin")] <- 1
      data.v11$meds_anx[data.v11$meds_7_text %in% c("Valuim", "you didnt list vallium under anti-anxiety", "Alphrazolam", 
                                                   "Hydroxizine--prescribed as needed for anxiety, only took one because of adverse reaction", 
                                                   "Hydrodoxine, a blood pressure medicine for anxiety that I don't remember the name of")] <- 1

# Create mood stabilizer
data.v11 <- data.v11 %>%
  mutate(meds_mood = case_when(meds_5 == 1 ~ 1,
                               meds_8 == 1 ~ 0,
                               meds_5 == 0 ~ 0))
      # Reconcile mood stabilizers from write-in in med_7_text
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]ithium")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]amictal")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[L|l]amotrigine")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[D|d]epakote")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[T|t]rileptal")] <- 1
      data.v11$meds_mood[str_detect(data.v11$meds_7_text, "[O|o]xcarbarzepine")] <- 1
      data.v11$meds_mood[data.v11$meds_7_text %in% c("Mood Stabilizer/Birth Control")] <- 1
      
# Create sleep med
data.v11 <- data.v11 %>%
  mutate(meds_sle = case_when(meds_6 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_6 == 0 ~ 0))
      # Reconcile sleep medications from write-ins in meds_7_text
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[X|x]yrem")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[S|s]leep")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[M|m]elatonin")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[Z|z]quil")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[N|n]yquil")] <- 1
      data.v11$meds_sle[str_detect(data.v11$meds_7_text, "[D|d]iphenhydramine")] <- 1
      data.v11$meds_sle[data.v11$meds_7_text %in% c("gabapentin, trazodone (both for sleep on separate occasions)", 
                                                    "generic sleep medication: melatonin supplements", 
                                                    "cold medicine to induce sleep")] <- 1
      
# Create "Other" meds
data.v11 <- data.v11 %>%
  mutate(meds_oth = case_when(meds_7 == 1 ~ 1,
                              meds_8 == 1 ~ 0,
                              meds_7 == 0 ~ 0)) 
      # Reconcile "other" entries from med_7_text
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]annabis")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[M|m]arijuana")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[W|w]eed")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[N|n]altrexone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[S|s]ynthroid")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]opomax")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]estosterone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[T|t]hyroid")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[H|h]ydrodocone")] <- 1 
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]lonidine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[G|g]uanfacine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[K|k]etamine")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[F|f]lexeril")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[P|p]ant[o|a]loc")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[S|s]pironolactone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[H|h]ormone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[O|o]xycodone")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[C|c]annabic")] <- 1
      data.v11$meds_oth[str_detect(data.v11$meds_7_text, "[B|b]eta")] <- 1
      data.v11$meds_oth[data.v11$meds_7_text %in% c("Canabis", "Vistiril", "Vistral", "Thyroxine", "The Reefer", "Topimitrate",
                                                    "topiramate", "Estradiol, Spironactone", "muscle relaxants to help me sleep",
                                                    "Sumatriptan for migraines", "Ondansetron" ,"Psilocybin Mushrooms", "clonodine",
                                                    "Muscle Relaxers", "anti-siezure medication", "Codeine", "Androgen Blockers",
                                                    "Anti-seizure to control racing thoughts", "Anti seizure",
                                                    "Buprenorphine (Suboxone) for opiate dependence", "Hydroxine", "Hydroxcyzine",
                                                    "Alyacen", "Mirapex", "hypothyroid medication", "Epilepsy drugs", "Birth control",
                                                    "Methadone", "Vimpat", "Contrave", "Benztropine", "Intuniv", "Medicinal marijuana",
                                                    "Pantaloc, for psychogenic vomiting", "medicinal Cannabis",
                                                    "Testosterone Enanthate", "testosterone", "Cannabis", "Medical marijuana")] <- 1
      
# Create a group for students who do not take any psychotropic meds
data.v11 <- data.v11 %>%
  mutate(meds_none = case_when(meds_none == 1 ~ 1,
                               meds_none == 0 ~ 0,
                               meds_sti == 1 ~ 0,
                               meds_dep == 1 ~ 0,
                               meds_psy == 1 ~ 0,
                               meds_anx == 1 ~ 0,
                               meds_mood == 1 ~ 0,
                               meds_sle == 1 ~ 0,
                               meds_oth == 1 ~ 0))

# tabulate and check outputs
table(data.v11$meds_sti)
table(data.v11$meds_dep)
table(data.v11$meds_psy)
table(data.v11$meds_anx)
table(data.v11$meds_mood)
table(data.v11$meds_sle)
table(data.v11$meds_oth)
table(data.v11$meds_none)

```

## Psychopharm medication count
Create 'meds_count' to be used in future regression to capture quantity and diversity of medications 
to see what is associated (i.e. the count is meant to find the common meds in suicidality)
```{r med_count}
data.v11$meds_1 <- as.numeric(data.v11$meds_1)
data.v11$meds_2 <- as.numeric(data.v11$meds_2)
data.v11$meds_3 <- as.numeric(data.v11$meds_3)
data.v11$meds_4 <- as.numeric(data.v11$meds_4)
data.v11$meds_5 <- as.numeric(data.v11$meds_5)
data.v11$meds_6 <- as.numeric(data.v11$meds_6)
data.v11$meds_7 <- as.numeric(data.v11$meds_7)
data.v11$meds_8 <- as.numeric(data.v11$meds_8)

data.v11$meds_1[is.na(data.v11$meds_1)] <- 0
data.v11$meds_2[is.na(data.v11$meds_2)] <- 0
data.v11$meds_3[is.na(data.v11$meds_3)] <- 0
data.v11$meds_4[is.na(data.v11$meds_4)] <- 0
data.v11$meds_5[is.na(data.v11$meds_5)] <- 0
data.v11$meds_6[is.na(data.v11$meds_6)] <- 0
data.v11$meds_7[is.na(data.v11$meds_7)] <- 0

# create meds_count var
data.v11 <- data.v11 %>%
  mutate(meds_count = meds_1 + meds_2 + meds_3 + meds_4 + meds_5 + meds_6 + meds_7)

table(data.v11$meds_count)  # check output
data.v11 = subset(data.v11, select= -c(meds_1:meds_7)) # clean up vars/columns
```

## Diathesis correlation matrix
```{r Diathesis correlation}

# Define diathesis vars for correlation matrix
data_dia <- data.v11 %>% select(age, sexual, gender, finpast, sib_freq, race, disab_1, ed_scoff, body_sr,
                                dx_dep, dx_any, dx_bi, dx_adhd, dx_psy, dx_ea, dx_tr, dx_sa, meds_count,
                                dx_pers, dx_none, meds_sti, meds_dep, meds_psy, meds_anx, meds_mood, meds_sle,
                                meds_oth, meds_none, ed_any)

# Corr_simple function
corr_simple <- function(data=data_dia, sig=0.6){
  # convert data to numeric in order to run correlations 
  # (i.e., convert to factor first to keep the integrity of the data - each value will become a num rather than NA)
  data_dia <- data_dia %>% mutate_if(is.character, as.factor)
  data_dia <-  data_dia %>% mutate_if(is.factor, as.numeric)
  # run a correlation and drop the insignificant ones
  corr <- cor(data_dia, use = "pairwise.complete.obs")                               
  corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
  corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
  corr <- as.data.frame(as.table(corr))                   # turn into a 3-column table
  corr <- na.omit(corr)                                   # remove the NA values from above 
  corr <- corr[order(-abs(corr$Freq)),]                   # sort by highest correlation
  corr <- subset(corr, abs(Freq) > sig)                   # select significant values 
  print(corr)
  
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}
corr_simple()

```
## EFA diathesis
```{r EFA diathesis}

data_dia <- data.v11 %>% select(age, sexual, gender, finpast, sib_freq, race, disab_1, ed_scoff, body_sr,
                                dx_dep, dx_any, dx_bi, dx_adhd, dx_psy, dx_ea, dx_tr, dx_sa, ed_any,
                                dx_pers, dx_none, meds_sti, meds_dep, meds_psy, meds_anx, meds_mood, meds_sle,
                                meds_oth, meds_none, meds_count)

# create corr matrix, with dummy vars
data_dia <- data_dia %>% 
  mutate_if(is.numeric, as.factor) %>% 
  mutate(
    age = as.numeric(age),
    meds_count = as.numeric(meds_count),
    ed_scoff = as.numeric(ed_scoff)
  ) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_dia   # check output

results_dia <- dummy_cols(data_dia, select_columns = c("age", "sexual", "gender", "finpast", "sib_freq", "race", "ed_any",
                                                       "disab_1", "ed_scoff", "body_sr", "audit_tot", "dx_dep", "dx_any",
                                                       "dx_bi", "dx_adhd", "dx_psy", "dx_ea", "dx_tr", "dx_sa", "dx_pers",
                                                       "dx_none", "meds_sti", "meds_dep", "meds_psy", "meds_anx", 
                                                       "meds_mood", "meds_sle", "meds_oth", "meds_none", "alc_tx", "meds_count"), 
                          remove_first_dummy = TRUE)

dia_pca <- PCA(results_dia)  # apply PCA
dia_pca$eig   # output matrix with eigenvalues

```

## CLM diathesis regression
Regression variables: age, sexual, gender, finpast, sib_freq, race, ed_scoff, body_sr, disab_1
dx_dep, dx_any, dx_bi, dx_psy, dx_ea, dx_tr, dx_sa, dx_adhd, dx_pers, dx_none, meds_sti, meds_dep, 
meds_psy, meds_anx, meds_mood, meds_sle, meds_other, meds_none,  meds_count 
```{r clm_dia}

clm_dia <- data.v11 %>% select(si_type, age, sexual, gender, finpast, sib_freq, race, disab_1, ed_scoff, 
                               body_sr, dx_dep, dx_any, dx_bi, dx_psy, dx_adhd, meds_count, ed_any,
                               dx_ea, dx_tr, dx_sa, dx_pers, dx_none, meds_sti, meds_dep, meds_psy, meds_anx, 
                               meds_mood, meds_sle, meds_oth, meds_none)

clm_dia <- NA_preproc(clm_dia)
clm_dia <- as.data.frame(clm_dia)

str(clm_dia)
set.seed(1000) 

index = sample(1:nrow(clm_dia), 0.7*nrow(clm_dia)) 

train_dia = clm_dia[index,]       # Create the training data for diathesis
test_dia = clm_dia[-index,]       # Create the test data for diathesis

dim(train_dia)
dim(test_dia)

m_dia <- clm(si_type ~ age + sexual + gender + finpast + sib_freq + race + disab_1 + ed_scoff + 
               meds_count + ed_any + dx_ea + dx_tr + dx_sa + dx_pers + meds_sti + meds_dep + meds_psy +
               body_sr + dx_dep + dx_bi + dx_psy + dx_adhd, data = clm_dia)
summary(m_dia)

# # sure fx
# dia_data <- m_dia[["model"]]
# 
# # residual plot
# dia_data$resid <- resids(m_dia, method = c("jitter"),
#       jitter.scale = c("response"))
# dia_data <- cbind(dia_data, m_dia[["fitted.values"]])
# ggplot(dia_data, aes(x = `active SI`, y = resid)) +
# geom_point(color = "purple", alpha = 0.15)
# geom_smooth(method = "lm")
# 
# dia_data %>% group_by(si_type_num) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))
# 
# dia_data %>% 
#   filter(si_type == "active SI") %>%
#   ggplot(aes(x = resid)) + geom_density(color = "magenta")

```


# ENVIRONMENT VARIABLES
Candidate variables: res_status, military, activ_yn, residenc, discrim, env_mh, ther_vis, school2, medtherc, medther, talk, 
religios, residenc, satisfied_overall, firstgen


```{r Env frame}
data.v12 <- data.v11
```

## US resident status
n = 942 ** too small to include
```{r U.S. Resident Status}
# aggregate all residency variables into 1 variable called 'res_status'
data.v12 <- data.v12 %>% 
  mutate(res_status = case_when(citizen == 1 ~ 1,
                                permanentres == 1 ~ 2, 
                                visa == 1 ~ 3,
                                undoc == 1 ~ 4,
                                otherdoc == 1 ~ 5))
table(data.v12$res_status)  # check output
data.v12 = subset(data.v12, select= -c(citizen, permanentres, visa, otherdoc, undoc))   # clean up var
```

## Discrimination
```{r discrim}
table(data.v12$discrim)
```

## MH campus environment
```{r env_mh}
table(data.v12$env_mh) 
```

## Total therapy visits
```{r ther_vis}
table(data.v11$ther_ever)
table(data.v11$ther_vis)

# Ther_ever is the parent variable to ther_vis. Combine for full sample.
data.v12 <- data.v11 %>%
  mutate(ther_vis = case_when(ther_ever == 0 ~ 0,
                              ther_vis == 0 ~ 0,
                              ther_vis == 1 ~ 1,
                              ther_vis == 2 ~ 2,
                              ther_vis == 3 ~ 3,
                              ther_vis == 4 ~ 4))
table(data.v12$ther_vis)   # check output
```

## Distress & social support
Group talk var into subcategories of none, informal, formal, or a combo of support
```{r talk}
data.v12$talk1_1[is.na(data.v12$talk1_1)] <- 0
data.v12$talk1_2[is.na(data.v12$talk1_2)] <- 0
data.v12$talk1_3[is.na(data.v12$talk1_3)] <- 0
data.v12$talk1_4[is.na(data.v12$talk1_4)] <- 0
data.v12$talk1_5[is.na(data.v12$talk1_5)] <- 0
data.v12$talk1_6[is.na(data.v12$talk1_6)] <- 0
data.v12$talk1_7[is.na(data.v12$talk1_7)] <- 0
data.v12$talk1_8[is.na(data.v12$talk1_8)] <- 0

# change talk_ var to numeric
data.v13 <- data.v12 %>% 
  # aggregate all talk vars under 3 'talk' categories -- none, informal help, formal help, formal+informal help
  mutate(talk = case_when(((talk1_1 == 1 | talk1_6 == 1 | talk1_7 == 1 | talk1_8 == 1) &                      # combined
                             (talk1_2 == 1 | talk1_3 == 1 | talk1_4 == 1 | talk1_5 == 1)) ~ 4,
                          (talk1_9 == 1 | talk1no == 1) ~ 1,                                                  # no support
                          (talk1_1 == 1 | talk1_6 == 1 | talk1_7 == 1 | talk1_8 == 1) ~ 2,                    # formal support
                          (talk1_2 == 1 | talk1_3 == 1 | talk1_4 == 1 | talk1_5 == 1) ~ 3)) %>%               # informal support
    mutate(talk = factor(talk, levels=1:4,
                               labels=c("No support", "Formal", "Informal", "Combined")))
table(data.v13$talk)  # check output
data.v13 = subset(data.v13, select= -c(talk1_1:talk1_9, talk1sig, talk1fr, talk1rel, talk1no, talk1sup, talk1rm, talk1pro, talk1fam)) # clean up
```

## Military status
```{r military}
table(data.v13$military)
```

## International student
```{r international}
table(data.v13$international)
```

## Therapy + Medication Treatment
```{r medtherc}
table(data.v13$medther)  # med or therapy
table(data.v13$medtherc) # meds + therapy
```

## Family support
```{r fam_support_aca}
table(data.v13$fam_support_aca)
```

## Overall school satisfaction
```{r satisfied_overall}
table(data.v13$satisfied_overall)
```

## Activity Involvement
```{r Activities}

  # set all NAs in acitvities equal to 0
data.v13$activ_ac[is.na(data.v13$activ_ac)] <- 0
data.v13$activ_athc[is.na(data.v13$activ_athc)] <- 0
data.v13$activ_athv[is.na(data.v13$activ_athv)] <- 0
data.v13$activ_athi[is.na(data.v13$activ_athi)] <- 0
data.v13$activ_cs[is.na(data.v13$activ_cs)] <- 0
data.v13$activ_cu[is.na(data.v13$activ_cu)] <- 0
data.v13$activ_da[is.na(data.v13$activ_da)] <- 0
data.v13$activ_fs[is.na(data.v13$activ_fs)] <- 0
data.v13$activ_gs[is.na(data.v13$activ_gs)] <- 0
data.v13$activ_gov[is.na(data.v13$activ_gov)] <- 0
data.v13$activ_mp[is.na(data.v13$activ_md)] <- 0
data.v13$activ_md[is.na(data.v13$activ_none)] <- 0
data.v13$activ_soc[is.na(data.v13$activ_soc)] <- 0
data.v13$activ_rel[is.na(data.v13$activ_rel)] <- 0
data.v13$activ_art[is.na(data.v13$activ_art)] <- 0
data.v13$activ_other[is.na(data.v13$activ_other)] <- 0

  # change class to numeric
data.v13$activ_ac <- as.numeric(data.v13$activ_ac)
data.v13$activ_athc <- as.numeric(data.v13$activ_athc)
data.v13$activ_athv <- as.numeric(data.v13$activ_athv)
data.v13$activ_athi <- as.numeric(data.v13$activ_athi)
data.v13$activ_cs <- as.numeric(data.v13$activ_cs)
data.v13$activ_cu <- as.numeric(data.v13$activ_cu)
data.v13$activ_da <- as.numeric(data.v13$activ_da)
data.v13$activ_fs <- as.numeric(data.v13$activ_fs)
data.v13$activ_gs <- as.numeric(data.v13$activ_gs)
data.v13$activ_gov <- as.numeric(data.v13$activ_gov)
data.v13$activ_mp <- as.numeric(data.v13$activ_md)
data.v13$activ_md <- as.numeric(data.v13$activ_none)
data.v13$activ_soc <- as.numeric(data.v13$activ_soc)
data.v13$activ_rel <- as.numeric(data.v13$activ_rel)
data.v13$activ_art <- as.numeric(data.v13$activ_art)
data.v13$activ_other <- as.numeric(data.v13$activ_other)

# Create one activity variable (dichotomous) based on involment in campus activities (org, sports, etc) or not involved
data.v14 <- data.v13 %>% 
  mutate(activ_yn = case_when(activ_none == 1 ~ 0,
                              (activ_ac == 1 | activ_athc == 1 | activ_athv == 1 | activ_athi == 1 | activ_cs == 1 | activ_cu |
                              activ_da == 1 | activ_fs == 1 | activ_gs == 1 | activ_gov == 1 | activ_mp == 1 | activ_md == 1 |
                              activ_rel == 1 | activ_soc == 1 | activ_art == 1 | activ_other == 1) ~ 1))

table(data.v14$activ_yn)     # tabulate and check output
data.v14 = subset(data.v14, select= -c(activ_none, activ_ac:activ_other))    # clean up vars
```

## School variable
```{r school2}

#insert each school as a dichotomous var
data.v14 <- data.v14 %>%
  mutate(var_AlbertaCAD = ifelse((school2 == "AlbertaCAD"), 1, 0)) %>%
  mutate(var_Boston = ifelse((school2 == "Boston"), 1, 0)) %>%
  mutate(var_CIA = ifelse((school2 == "CIA"), 1, 0)) %>%
  mutate(var_CalPoly = ifelse((school2 == "CalPoly"), 1, 0)) %>%
  mutate(var_ColoradoCollege = ifelse((school2 == "ColoradoCollege"), 1, 0)) %>%
  mutate(var_ColoradoMountain = ifelse((school2 == "ColoradoMountain"), 1, 0)) %>%
  mutate(var_ColumbusCAD = ifelse((school2 == "ColumbusCAD"), 1, 0)) %>%
  mutate(var_Delaware = ifelse((school2 == "Delaware"), 1, 0)) %>%
  mutate(var_DetroitMercy = ifelse((school2 == "DetroitMercy"), 1, 0)) %>%
  mutate(var_Emerson = ifelse((school2 == "Emerson"), 1, 0)) %>%
  mutate(var_GeorgeMason = ifelse((school2 == "GeorgeMason"), 1, 0)) %>%
  mutate(var_GeorgiaTech = ifelse((school2 == "GeorgiaTech"), 1, 0)) %>%
  mutate(var_Kalamazoo = ifelse((school2 == "Kalamazoo"), 1, 0)) %>%
  mutate(var_Kansas = ifelse((school2 == "Kansas"), 1, 0)) %>%
  mutate(var_MDInstituteCollegeArt = ifelse((school2 == "MDInstituteCollegeArt"), 1, 0)) %>%
  mutate(var_MassArt = ifelse((school2 == "MassArt"), 1, 0)) %>%
  mutate(var_Merrimack = ifelse((school2 == "Merrimack"), 1, 0)) %>%
  mutate(var_MichiganState = ifelse((school2 == "MichiganState"), 1, 0)) %>%
  mutate(var_MinnesotaCAD = ifelse((school2 == "MinnesotaCAD"), 1, 0)) %>%
  mutate(var_MontclairState = ifelse((school2 == "MontclairState"), 1, 0)) %>%
  mutate(var_NHInstituteArt = ifelse((school2 == "NHInstituteArt"), 1, 0)) %>%
  mutate(var_NevadaReno = ifelse((school2 == "NevadaReno"), 1, 0)) %>%
  mutate(var_OaklandCC = ifelse((school2 == "OaklandCC"), 1, 0)) %>%
  mutate(var_OklahomaCityCC = ifelse((school2 == "OklahomaCityCC"), 1, 0)) %>%
  mutate(var_PaloAltoCollege = ifelse((school2 == "PaloAltoCollege"), 1, 0)) %>%
  mutate(var_PennCollegeTech = ifelse((school2 == "PennCollegeTech"), 1, 0)) %>%
  mutate(var_Pratt = ifelse((school2 == "RedlandsCC"), 1, 0)) %>%
  mutate(var_RhodeIsland = ifelse((school2 == "RhodeIsland"), 1, 0)) %>%
  mutate(var_RedlandsCC = ifelse((school2 == "RedlandsCC"), 1, 0)) %>%
  mutate(var_Ringling = ifelse((school2 == "Ringling"), 1, 0)) %>%
  mutate(var_Rollins = ifelse((school2 == "Rollins"), 1, 0)) %>%
  mutate(var_SAIC = ifelse((school2 == "SAIC"), 1, 0)) %>%
  mutate(var_Sewanee = ifelse((school2 == "Sewanee"), 1, 0)) %>%
  mutate(var_Smith = ifelse((school2 == "Smith"), 1, 0)) %>%
  mutate(var_SouthernNazarene = ifelse((school2 == "SouthernNazarene"), 1, 0)) %>%
  mutate(var_SouthwesternOklahomaState = ifelse((school2 == "SouthwesternOklahomaState"), 1, 0)) %>%
  mutate(var_StJohns = ifelse((school2 == "StJohns"), 1, 0)) %>%
  mutate(var_StMarys = ifelse((school2 == "StMarys"), 1, 0)) %>%
  mutate(var_StRose = ifelse((school2 == "StRose"), 1, 0)) %>%
  mutate(var_TrumanState = ifelse((school2 == "TrumanState"), 1, 0)) %>%
  mutate(var_TuftsHealthSciences = ifelse((school2 == "TuftsHealthSciences"), 1, 0)) %>%
  mutate(var_TuftsMainCampus = ifelse((school2 == "TuftsMainCampus"), 1, 0)) %>%
  mutate(var_UChicago = ifelse((school2 == "UChicago"), 1, 0)) %>%
  mutate(var_UFlorida = ifelse((school2 == "UFlorida"), 1, 0)) %>%
  mutate(var_UMDearborn = ifelse((school2 == "UMDearborn"), 1, 0)) %>%
  mutate(var_UMich = ifelse((school2 == "UMich"), 1, 0)) %>%
  mutate(var_UNCSchoolArt = ifelse((school2 == "UNCSchoolArt"), 1, 0)) %>%
  mutate(var_UTKnoxville = ifelse((school2 == "UTKnoxville"), 1, 0)) %>%
  mutate(var_UniversityofSouthernCalifornia = ifelse((school2 == "UniversityofSouthernCalifornia"), 1, 0)) %>%  
  mutate(var_Utah = ifelse((school2 == "Utah"), 1, 0)) %>%
  mutate(var_VATech = ifelse((school2 == "VATech"), 1, 0)) %>%
  mutate(var_VirginiaCommonwealth = ifelse((school2 == "VirginiaCommonwealth"), 1, 0)) %>%
  mutate(var_WakeForest = ifelse((school2 == "WakeForest"), 1, 0)) %>%
  mutate(var_Watkins = ifelse((school2 == "Watkins"), 1, 0)) %>%
  mutate(var_WestVirginia = ifelse((school2 == "WestVirginia"), 1, 0)) %>%
  mutate(var_WesternCarolina = ifelse((school2 == "WesternCarolina"), 1, 0)) %>%
  mutate(var_WesternMichigan = ifelse((school2 == "WesternMichigan"), 1, 0)) %>%
  mutate(var_WesternWA = ifelse((school2 == "WesternWA"), 1, 0)) %>%
  mutate(var_Xavier = ifelse((school2 == "Xavier"), 1, 0))

```

### School type levels

(1) 4-year: var_Boston, var_CalPoly, var_ColoradoCollege, var_ColoradoMountain, var_Delaware, var_DetroitMercy, var_Emerson, var_GeorgeMason, var_GeorgiaTech, var_Kalamazoo, var_Kansas, var_Merrimack, var_MichiganState, var_MontclairState, var_NevadaReno, var_PennCollegeTech, var_RhodeIsland, var_Rollins, var_Sewanee, var_Smith, var_SouthernNazarene, var_SouthwesternOklahomaState, var_StJohns, var_StMarys, var_StRose, var_TrumanState, var_TuftsHealthSciences, var_TuftsMainCampus, var_UChicago, var_UFlorida, var_UMDearborn, var_UMich, var_UTKnoxville,  var_UniversityofSouthernCalifornia, var_Utah, var_VATech, var_VirginiaCommonwealth, var_WakeForest,var_WestVirginia, var_WesternCarolina, var_WesternMichigan, var_WesternWA, var_Xavier

(2) Community College: var_OaklandCC, var_OklahomaCityCC, var_PaloAltoCollege, var_RedlandsCC 

(3) Art & Design, Culinary: var_AlbertaCAD, var_CIA, var_ColumbusCAD, var_MDInstituteCollegeArt, var_MassArt, var_MinnesotaCAD, var_NHInstituteArt, var_Pratt, , var_Ringling, var_SAIC, var_UNCSchoolArt, var_Watkins, 

```{r school2_type}
data.v14 <- data.v14 %>%
  # 4 year college
  mutate(school2_4yr = ifelse(var_Boston + var_CalPoly + var_ColoradoCollege + var_ColoradoMountain + var_Delaware +
                                var_DetroitMercy + var_Emerson + var_GeorgeMason + var_GeorgiaTech + var_Kalamazoo + 
                                var_Kansas + var_Merrimack + var_MichiganState + var_MontclairState + var_NevadaReno +
                                var_PennCollegeTech + var_RhodeIsland + var_Rollins + var_Sewanee + var_Smith +
                                var_SouthernNazarene + var_SouthwesternOklahomaState + var_StJohns + var_StMarys + var_StRose +
                                var_TrumanState + var_TuftsHealthSciences + var_TuftsMainCampus + var_UChicago + var_UFlorida +
                                var_UMDearborn + var_UMich + var_UTKnoxville + var_UniversityofSouthernCalifornia + var_Utah +
                                var_VATech + var_VirginiaCommonwealth + var_WakeForest + var_WestVirginia + 
                                var_WesternCarolina + var_WesternMichigan + var_WesternWA + var_Xavier == 1, TRUE,
                              ifelse(
                                var_Boston + var_CalPoly + var_ColoradoCollege + var_ColoradoMountain + var_Delaware +
                                  var_DetroitMercy + var_Emerson + var_GeorgeMason + var_GeorgiaTech + var_Kalamazoo +
                                  var_Kansas + var_Merrimack + var_MichiganState + var_MontclairState + var_NevadaReno +
                                  var_PennCollegeTech + var_RhodeIsland + var_Rollins + var_Sewanee + var_Smith +
                                  var_SouthernNazarene + var_SouthwesternOklahomaState + var_StJohns + var_StMarys + 
                                  var_StRose + var_TrumanState + var_TuftsHealthSciences + var_TuftsMainCampus + var_UChicago +
                                  var_UFlorida + var_UMDearborn + var_UMich + var_UTKnoxville +
                                  var_UniversityofSouthernCalifornia + var_Utah + var_VATech + var_VirginiaCommonwealth +
                                  var_WakeForest + var_WestVirginia + var_WesternCarolina + var_WesternMichigan +
                                  var_WesternWA + var_Xavier == 0, FALSE, NA))) %>%
    
  # Community colleges
    mutate(school2_cc = ifelse(var_OaklandCC + var_OklahomaCityCC + var_PaloAltoCollege + var_RedlandsCC == 1, TRUE, 
                               ifelse(var_OaklandCC + var_OklahomaCityCC + var_PaloAltoCollege + var_RedlandsCC == 0, FALSE, NA))) %>%
    
  # Art, design, culinary
    mutate(school2_ad = ifelse(var_AlbertaCAD + var_CIA + var_ColumbusCAD + var_MDInstituteCollegeArt + var_MassArt +
                                 var_MinnesotaCAD + var_NHInstituteArt + var_Pratt + var_Ringling + var_SAIC + 
                                 var_UNCSchoolArt + var_Watkins == 1, TRUE, 
                               ifelse(
                                 var_AlbertaCAD + var_CIA + var_ColumbusCAD + var_MDInstituteCollegeArt + var_MassArt +
                                   var_MinnesotaCAD + var_NHInstituteArt + var_Pratt + var_Ringling + var_SAIC +
                                   var_UNCSchoolArt + var_Watkins == 0, FALSE, NA)))

# Binary variable for school type
table(data.v14$school2_4yr)
table(data.v14$school2_cc)
table(data.v14$school2_ad)

#  create one school type variables
data.v14 <- data.v14 %>%
  mutate(school2_type = case_when(school2_4yr == TRUE ~ 1,
                                  school2_cc == TRUE ~ 2,
                                  school2_ad == TRUE ~ 3)) %>%
  mutate(school2_type = factor(school2_type, levels=1:3,
                               labels=c("4-yr colleges", "commununity colleges", "art & design schools")))
table(data.v14$school2_type)
      

# Clean-up school2 var
data.v14 = subset(data.v14, select= -c(var_Boston, var_CalPoly, var_ColoradoCollege, var_ColoradoMountain, var_Delaware,
                                       var_DetroitMercy, var_Emerson, var_GeorgeMason, var_GeorgiaTech, var_Kalamazoo,
                                       var_Kansas, var_Merrimack, var_MichiganState, var_MontclairState, var_NevadaReno,
                                       var_PennCollegeTech, var_RhodeIsland, var_Rollins, var_Sewanee, var_Smith,
                                       var_SouthernNazarene, var_SouthwesternOklahomaState, var_StJohns, var_StMarys,
                                       var_StRose, var_TrumanState, var_TuftsHealthSciences, var_TuftsMainCampus, var_UChicago,
                                       var_UFlorida, var_UMDearborn, var_UMich, var_UTKnoxville, 
                                       var_UniversityofSouthernCalifornia, var_Utah, var_VATech, var_VirginiaCommonwealth,
                                       var_WakeForest, var_WestVirginia, var_WesternCarolina, var_WesternMichigan,
                                       var_WesternWA, var_Xavier, var_OaklandCC, var_OklahomaCityCC, var_PaloAltoCollege,
                                       var_RedlandsCC, var_AlbertaCAD, var_CIA, var_ColumbusCAD, var_MDInstituteCollegeArt,
                                       var_MassArt, var_MinnesotaCAD, var_NHInstituteArt, var_Pratt, var_Ringling, var_SAIC,
                                       var_UNCSchoolArt, var_Watkins))


```

## Environmental correlation
```{r}
# Define environment vars for correlation matrix
data_env <- data.v14 %>% select(discrim, religios, env_mh, residenc, ther_vis, activ_yn, divers, 
                                  school2_type, satisfied_overall, ther_vis, 
                                  talk, military, international)
data <- data_env
sig <- 0.6

 # Convert data to numeric in order to run correlations (i.e., convert to factor first to keep the integrity of the data 
 # Each value will become a number rather than turn into NA_
data_env <- data_env %>% mutate_if(is.character, as.factor)
data_env <-  data_env %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_env, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
# turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")

```

##EFA Environment
```{r}
data_env <- data.v14 %>% select(discrim, religios, env_mh, residenc, ther_vis, activ_yn, divers, 
                                school2_type, satisfied_overall, ther_vis, talk, military, international)

# create corr matrix, with dummy vars
data_env <- data_env %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_env   # check output

# impute to perform FAMD calculation
# require(missMDA)
# data(data_env)
# res.impute <- imputeFAMD(data_env, ncp=3)
# res.famd <- FAMD(res.pca, tab.disj=res.impute$tab.disj)
# 
# fviz_screeplot(res.famd)
# fviz_famd_var(res.famd)

# PCA map with eigenvalues
results_env <- dummy_cols(data_env, select_columns = c("discrim", "religios", "env_mh", "residenc", "ther_vis",
                                                       "activ_yn", "divers", "school2_type", "satisfied_overall", "ther_vis", 
                                                       "talk", "military", "international"),
                          remove_first_dummy = TRUE)

env_pca <- PCA(results_env)  # apply PCA
env_pca$eig   # output matrix with eigenvalues

```

## CLM environmental regression
Variables for consideration: discrim + religios + env_mh + residenc + ther_vis + activ_yn + divers + school2_4yr + school2_cc + school2_ad + satisfied_overall + ther_vis + talk + military + international
```{r CLM environment}

clm_env <- data.v14 %>% select(si_type, discrim, religios, env_mh, residenc, ther_vis, activ_yn, divers, school2_type, 
                   satisfied_overall, ther_vis, talk, military, international)
clm_env <- NA_preproc(clm_env)
clm_env <- as.data.frame(clm_env)

str(clm_env)
set.seed(1000) 
 
index = sample(1:nrow(clm_env), 0.7*nrow(clm_env)) 

train_env = clm_env[index,] # Create the training data 
test_env = clm_env[-index,] # Create the test data

dim(train_env)
dim(test_env)

m_env <- clm(si_type ~ discrim + religios + env_mh + residenc + ther_vis + activ_yn + divers + school2_type + 
                   satisfied_overall + ther_vis + talk + military + international, data = clm_env)
summary(m_env)

# sure fx
env_data <- mod_env[["model"]]

# residual plot
env_data$resid <- resids(mod_env, method = c("jitter"),
    jitter.scale = c("response"))
env_data <- cbind(env_data, mod_env[["fitted.values"]])
ggplot(env_data, aes(x = `passive SI`, y = resid)) +
geom_point(color = "purple", alpha = 0.15) +
geom_smooth(method = "lm")

env_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))

env_data %>% 
  filter(si_type == "SA") %>% 
  ggplot(aes(x = resid)) + geom_density(color = "magenta")

```


# LIFE EVENTS
Candidate variables: ins_cover, gpa_sr, abuse_year, psyhx, isi_risk, binge_fr, fincur, drug_mar, drug_poly, drug_coc, drug_other, drug_met, drug_none, drug_opi, drug_her, drug_stim, drug_ect, assault_sex, assault_phys, assault_emo, isi_tot


```{r LE frame}
data.v15 <- data.v14
```

## Insurance coverage
```{r ins_cover}

# create one insurance var from related questions in 'ins_' module
data.v15 <- data.v15 %>% 
  mutate(ins_cover = case_when((ins_1 == 1 | ins_mh == 4 | ins_mh == 5 | ins_ade == 3) ~ 0,
                               (ins_1 == 0 | ins_mh == 1 | ins_mh == 2 | ins_ade == 2) ~ 1))
table(data.v15$ins_cover, useNA = "always")   # check output
#data.v15 = subset(data.v15, select= -c(ins_1, ins_mh, ins_ade, ins_inade)) # clean up var
```

## Illicit drug use
Creates illicit drug use var to aggregate all drug answers into one, create drug_poly for >1 drug (limitation -- 30 days reflection)
```{r drug_type}
data.v15$drug_mar[is.na(data.v15$drug_mar)] <- 0
data.v15$drug_coc[is.na(data.v15$drug_coc)] <- 0
data.v15$drug_her[is.na(data.v15$drug_her)] <- 0
data.v15$drug_met[is.na(data.v15$drug_met)] <- 0
data.v15$drug_stim[is.na(data.v15$drug_stim)] <- 0
data.v15$drug_ect[is.na(data.v15$drug_ect)] <- 0
data.v15$drug_opi[is.na(data.v15$drug_opi)] <- 0
data.v15$drug_other[is.na(data.v15$drug_other)] <- 0
data.v15$drug_none[is.na(data.v15$drug_none)] <- 0
  
# change drug_ var to numeric
data.v16 <- data.v15 %>% 
  mutate(drug_sum = as.numeric(drug_mar) + as.numeric(drug_coc) + as.numeric(drug_her) + as.numeric(drug_met) + as.numeric(drug_stim) + as.numeric(drug_ect) + 
           as.numeric(drug_opi) + as.numeric(drug_other)) %>%

  # If >1 drug is selected, create/add selection to 'drug_poly'
  mutate(drug_poly = ifelse(drug_sum > 1, 1, 0)) %>%
  
  # drug_[subtype] specification, check for sub 1/0
  mutate(drug_type = case_when(drug_poly == 1 ~ 9,
                               drug_mar == 1 ~ 1,
                               drug_coc == 1 ~ 2, 
                               drug_her == 1 ~ 3,
                               drug_met == 1 ~ 4,
                               drug_stim == 1 ~ 5,
                               drug_ect == 1 ~ 6,
                               drug_opi == 1 ~ 7,
                               drug_other == 1 ~ 8,
                               drug_none == 1 ~ 10)) %>%
  mutate(drug_type = factor(drug_type, levels=1:10,
                    labels=c("Marijuana", "Cocaine", "Heroin", "Meth", "Stimulants", "Ecstasy", "Opioids", "Other", "Poly use", "None")))
table(data.v16$drug_type)   # tabulate and check output
data.v16 = subset(data.v16, select= -c(drug_sum)) # clean up var

# sort into uppers and downers
data.v16 <- data.v16 %>%
  mutate(drug_type = case_when(drug_poly == 1 ~ 3,
                               drug_mar == 1 ~ 1,
                               drug_coc == 1 ~ 2, 
                               drug_her == 1 ~ 1,
                               drug_met == 1 ~ 2,
                               drug_stim == 1 ~ 2,
                               drug_ect == 1 ~ 2,
                               drug_opi == 1 ~ 7,
                               drug_other == 1 ~ 4,
                               drug_none == 1 ~ 0)) %>%
  mutate(drug_type = factor(drug_type, levels=0:4,
                    labels=c("None", "Downer", "Upper", "Poly", "Other")))
table(data.v16$drug_type)   # tabulate and check output
data.v16 = subset(data.v16, select= -c(drug_mar, drug_coc, drug_her, drug_poly, drug_met, drug_stim, drug_ect, drug_opi, drug_other)) # clean up var

```

## Current financial stressors
```{r fincur}
table(data.v16$fincur)
```

## ISI scale
** too small
```{r isi_tot}
table(data.v16$isi_1)
table(data.v16$isi_2)

```

## Abuse
Transformed: Abuse_year~ abuse reported in the last year only
```{r abuse_year}
table(data.v16$abuse_life)
table(data.v16$abuse_recent)

# create dichotomous variable called 'abuse_year' to reflect abuse that occurred in last 12 months
data.v17 <- data.v16 %>%
  mutate(abuse_year = case_when(abuse_recent == 1 ~ 1,
                                abuse_recent == 2 ~ 1,
                                abuse_recent == 3 ~ 1,
                                abuse_recent == 4 ~ 0,
                                abuse_recent == 5 ~ 0,
                                abuse_life == 0 ~ 0))
data.v17$abuse_year <- as.numeric(data.v17$abuse_year)
table(data.v17$abuse_year)   # check output
data.v17 = subset(data.v17, select= -c(abuse_life, abuse_recent)) # clean up var
```

## Recent assault history
```{r assault_[type]}

# set all assault NAs equal to 0
data.v17$assault_sex[is.na(data.v17$assault_sex)] <- 0
data.v17$assault_emo[is.na(data.v17$assault_emo)] <- 0
data.v17$assault_phys[is.na(data.v17$assault_phys)] <- 0

data.v17 <- data.v17 %>%
  mutate(assault_sex = case_when(assault_sex == 2 ~ 0,
                                 assault_sex == 1 ~ 1))

table(data.v17$assault_sex) # tabulate and check output
table(data.v17$assault_emo)
table(data.v17$assault_phys)

```

## Binging frequency
```{r binge_fr}
# set all binge_fr values equal to 0
data.v17$binge_fr_f[is.na(data.v17$binge_fr_f)] <- 0
data.v17$binge_fr_m[is.na(data.v17$binge_fr_m)] <- 0
data.v17$binge_fr_o[is.na(data.v17$binge_fr_o)] <- 0

# create one binge var
data.v18 <- data.v17 %>% 
  mutate(binge_fr = as.numeric(binge_fr_f) + as.numeric(binge_fr_m) + as.numeric(binge_fr_o)) %>%
  mutate(binge_fr = as.character(binge_fr))

table(data.v18$binge_fr, useNA = "always") # check output
data.v18 = subset(data.v18, select= -c(binge_fr_f, binge_fr_m, binge_fr_o)) # clean up var

```

## Recent Psychiatric Hospitalization, last 12 months
```{r psyhx}
# Create variable for any psychiatric hospitalization in the last 12 months
data.v18 <- data.v18 %>% 
  mutate(psyhx = case_when((prov_pes == 1 | prov_inp == 1 | prov_par == 1) ~ 1, 
                           (prov_pes == 0 & prov_inp == 0 & prov_par == 0) ~ 0,
                           ther_ever == 0 ~ 0))
table(data.v18$psyhx, useNA = "always")   # check output
data.v18 = subset(data.v18, select= -c(prov_pes, prov_inp, prov_par)) # clean up var

```

## Current GPA
```{r gpa_sr}
 table(data.v18$gpa_sr, useNA = "always")
```
 
## LE correlation
```{r LE correlation}

# Define life event var candidates for correlation matrix
data_le <- data.v18 %>% select(ins_cover, abuse_year, gpa_sr, aca_impa, fincur, assault_phys, 
                               binge_fr, assault_sex, assault_emo, deprawsc, psyhx, drug_type)
data <- data_le
sig <- 0.6

 # convert data to numeric in order to run correlations (i.e., convert to factor first to keep 
 # the integrity of the data - each value will become a number rather than turn into NA_
data_le <- data_le %>% mutate_if(is.character, as.factor)
data_le <-  data_le %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_le, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")

```
## EFA LE
```{r}
data_le <- data.v18 %>% select(ins_cover, abuse_year, gpa_sr, aca_impa, fincur, assault_phys, 
                               binge_fr, assault_sex, assault_emo, deprawsc, psyhx, drug_type)

# create corr matrix, with dummy vars
data_le <- data_le %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_le   # check output

results_le <- dummy_cols(data_le, select_columns = c("ins_cover", "abuse_year", "gpa_sr", "aca_impa", 
                                                     "fincur", "assault_phys", "binge_fr", "assault_sex", 
                                                     "assault_emo", "deprawsc", "psyhx", "drug_type"),
                         remove_first_dummy = TRUE)

le_pca <- PCA(results_le)  # apply PCA
le_pca$eig   # output matrix with eigenvalues
```

## CLM Life event regression
Regression variables: ins_cover, gpa_sr, aca_impa, abuse_year, psyhx, binge_fr, fincur, 
drug_type, assault_sex, assault_phys, assault_emo, isi_tot
```{r clm_LE}
clm_le <- data.v18 %>% select(si_type, ins_cover, abuse_year, gpa_sr, aca_impa, fincur, assault_phys, 
                              binge_fr, assault_sex, assault_emo, deprawsc, psyhx, drug_type)

clm_le <- NA_preproc(clm_le)
clm_le <- as.data.frame(clm_le)

str(clm_le)
set.seed(1000) 

index = sample(1:nrow(clm_le), 0.7*nrow(clm_le))

train_le = clm_le[index,] # Create the training data 
test_le = clm_le[-index,] # Create the test data

dim(train_le)
dim(test_le)

m_le <- clm(si_type ~., data = train_le)
summary(m_le)

# sure fx
le_data <- mod_le[["model"]]

# residual plot
le_data$resid <- resids(mod_le, method = c("jitter"),
     jitter.scale = c("response"))
le_data <- cbind(le_data, mod_le[["fitted.values"]])
ggplot(le_data, aes(x = `passive SI`, y = resid)) +
geom_point(color = "purple", alpha = 0.15) +
geom_smooth(method = "lm")

le_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))

le_data %>% 
 filter(si_type == "SA") %>%
 ggplot(aes(x = resid)) + geom_density(color = "magenta")

```


# DEFEAT
Candidate variables: aca_impa, failed, men_ill_ash, deprawsc


```{r Defeat frame}
data.v19 <- data.v18
```

## PHQ risk/severity
```{r PHQ}
table(data.v19$deprawsc, useNA = "always")
```

## Academic impact
```{r aca_imp}
table(data.v19$aca_impa, useNA = "always")
```

## Academic failure
```{r failed}
table(data.v19$failed, useNA = "always")

# sample too small in the elective, remove
data.v19 = subset(data.v19, select= -c(failed)) # clean up var

```

## Ashamed
n= 8916 ** too small
```{r men_ill_ash}
table(data.v19$men_ill_ash, useNA = "always")

# sample too small in the elective, remove
data.v19 = subset(data.v19, select= -c(men_ill_ash)) # clean up var
```

## Defeat correlation
```{r Defeat cor}

# Define defeat variable candidates for correlation matrix
data_def <- data.v19 %>% select(aca_impa, deprawsc)
data <- data_def
sig <- 0.6

 # convert data to numeric in order to run correlations (i.e., convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA_
data_def <- data_def %>% mutate_if(is.character, as.factor)
data_def <-  data_def %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_def, use = "pairwise.complete.obs")                               
corr[lower.tri(corr, diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                           # drop perfect correlations
corr <- as.data.frame(as.table(corr))                    #turn into a 3-column table
corr <- na.omit(corr)                                    #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                    #sort by highest correlation
print(corr)

```

## EFA defeat
```{r}
data_def <- data.v19 %>% select(aca_impa, deprawsc)

# create corr matrix, with dummy vars 
data_def <- data_def %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_def   # check output

results_def <- dummy_cols(data_def, select_columns = c("aca_impa", "deprawsc"),
                         remove_first_dummy = TRUE)

def_pca <- PCA(results_def)  # apply PCA
def_pca$eig   # output matrix with eigenvalues
```

## CLM defeat
```{r CLM defeat}

clm_def <- data.v19 %>% select(si_type, aca_impa, deprawsc)

clm_def <- NA_preproc(clm_def)
clm_def <- as.data.frame(clm_def)

str(clm_def)
set.seed(1000) 

index = sample(1:nrow(clm_def), 0.7*nrow(clm_def))

train_def = clm_def[index,] # Create the training data 
test_def = clm_def[-index,] # Create the test data

dim(train_def)
dim(test_def)

m_def <- clm(si_type ~., data = train_def)
summary(m_def)

# # sure fx
# le_data <- mod_le[["model"]]
# 
# # residual plot
# le_data$resid <- resids(mod_le, method = c("jitter"),
#      jitter.scale = c("response"))
# le_data <- cbind(le_data, mod_le[["fitted.values"]])
# ggplot(le_data, aes(x = `passive SI`, y = resid)) +
# geom_point(color = "purple", alpha = 0.15) +
# geom_smooth(method = "lm")
# 
# le_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))
# 
# le_data %>% 
#  filter(si_type == "SA") %>%
#  ggplot(aes(x = resid)) + geom_density(color = "magenta")

```


# THREATS TO SELF-MODERATION
Candidate variables: persist, anx_score, dep_secret, AAQ_tot, achieve4


```{r TSM frame}
data.v20 <- data.v19
```

## GAD7 total
```{r anx_score}
table(data.v20$anx_score, useNA = "always")
```

## Secretive MH
```{r dep_secret}
table(data.v20$dep_secret, useNA = "always")
```

## AAQ total (elective module)
```{r AAQ_tot}
# create AAQ_tot variable to sum up imputed AAQ questions
data.v20 <- data.v20 %>%
  mutate(AAQ_tot = as.numeric(AAQ_1) + as.numeric(AAQ_2) + as.numeric(AAQ_3) + as.numeric(AAQ_4) + as.numeric(AAQ_5) + as.numeric(AAQ_6) + as.numeric(AAQ_7))

table(data.v20$AAQ_tot, useNA = "always")
```

## Achievement
n = 594  ** too small
```{r achieve4}
table(data.v20$achieve4, useNA = "always")
```

## Persist
```{r persist}
table(data.v20$persist, useNA = "always")
```

## Threats to self-moderation correlation
```{r}

# Define TSM variable candidates for correlation matrix
data_tsm <- data.v20 %>% select(persist, anx_score, dep_secret, achieve4, AAQ_tot)
data <- data_tsm
sig <- 0.6

 # convert data to numeric in order to run correlations (i.e., convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA_
data_tsm <- data_tsm %>% mutate_if(is.character, as.factor)
data_tsm <-  data_tsm %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_tsm, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                          # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
  
```

##EFA TSM
```{r EFA TSM}
data_tsm <- data.v20 %>% select(persist, anx_score, dep_secret, achieve4, AAQ_tot)

# create corr matrix, with dummy vars 
data_tsm <- data_tsm %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_tsm   # check output

results_tsm <- dummy_cols(data_tsm, select_columns = c("persist", "anx_score", "dep_secret", "achieve4", "AAQ_tot"),
                         remove_first_dummy = TRUE)

tsm_pca <- PCA(results_tsm)  # apply PCA
tsm_pca$eig   # output matrix with eigenvalues
```

## CLM TSM
```{r CLM TSM}
data.v20$persist <- as.factor(data.v19$persist)
data.v20$anx_score <- as.integer(data.v19$anx_score)
data.v20$dep_secret <- as.factor(data.v19$dep_secret)
data.v20$AAQ_tot <- as.integer(data.v19$AAQ_tot)

clm_tsm <- data.v20 %>% select(si_type, persist, anx_score, dep_secret, AAQ_tot)

clm_tsm <- NA_preproc(clm_tsm)
clm_tsm <- as.data.frame(clm_tsm)

str(clm_tsm)
set.seed(1000)

index = sample(1:nrow(clm_tsm), 0.7*nrow(clm_tsm))

train_tsm = clm_tsm[index,] # Create the training data
test_tsm = clm_tsm[-index,] # Create the test data

dim(train_tsm)
dim(test_tsm)

m_tsm <- clm(si_type ~., data = train_tsm)
summary(m_tsm)

# sure fx
tsm_data <- mod_tsm[["model"]]

# residual plot
tsm_data$resid <- resids(mod_tsm, method = c("jitter"),
      jitter.scale = c("response"))
tsm_data <- cbind(tsm_data, mod_tsm[["fitted.values"]])
ggplot(tsm_data, aes(x = `passive SI`, y = resid)) +
geom_point(color = "purple", alpha = 0.15) +
geom_smooth(method = "lm")

tsm_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))

tsm_data %>% 
  filter(si_type == "SA") %>%
  ggplot(aes(x = resid)) + geom_density(color = "magenta")

```


# ENTRAPMENT
Candidate variables: dep_impa, gad7_impa, perneed_cur


```{r Entrap frame}
data.v21 <- data.v20
```

## Depression impact
```{r dep_impac}
table(data.v21$dep_impa, useNA = "always")
```

## Anxiety impact
```{r gad7_impa}
table(data.v21$gad7_impa, useNA = "always")
```

## Current needs
```{r perneed_cur}
table(data.v21$percneed_cur, useNA = "always")
```

## Entrapment correlation
```{r Entrapment corr}

# Define entrapment variable candidates for correlation matrix
data_entrap <- data.v21 %>% select(dep_impa, gad7_impa, percneed_cur)
data <- data_entrap
sig <- 0.6

 # convert data to numeric in order to run correlations (i.e., convert to factor first to keep the integrity of the data
 # each value will become a number rather than turn into NA_
data_entrap <- data_entrap %>% mutate_if(is.character, as.factor)
data_entrap <-  data_entrap %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_entrap, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                                   # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)

```

## EFA Entrapment
```{r CFA entrap}
data_entrap <- data.v21 %>% select(dep_impa, gad7_impa, percneed_cur)

# create corr matrix, with dummy vars 
data_entrap <- data_entrap %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_entrap   # check output

results_entrap <- dummy_cols(data_entrap, select_columns = c("dep_impa", "gad7_impa", "percneed_cur"),
                         remove_first_dummy = TRUE)

entrap_pca <- PCA(results_entrap)  # apply PCA
entrap_pca$eig   # output matrix with eigenvalues
```

##  CLM Entrapment regression
```{r clm_entrap}
clm_entrap <- data.v21 %>% select(si_type, dep_impa, gad7_impa, percneed_cur)

str(clm_entrap)
set.seed(1000) 

index = sample(1:nrow(clm_entrap), 0.7*nrow(clm_entrap)) 

train_entrap = clm_entrap[index,] # Create the training data 
test_entrap = clm_entrap[-index,] # Create the test data

dim(train_entrap)
dim(test_entrap)

m_entrap <- clm(si_type ~., data = train_entrap)
summary(m_entrap)

# sure fx
entrap_data <- mod_entrap[["model"]]

# residual plot
entrap_data$resid <- resids(mod_entrap, method = c("jitter"),
       jitter.scale = c("response"))
entrap_data <- cbind(entrap_data, mod_entrap[["fitted.values"]])
ggplot(entrap_data, aes(x = `passive SI`, y = resid)) + 
  geom_point(color = "purple", alpha = 0.15) + 
  geom_smooth(method = "lm")

entrap_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))
entrap_data %>% filter(si_type == "SI") %>% ggplot(aes(x = resid)) + geom_density(color = "magenta")


```

## StepAIC entrapment
```{r StepAIC entrapment}

models <- regsubsets(si_type_num ~., data = datatrain_entrap, nvmax = 5,
                     method = "seqrep")
summary(models)

# Set seed for reproducibility
set.seed(1000)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(si_type_num ~., data = datatrain_entrap,
                    method = "lmStepAIC", 
                    trControl = train.control,
                    trace = FALSE
                    )
# Model accuracy
step.model$results
# Final model coefficients
step.model$finalModel
# Summary of the model
summary(step.model$finalModel)
coef(step.model$finalModel, 4)
```


# MOTIVATIONAL MODERATORS
Candidate variables: flourish, belong2, belong8, belong9, achieve1, achieve14, achieve18, stig_pcv_2, stig_pcv_3, friend_devel, inf, brs_tot


```{r MM frame}
data.v22 <- data.v21
```

## Belonging
```{r belong}
table(data.v22$belong1)
table(data.v22$belong2)
table(data.v22$belong8)
table(data.v22$belong9)
```

## Achievement
** too small for inclusion in model
```{r achieve_1_14_18}
table(data.v22$achieve1, useNA = "always")
table(data.v22$achieve14, useNA = "always")
table(data.v22$achieve18, useNA = "always")

# sample too small in the elective, remove
data.v22 = subset(data.v22, select= -c(achieve1, achieve14, achieve18)) # clean up var
```

## Flourishing Scale
```{r flourish}
table(data.v22$flourish, useNA = "always")
```

## Informal Mental Health Support
```{r inf}
data.v22$inf_1[is.na(data.v22$inf_1)] <- 0
data.v22$inf_2[is.na(data.v22$inf_2)] <- 0
data.v22$inf_3[is.na(data.v22$inf_3)] <- 0
data.v22$inf_4[is.na(data.v22$inf_4)] <- 0
data.v22$inf_5[is.na(data.v22$inf_5)] <- 0
data.v22$inf_6[is.na(data.v22$inf_6)] <- 0
data.v22$inf_7[is.na(data.v22$inf_7)] <- 0

# change inf_ var to numeric
data.v22 <- data.v22 %>% 
  mutate(inf_sum = as.numeric(inf_1) + as.numeric(inf_2) + as.numeric(inf_3) + as.numeric(inf_4) + as.numeric(inf_5) +
                   as.numeric(inf_6) + as.numeric(inf_7)) %>%
  mutate(inf_multi = ifelse(inf_sum > 1, 1, 0)) %>%
  
  # merge all var into inf_type
  mutate(inf = case_when(inf_multi == 1 ~ 9,
                        inf_1 == 1 ~ 2,
                        inf_2 == 1 ~ 3,
                        inf_3 == 1 ~ 4,
                        inf_4 == 1 ~ 5,
                        inf_5 == 1 ~ 6,
                        inf_6 == 1 ~ 7,
                        inf_7 == 1 ~ 8,
                        inf_8 == 1 ~ 1))
table(data.v22$inf)

# clean up vars
data.v22 = subset(data.v22, select= -c(inf_1:inf_help, inf_sum))
```

## BRS
```{r BRS_tot}
data.v22 <- data.v22 %>%
  mutate(brs_tot = as.integer(BRS_1) + as.integer(BRS_2) + as.integer(BRS_3) + as.integer(BRS_4) + as.integer(BRS_5))
table(data.v22$brs_tot)
```

## Friendship development
```{r friend_devel}
table(data.v22$friend_devel, useNA = "always")

# sample too small in the elective, remove
#data.v22 = subset(data.v22, select= -c(friend_devel)) # clean up var
```

## Perceived MH as failure
```{r stig_pcv_2}
table(data.v22$stig_pcv_2, useNA = "always")
```

## Perceived MH tx as negative
```{r stig_pcv_3}
table(data.v22$stig_pcv_3, useNA = "always")
```

## Motivational Moderator correlation
```{r}

# Define MM variable candidates for correlation matrix
data_mm <- data.v22 %>% select(inf, flourish, belong1, belong2, belong8, belong9, stig_pcv_2, stig_pcv_3, brs_tot)
data <- data_mm
sig <- 0.6

 # convert data to numeric in order to run correlations (i.e., convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA_
data_mm <- data_mm %>% mutate_if(is.character, as.factor)
data_mm <-  data_mm %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(data_mm, use = "pairwise.complete.obs")                               
corr[lower.tri(corr,diag=TRUE)] <- NA                   # prepare to drop duplicates and correlations of 1    
corr[corr %in% c(-1, 1)] <- NA                                   # drop perfect correlations
corr <- as.data.frame(as.table(corr))                   #turn into a 3-column table
corr <- na.omit(corr)                                   #remove the NA values from above 
corr <- corr[order(-abs(corr$Freq)),]                   #sort by highest correlation
print(corr)
  
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- acast(corr, Var1~Var2, value.var="Freq")  
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")

```

## EFA MM
```{r EFA MM}
data_mm <- data.v22 %>% select(inf, flourish, belong1, belong2, belong8, belong9, stig_pcv_2, stig_pcv_3, brs_tot)

# create corr matrix, with dummy vars 
data_mm <- data_mm %>% 
  mutate_if(is.numeric, as.factor) %>% 
  dummy_cols(remove_first_dummy = TRUE) %>% 
  data.matrix()
data_mm   # check output

results_mm <- dummy_cols(data_mm, select_columns = c("inf", "flourish", "belong1", "belong2", "belong8",
                                                     "belong9", "stig_pcv_2", "stig_pcv_3", "brs_tot"),
                         remove_first_dummy = TRUE)

mm_pca <- PCA(results_mm)  # apply PCA
mm_pca$eig   # output matrix with eigenvalues
```

## CLM Motivational moderators regression
Variables for consideration:  inf, flourish, belong1, belong2, belong8, belong9, stig_pcv_2, stig_pcv_3, brs_tot
```{r CLM mm}

clm_mm <- data.v22 %>% select(si_type, inf, flourish, belong1, belong2, belong8, belong9, stig_pcv_2, stig_pcv_3, brs_tot)

clm_mm <- NA_preproc(clm_mm)
clm_mm <- as.data.frame(clm_mm)

str(clm_mm)
set.seed(1000) 

index = sample(1:nrow(clm_mm), 0.7*nrow(clm_mm))

train_mm = clm_mm[index,] # Create the training data 
test_mm = clm_mm[-index,] # Create the test data

dim(train_mm)
dim(test_mm)

m_mm <- clm(si_type ~., data = train_mm)
summary(m_mm)

# sure fx
mm_data <- mod_mm[["model"]]

# residual plot
mm_data$resid <- resids(mod_mm, method = c("jitter"),
       jitter.scale = c("response"))
mm_data <- cbind(mm_data, mod_mm[["fitted.values"]])
ggplot(mm_data, aes(x = `passive SI`, y = resid)) + 
  geom_point(color = "purple", alpha = 0.15) + 
  geom_smooth(method = "lm")

mm_data %>% group_by(si_type) %>% summarize(mean = mean(resid, na.rm = TRUE), sd = sd(resid, na.rm = TRUE))

mm_data %>% 
  filter(si_type == "SA") %>% 
  ggplot(aes(x = resid)) + 
  geom_density(color = "magenta")

```


# MISSINGNESS


## Absolute missingness, by school
```{r Structural missingness, by school}
## Absolute missingness by school module choice

data.v23 <- data.v22

# school structural analysis
var_names <- c(
  # diathesis
  "age", "sexual", "gender", "finpast", "sib_freq", "race", "body_sr", "ed_scoff", "audit_tot", "dx_dep", "dx_bi", "dx_psy", 
  "dx_ea", "dx_tr", "dx_sa", "dx_adhd", "dx_pers", "dx_none", "meds_sti", "meds_dep", "meds_psy", "meds_anx", "meds_mood",
  "meds_sle", "meds_oth", "meds_none", "meds_count", "disab_1", "alc_tx", "wcs_tot",
  # environment
  "env_mh", "fam_support_aca", "ther_vis", "divers", "satisfied_overall", "residenc", "talk", "religios", 
  "discrim", "school2_type",
  # life events
  "ins_cover", "gpa_sr", "assault_phys", "binge_fr", "assault_sex", "assault_emo", "psyhx", 
  "fincur", "drug_type", "drug_none",
  # defeat
  "aca_impa", "deprawsc", 
  # threats to self-moderation
  "persist", "anx_score", "dep_secret", "AAQ_tot", "achieve4",
  # entrapment
  "dep_impa", "gad7_impa", "percneed_cur",
  # motivational moderators
  "inf", "flourish", "belong1", "belong2", "belong8", "belong9", "stig_pcv_2", "stig_pcv_3", "BRS_tot")  
# "abuse_year" now working, "activ_yn

var_names <- sort(var_names)
# build table for absolute missingness based on school2
# create skeleton of missingness matrix
school2 <- count(data.v23, school2)
school2_ncol <- ncol(school2)

for (j in 1:length(var_names)) {
  school2[, school2_ncol + j] <- 999
}
# create column headers and vectors for school, total students, then each variable count
names <- c("school2_values", "total", var_names)
names(school2) <- names

# input values for each school missingness across all variables
for (j in 1:(length(var_names))) {
for (i in 1:nrow(school2)) {   # i corresponds to rows, j to columns
  school2[i, 2 + j] <- sum(!is.na(data.v23[data.v23$school2 == school2[[i,1]], var_names[j]]))
}
}
# add percentage in each column
school2_m <- as.matrix(school2[-1]) 
  
props <- school2_m[,2:ncol(school2_m)] / school2_m[,1]
school2 <- data.frame(school2[, 1:2], props)

# view missingness matrix
view(school2)
# save matrix
write_xlsx(school2, path = "sch2_missing.xlsx", col_names = TRUE, format_headers = TRUE)
```

## Missingness of each overall variable
```{r Missingness of each overall var}

# Create dataframe to store data 
missing_rates <- data.frame("variable" = var_names,
           "total" = rep(999, length(var_names)),
           "prop_present" = rep(999, length(var_names)))

# loop to replicate row for all var
for (i in 1:nrow(missing_rates)) {
  missing_rates[i, 2] = sum(!is.na(data.v23[[i]]))
  missing_rates[i, 3] = missing_rates[i, 2] / nrow(data.v23)
}
# view matrix
view(missing_rates)
# save matrix
write_xlsx(missing_rates, path = "overall_missing.xlsx", col_names = TRUE, format_headers = TRUE)
  
```

## Missingness of each var w/in each var
```{r Missingness between vars}

missing_dfs <- list()
for (k in 1:length(var_names)) {

#build table for absolute missingness based on school2
#create skeleton of missingness matrix
var <- count(data.v23, data.v23[var_names[k]])
var_ncol <- ncol(var)

for (j in 1:length(var_names)) {
  var[, var_ncol + j] <- 999
}
# create column headers and vectors for school, total students, then each variable count
names <- c(paste0(var_names[k], "_values"), "total", var_names)
names(var) <- names

# input values for each school missingness across all variables
for (j in 1:(length(var_names))) {
for (i in 1:nrow(var)) {   # i corresponds to rows, j to columns
  var[i, 2 + j] <- sum(!is.na(data.v23[data.v23[[var_names[k]]] == var[[i,1]], var_names[j]]))
}
}
# add percentage in each column
var_m <- as.matrix(var[-1]) 
props <- var_m[,2:ncol(var_m)] / var_m[,1]
var <- data.frame(var[, 1:2], props)
replace_na("NA")
missing_dfs[[k]] <- var
}

names(missing_dfs) <- var_names
save(missing_dfs, file = "missing_dfs.RData")
load("missing_dfs.RData")
write_xlsx(missing_dfs, path = "missing.xlsx", col_names = TRUE, format_headers = TRUE)

```

## Missingness regression
```{r correlation and missingness rates}

# missing_mods <- list()
# for (i in 1:length(missing_dfs)) {
#   for (j in 1:(ncol(missing_dfs[[1]]) - 2)) {
#     mod <- glm(missing_dfs[[i]][,j + 2] ~ missing_dfs[[i]][,1])
#     missing_mods[[i]] <- coef(mod)
#   }
# }
# names(missing_mods) <- names(missing_dfs)
```


# DATA DIVISION


```{r Data division}

set.seed(1000) 
split = sample.split(data.v23$school2, SplitRatio = 0.50)
 
# Create training and testing sets
traindf.v23 <- subset(data.v23, split == TRUE)
testdf.v23 <- subset(data.v23, split == FALSE)
 
#Check output
dim(data.v23)
dim(traindf.v23)
dim(testdf.v23)
```


# SEM ANALYSIS


```{r SEM}

SI_model_tr <- '# measurement model
                dia_sem =~ age + sexual + gender + finpast + sib_freq + 
                    race + body_sr + ed_scoff + audit_tot + dx_dep + 
                    dx_bi + dx_psy + dx_ea + dx_tr + dx_sa + dx_adhd +
                    dx_pers + dx_none + meds_sti + meds_dep + meds_psy +
                    meds_anx + meds_mood + meds_sle + meds_oth + meds_none +
                    meds_count + disab_1 + alc_tx
                env_sem =~ env_mh + fam_support_aca + ther_vis + divers +
                    satisfied_overall + residenc + talk + religios + discrim +
                    school2_type
                le_sem =~ ins_cover + gpa_sr + assault_phys + binge_fr + assault_sex + 
                    assault_emo + psyhx + fincur + drug_type + drug_none
                def_sem =~ aca_impa + deprawsc
                tsm_sem =~ persist + anx_score + dep_secret + AAQ_tot
                entrap_sem =~ dep_impa + gad7_impa + perneed_cur
                mm_sm =~ flourish + belong + inf + stig_pcv_2 + stig_pcv_3 +
                    brs_tot
                si_sem =~ si_type
                    
                # regressions
                  # pre-motivational
                      dia_sem ~ env_sem
                      env_sem ~ le_sem
                      le_sem ~ dia_sem
                      def_sem ~ dia_sem + env_sem + le_sem
    # motivational
    entrap_sem ~ def_sem + tsm_sem
    si_sem ~ entrap_sem + mm_sem
    si_sem ~ def_sem + tsm_sem + entrap_sem + mm_sem
    
  # residual correlations
    med_count ~~ meds_sti + meds_dep + meds_psy + meds_anx + meds_mood + meds_sle + meds_oth + meds_none + meds_count
    dx_any ~~ dx_dep + dx_bi + dx_psy + dx_ea + dx_tr + dx_sa + dx_adhd + dx_pers + dx_non  '

fit_train <- sem(SI_model_tr, data=traindf.v23)
summary(fit_train, fit.measures=TRUE)

# test data comparator
SI_model_tr <- ' # measurement model
                  dia_sem =~ testdf.v23$age + testdf.v23$sexual + testdf.v23$gender + testdf.v23$finpast + testdf.v23$sib_freq + 
                    testdf.v23$race + testdf.v23$body_sr + testdf.v23$ed_scoff + testdf.v23$audit_tot + testdf.v23$dx_dep + 
                    testdf.v23$dx_bi + testdf.v23$dx_psy + testdf.v23$dx_ea + testdf.v23$dx_tr + testdf.v23$dx_sa + testdf.v23$dx_adhd +
                    testdf.v23$dx_pers + testdf.v23$dx_none + testdf.v23$meds_sti + testdf.v23$meds_dep + testdf.v23$meds_psy +
                    testdf.v23$meds_anx + testdf.v23$meds_mood + testdf.v23$meds_sle + testdf.v23$meds_oth + testdf.v23$meds_none +
                    testdf.v23$meds_count + testdf.v23$disab_1 + testdf.v23$alc_tx
                  env_sem =~ testdf.v23$env_mh + testdf.v23$fam_support_aca + testdf.v23$ther_vis + testdf.v23$divers +
                    testdf.v23$satisfied_overall + testdf.v23$residenc + testdf.v23$talk + testdf.v23$religios + testdf.v23$discrim +
                    testdf.v23$school2_type
                  le_sem =~ testdf.v23$ins_cover + testdf.v23$gpa_sr + testdf.v23$assault_phys + testdf.v23$binge_fr + testdf.v23$assault_sex + 
                    testdf.v23$assault_emo + testdf.v23$psyhx + testdf.v23$fincur + testdf.v23$drug_type + testdf.v23$drug_none
                  def_sem =~ testdf.v23$aca_impa + testdf.v23$deprawsc
                  tsm_sem =~ testdf.v23$persist + testdf.v23$anx_score + testdf.v23$dep_secret + testdf.v23$AAQ_tot
                  entrap_sem =~ testdf.v23$dep_impa + testdf.v23$gad7_impa + testdf.v23$perneed_cur
                  mm_sm =~ testdf.v23$flourish + testdf.v23$belong + testdf.v23$inf + testdf.v23$stig_pcv_2 + testdf.v23$stig_pcv_3 +
                    testdf.v23$brs_tot 
                
                # regressions
                    dia_sem ~ env_sem
                    env_sem ~ le_sem
                    le_sem ~ dia_sem
                    def_sem ~ dia_sem + env_sem + le_sem
                    entrap_sem ~ def_sem + tsm_sem
                    si_sem ~ entrap_sem + mm_sem
                    si_sem ~ def_sem + tsm_sem + entrap_sem + mm_sem
    
                # residual correlations
                    med_count ~~ testdf.v23$meds_sti + testdf.v23$meds_dep + testdf.v23$meds_psy + testdf.v23$meds_anx + testdf.v23$meds_mood +
                                 testdf.v23$meds_sle + testdf.v23$meds_oth + testdf.v23$meds_none + testdf.v23$meds_count
                    dx_any ~~ testdf.v23$dx_dep + testdf.v23$dx_bi + testdf.v23$dx_psy + testdf.v23$dx_ea + testdf.v23$dx_tr + testdf.v23$dx_sa +
                                 testdf.v23$dx_adhd + testdf.v23$dx_pers + testdf.v23$dx_non  '


# Training data analyses for indirect effects
set.seed(1000)
X <- rnorm(100)
M <- 0.5*X + rnorm(100)  # what are X, Y, M??????
Y <- 0.7*M + rnorm(100)
Data <- data.frame(X = X, Y = Y, M = M)
model <- ' # direct effect
           Y ~ c*X
           # mediator
           M ~ a*X
           Y ~ b*M
           # indirect effect (a*b)
           ab := a*b
           # total effect
           total := c + (a*b)  '

fit_train <- sem(model, data=data.v23)
summary(fit_train, fit.measures=TRUE)


fit_test <- sem(SI_model_tr, data=testdf.v23)
summary(fit_test, fit.measures=TRUE)
```


# MODIFICATION INDICES


```{r mod_indices}

mi_train <- modindices(fit_train, sort = TRUE, maximum.number = 5)  # might need to use lavTestScore() function

mi_test <- modindices(fit_test, sort = TRUE, maximum.number = 5)

```



```{r FAMD}
res.famd <- FAMD(wine[,c(1,2, 16, 22, 29, 28, 30,31)], graph = FALSE)
 
 
 
```


# PARAMETER ESTIMATES


```{r parameter_est}

#train
parameterEstimates(fit_train)
fitted(fit_train)
resid(fit_train)
fitMeasures(fit_train, c("cfi","rmsea","srmr"))
lavInspect(fit_train, what = "start")
lavInspect(fit_train, what = "list")

# test
parameterEstimates(fit_test)
fitted(fit_test)
resid(fit_test)
fitMeasures(fit_test, c("cfi","rmsea","srmr"))
lavInspect(fit_test, what = "start")
lavInspect(fit_test, what = "list")

```


```{r parameter_est}
